### 2015-10-31 models/priors-consortium.Rmd
Print posterior over parameters
Posterior predictives still need some thought.
How do you get out a posterior predictive for the dirichlet?

The posterior over parameters examined using 50k samples are only vaguely reasonable.
The independence between states (dirichlet prior) may not be valid.

Also, still need to analyze the human data to be able to compare it to the posterior predictive.

---
title: "priors-consortium"
author: "mht"
date: "October 31, 2015"
output: html_document
---

```{r helpers}

library(coda)
library(hydroGOF)
estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}
HPDhi<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}
HPDlo<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
options("scipen"=10)   
```

# Human data

```{r data.load}
d.number <- read.csv("~/Documents/research/priors/prag/data/number_dat.csv")
d.binned <- read.csv("~/Documents/research/priors/prag/data/bin_dat_noZeros.csv")
d.choice <- read.csv("~/Documents/research/priors/prag/data/choice_dat.csv")
```

```{r data.number}
qplot(data=d.number, x = chosen_bin, geom='histogram')+
  facet_wrap(~tag)

```

```{r data.binned}
d.bin.tidy<- d.binned %>%
  group_by(tag, bin_num) %>%
  multi_boot_standard(column="nresponse")

ggplot(d.bin.tidy, aes(x=bin_num, y = mean))+
  geom_bar(stat='identity', position=position_dodge())+
  geom_errorbar(aes(ymin=ci_lower, ymax = ci_upper))+
  facet_wrap(~tag)
```

```{r data.choice}

```

# Model

```{r model.load}
fpath<- "~/Documents/research/priors/models/results/"
prefix<- "priors-base-IncrMH"
samples = 50000
burn = samples /2
m<-read.csv(paste(fpath, prefix, samples, "burn", burn,".csv", sep=''))
m.samples <- m[rep(row.names(m), m$Probability*samples), 1:18]
```

# Posterior over parameters

## Data analysis parameters

```{r posterior.parameters}

m.params <- m.samples %>%
  filter(Type=='parameter')

ggplot(m.params, aes(x=Value))+
  geom_histogram()+
  facet_wrap(~Parameter, scales='free')

```

## Parameters governing the big prior in the sky

```{r}
m.pp <- m.samples %>%
  filter(Type == 'posterior' & Parameter=='priorVector') %>%
  gather(Bin, Value, starts_with("X")) %>%
  group_by(Parameter, Item, Bin) %>%
  summarize(MAP = estimate_mode(Value),
            credHigh = HPDhi(Value),
            credLow = HPDlo(Value)) %>%
  mutate(Bin = as.numeric(as.character(gsub("X","",Bin))))

ggplot(m.pp, aes(x=Bin, y=MAP))+
  geom_bar(stat='identity', position=position_dodge())+
  geom_errorbar(aes(ymin=credLow, ymax=credHigh))+
  facet_wrap(~Item)

```

### 2016-01-25 prag_alt/helpers/helpers.R
dev folders for prag model

@@ -1,0 +1,224 @@
+library('dplyr')
+library('ggplot2')
+library('reshape2')
+
+logit = function(x, k = 1) k * log(x / (1-x))
+logistic = function(x, k = 1, zero = 0) 1 / (1 + exp(-1/k*(x - zero)))
+add.margin <- function(x, eps = 0.000001) (1 - 2*eps)*x + eps 
+# calibrate <- function(x) (x - min(x)) / (max(x) - min(x))
+normalize <- function(x) (x / sum(x))
+
+
+# Computes highest density interval from a sample of representative values,
+#   estimated as shortest credible interval.
+# Arguments:
+#   sampleVec
+#     is a vector of representative values from a probability distribution.
+#   credMass
+#     is a scalar between 0 and 1, indicating the mass within the credible
+#     interval that is to be estimated.
+# Value:
+#   HDIlim is a vector containing the limits of the HDI
+HDIofMCMC = function( sampleVec , credMass = 0.95 ) {
+  sortedPts = sort( sampleVec )
+  ciIdxInc = ceiling( credMass * length( sortedPts ) )
+  nCIs = length( sortedPts ) - ciIdxInc
+  ciWidth = rep( 0 , nCIs )
+  for ( i in 1:nCIs ) {
+    ciWidth[ i ] = sortedPts[ i + ciIdxInc ] - sortedPts[ i ]
+  }
+  HDImin = sortedPts[ which.min( ciWidth ) ]
+  HDImax = sortedPts[ which.min( ciWidth ) + ciIdxInc ]
+  HDIlim = c( HDImin , HDImax )
+  return( HDIlim )
+}
+
+
+# get the all the samples into a data.frame
+# first get the parameters, and the the posterior predictive values
+# otherwise it messes with the number of bins; until now: only sliderPPC
+construct_ppvs <- function(samples, ppv = 'y.sliderPPC') {
+  slist <- samples$BUGSoutput$sims.list
+  ppv_df <- slist[[ppv]]
+  ppv_df <- tbl_df(melt(ppv_df))
+  
+  # poor man's solution as of yet
+  if (ppv == 'y.sliderPPC') {
+    bins <- rep(bin_dat$bin_num, each = max(ppv_df$Var1))
+    items <- rep(bin_dat$tag, each = max(ppv_df$Var1))
+    workerid = rep(bin_dat$workerid, each = max(ppv_df$Var1))
+    ppv_df <- ppv_df %>%
[…]

### 2016-01-25 prag_alt/main.R

@@ -1,7 +1,4 @@
 library('R2jags')
-library('rjags')
-library('runjags')
-library('ggmcmc')
 source('helpers/helpers.R')
 
 
@@ -41,17 +38,21 @@ data_choice <- with(choice_dat,
 data_aggr <- c(data_gen, data_slider, data_number, data_choice)
 params <- c('w', 
             'a', 
-            'sigma', 
-            'item.pop', 
+            'tau', 
+             'item.pop', 
             'k.skewGlobal', 
             'b',
-            "y.choicePPC", "y.numberPPC", "y.sliderPPC")
+            'y.sliderPPC',
+            'y.numberPPC') 
+            # 'y.choicePPC')
+            
 
 samples <- jags(data_aggr, parameters.to.save = params,
-                model.file = 'models/model.txt', n.chains = 2, n.iter = 500, 
-                n.burnin = 300, n.thin = 2, DIC = FALSE)
+                model.file = 'models/model.txt', n.chains = 2, n.iter = 50000, 
+                n.burnin = 30000, n.thin = 2, DIC = FALSE)
 
-stop()
+mys = samples$BUGSoutput$sims.list
+mysDF = as.data.frame(mys)
 
 csamples <- clean_samples(samples, ppv = 'none')
 

### 2016-01-25 prag_alt/main2.R

@@ -1,14 +1,22 @@
-library('R2jags')
-library('rjags')
-library('runjags')
-library('ggmcmc')
+# library('R2jags')
+library(jagsUI) # for parallel computing
+library(ggmcmc)
+library(coda)
+library(BEST)
 source('helpers/helpers.R')
+source('~/Desktop/data/svn/ProComPrag/dev_tmp/typicality_quantifiers/model/helpers.r')
 
 
 dat <- read.csv('data/exp1.csv')
 bin_dat <- read.csv('data/bin_dat.csv')
 choice_dat <- read.csv('data/choice_dat.csv')
 number_dat <- read.csv('data/number_dat.csv')
+
+# focus = "coffee"
+# dat = droplevels(filter(dat, tag == focus))
+# bin_dat = droplevels(filter(bin_dat, tag == focus))
+# choice_dat = droplevels(filter(choice_dat, tag == focus))
+# number_dat = droplevels(filter(number_dat, tag == focus))
 
 # data useful for all dependent measures
 data_gen <- with(dat,
@@ -17,7 +25,7 @@ data_gen <- with(dat,
 
 # data specific for the slider measure
 data_slider <- with(bin_dat, # use normalized response
-               list('y.slider' = logit(add.margin(nresponse)),
+               list('y.slider' = response,
                     'bin.num' = bin_num,
                     'item.slider' = as.numeric(tag), 
                     'worker.slider' = workerid + 1,
@@ -43,15 +51,43 @@ params <- c('w',
             'a', 
             'sigma', 
             'item.pop', 
-            'k.skewGlobal', 
-            'b',
-            "y.choicePPC", "y.numberPPC", "y.sliderPPC")
+            'k.skewGlobal',
+            'threshold',
+            'b'
+            #'y.sliderPPC',
+            #'y.numberPPC'
+            # 'y.choicePPC'
+            )
+burnin = 5000
+iter = 5000
+out <- jags(data = data_aggr,
+            inits = NULL,
+            parameters.to.save = params,
+            codaOnly = c('y.sliderPPC', 'y.numberPPC', 'y.choicePPC'),
+            model.file = 'models/model.jags.R',
+            n.chains = 2,
+            n.adapt = 1000,
+            n.iter = iter + burnin,
+            n.burnin = burnin,
+            n.thin = 2, 
+            DIC = TRUE,
[…]

### 2016-01-25 prag_alt/models/model.jags.R

@@ -2,16 +2,16 @@ model {
     ## SLIDER
     for (i in 1:N.slider) {
         p[i] <- subj[item.slider[i], worker.slider[i], bin.num[i]]
-        y[i] <- k.skew[worker.slider[i]] * log(p[i] / (1 - p[i]))
-        y.slider[i] ~ dnorm(y[i], tau)
-        y.sliderPPC[i] ~ dnorm(y[i], tau)
+        pred[i] = 1/ (1 + exp(-k.skewGlobal*(log(p[i]/(1-p[i]))-threshold)))
+        y.slider[i] ~ dnorm(pred[i], tau) T(0,1)
+        # y.sliderPPC[i] ~ dnorm(y[i], tau)
     }
     
     ## NUMBERS
-    for (i in 1:N.number) {
-        y.number[i] ~ dcat(pow(subj[item.number[i], worker.number[i], 1:15], a))
-        y.numberPPC[i] ~ dcat(pow(subj[item.number[i], worker.number[i], 1:15], a))
-    }
+   for (i in 1:N.number) {
+       y.number[i] ~ dcat(pow(subj[item.number[i], worker.number[i], 1:15], a))
+       y.numberPPC[i] ~ dcat(pow(subj[item.number[i], worker.number[i], 1:15], a))
+   }
 
     ## CHOICE
     for (i in 1:N.choice) {
@@ -33,7 +33,7 @@ model {
         prob[i] <- probT[i,2] / sum(probT[i,])
         
         y.choice[i] ~ dbern(prob[i])                            
-        y.choicePPC[i] ~ dbern(prob[i])
+#         y.choicePPC[i] ~ dbern(prob[i])
                               
     }
     
@@ -41,8 +41,9 @@ model {
     w ~ dgamma(2, 0.1)
     a ~ dgamma(2, 1)
     b ~ dgamma(2, 1)
-    sigma ~ dunif(0, .00001) 
+    sigma ~ dgamma(.0001, .0001)
     tau <- 1/sigma^2
+    threshold ~ dnorm(0,1)
     
     # priors for logistic skew
     # for (i in 1:n.subj) { k.skew[i] ~ dgamma(2, 1) }
@@ -51,11 +52,11 @@ model {
     
     # population item priors
     for (j in 1:n.items) {
-        item.pop[1:15, j] ~ ddirch(ones[])
+        item.pop[j,1:15] ~ ddirch(ones[])
         
         # subject specific priors
[…]

### 2016-01-25 prag_alt/prepare.R

@@ -1,0 +1,149 @@
+library('rjson')
+library('dplyr')
+library('stringr')
+source('helpers/helpers.R')
+
+dat <- read.csv('data/exp1.csv')
+
+############
+# bin data #
+############
+bin_dat <- dat %>% 
+  filter(measure == 'binned_histogram') %>%
+  mutate(response = as.numeric(as.character(response)))
+
+all_bins <- as.character(unique(bin_dat$bins))
+all_bins <- c(all_bins, all_bins[1])
+all_bins <- lapply(all_bins, function(str) {
+  str <- gsub("u", "", str)
+  str <- gsub("'", "\"", str)
+  str <- gsub(".xb0", "", str)
+  fromJSON(str)
+})
+names(all_bins) = c("joke", "movies", "tv", "coffee", "watch", "commute", "laptop", "marbles")
+
+bin_dat$bin_num = sapply(1:nrow(bin_dat), function(i) {
+  bin = as.character(bin_dat$bin)[i]
+  tag = as.character(bin_dat$tag)[i]
+  which(all_bins[[tag]] == bin)
+})
+
+bin_dat <- select(bin_dat, workerid, tag, bin_num, response)
+bin_dat <- bin_dat %>% group_by(tag, workerid) %>% mutate(nresponse = normalize(response)) %>% data.frame()
+# write.csv(bin_dat, file = "data/bin_dat.csv", row.names = FALSE )
+
+###############
+# number data #
+###############
+number_dat <- dat %>%
+  filter(measure == 'give_number') %>% 
+  mutate(response = as.numeric(as.character(response)))
+
+steps <- list('coffee' = 13, 'movies' = 16, 'tv' = 3, 'commute' = 7, 'laptop' = 500,
+              'watch' = 50, 'marbles' = 1, 'joke' = 1)
+
+# TEST IF THIS WORKS! edge cases: last bin (or more), first bin (0-3, 44 or less etc.)
+number_dat$chosen_bin <- sapply(1:nrow(number_dat), function(i) {
+  res <- number_dat$response[i]
+  tag <- as.character(number_dat$tag)[i]
+  bins <- all_bins[[tag]]
+  names(bins) <- 1:15
+  bins <- regmatches(bins, gregexpr('\\(?[0-9,.]+', bins)) # remove everything except numbers
+  bins.log <- numeric(15)
+  
+  # ugh
+  flag <- FALSE
[…]

### 2016-01-25 prag_oldDev/helpers/helpers.R

@@ -1,0 +1,216 @@
+library('dplyr')
+library('ggplot2')
+library('reshape2')
+
+
+
+logit = function(x, k = 1) k * log(x / (1-x))
+logistic = function(x, k = 1, zero = 0) 1 / (1 + exp(-1/k*(x - zero)))
+add.margin <- function(x, eps = 0.000001) (1 - 2*eps)*x + eps 
+# calibrate <- function(x) (x - min(x)) / (max(x) - min(x))
+normalize <- function(x) (x / sum(x))
+
+
+# Computes highest density interval from a sample of representative values,
+#   estimated as shortest credible interval.
+# Arguments:
+#   sampleVec
+#     is a vector of representative values from a probability distribution.
+#   credMass
+#     is a scalar between 0 and 1, indicating the mass within the credible
+#     interval that is to be estimated.
+# Value:
+#   HDIlim is a vector containing the limits of the HDI
+HDIofMCMC = function( sampleVec , credMass = 0.95 ) {
+  sortedPts = sort( sampleVec )
+  ciIdxInc = ceiling( credMass * length( sortedPts ) )
+  nCIs = length( sortedPts ) - ciIdxInc
+  ciWidth = rep( 0 , nCIs )
+  for ( i in 1:nCIs ) {
+    ciWidth[ i ] = sortedPts[ i + ciIdxInc ] - sortedPts[ i ]
+  }
+  HDImin = sortedPts[ which.min( ciWidth ) ]
+  HDImax = sortedPts[ which.min( ciWidth ) + ciIdxInc ]
+  HDIlim = c( HDImin , HDImax )
+  return( HDIlim )
+}
+
+
+# get the all the samples into a data.frame
+# first get the parameters, and the the posterior predictive values
+# otherwise it messes with the number of bins; until now: only sliderPPC
+construct_ppvs <- function(samples, ppv = 'y.sliderPPC') {
+  slist <- samples$BUGSoutput$sims.list
+  ppv_df <- slist[[ppv]]
+  ppv_df <- tbl_df(melt(ppv_df))
+  
+  # poor man's solution as of yet
+  if (ppv == 'y.sliderPPC') {
+    bins <- rep(bin_dat$bin_num, each = max(ppv_df$Var1))
+    items <- rep(bin_dat$tag, each = max(ppv_df$Var1))
+    workerid = rep(bin_dat$workerid, each = max(ppv_df$Var1))
+    ppv_df <- ppv_df %>%
[…]

### 2016-01-25 prag_oldDev/main.R

@@ -1,44 +1,20 @@
-library('R2jags')
-library('rjags')
-library('runjags')
+library('coda')
 library('ggmcmc')
+library('jagsUI') # for parallel computing
 source('helpers/helpers.R')
+source('~/Desktop/data/svn/ProComPrag/dev_tmp/typicality_quantifiers/model/helpers.r')
+source('process_data.R')
 
 
-dat <- read.csv('data/exp1.csv')
-bin_dat <- read.csv('data/bin_dat.csv')
-choice_dat <- read.csv('data/choice_dat.csv')
-number_dat <- read.csv('data/number_dat.csv')
+saveFlag = FALSE
 
-# data useful for all dependent measures
-data_gen <- with(dat,
-            list('n.items' = nlevels(tag), 'ones' = rep(1, 15),
-                 'n.subj' = length(unique(workerid))))
 
-# data specific for the slider measure
-data_slider <- with(bin_dat, # use normalized response
-               list('y.slider' = logit(add.margin(nresponse)),
-                    'bin.num' = bin_num,
-                    'item.slider' = as.numeric(tag), 
-                    'worker.slider' = workerid + 1,
-                    'N.slider' = nrow(bin_dat)))
 
-# data specific for the number measure
-data_number <- with(number_dat,
-               list('y.number' = chosen_bin, 
-                    'item.number' = as.numeric(tag),
-                    'worker.number' = workerid + 1,
-                    'N.number' = nrow(number_dat)))
 
-# data specific for the choice measure
-data_choice <- with(choice_dat,
-               list('y.choice' = as.numeric(higher_chosen),
-                    'item.choice' = as.numeric(as.factor(tag)),
-                    'N.choice' = nrow(choice_dat), 'worker.choice' = workerid + 1,
-                    'higher' = sapply(1:nrow(choice_dat), function(i) max(chosen_bin[i], unchosen_bin[i])),
-                    'lower' = sapply(1:nrow(choice_dat), function(i) min(chosen_bin[i], unchosen_bin[i]))))
-
-data_aggr <- c(data_gen, data_slider, data_number, data_choice)
+data_aggr <- list(y.slider = y.slider, y.numer = y.number, y.choice = y.choice, 
+                  higher = higher, lower = lower,
+                  nSubjs = 20, nItems = 8, nBins = 15, nLightConds = 5,
+                  ones = rep(1,15))
 params <- c('w', 
[…]

### 2016-01-25 prag_oldDev/models/model.jags.R

@@ -1,0 +1,60 @@
+model {
+  ## SLIDERS
+  for (i in 1:nSubjs) {
+    for (j in 1:nItems) {
+      for (k in 1:nBins){
+        y[i,j,k] <- k.skew[i] * log(subj[i,j,k] / (1 - subj[i,j,k]))
+        y.slider[i,j,k]    ~ dnorm(y[i,j,k], tau)
+        y.sliderPPC[i,j,k] ~ dnorm(y[i,j,k], tau)   
+      }
+    }
+  }
+  ## NUMBERS
+  for (i in 1:nSubjs) {
+    for (j in 1:nItems) {
+        y.number[i,j,1]    ~ dcat(pow(subj[i,j, 1:15], a))
+        y.numberPPC[i,j,1] ~ dcat(pow(subj[i,j, 1:15], a))
+    }
+  }
+
+  ## CHOICE
+  for (i in 1:nSubjs) {
+    for (j in 1:nItems) {
+      for (b in 1:15) {
+        pTemp[i,j,b] <- exp( 1000 * subj[i,j,b] )
+        pTempNorm[i,j,b] = pTemp[i,j,b] / sum(pTemp[i,j,1:15])
+      }
+      mode[i,j] ~ dcat(pTempNorm[i,j,1:15])
+      for (k in 1:nLightConds) {
+        # difference between choice alternatives and modal bin
+        d.lo[i,j,k] <- abs(mode[i,j] - lower[i,j,k])
+        d.hi[i,j,k] <- abs(mode[i,j] - higher[i,j,k])
+        probT[i,j,k,1] <- exp(b * (15 - d.lo[i,j,k]))
+        probT[i,j,k,2] <- exp(b * (15 - d.hi[i,j,k]))
+        prob[i,j,k] <- probT[i,j,k,2] / sum(probT[i,j,k,1:2])
+        y.choice[i,j,k] ~ dbern(prob[i,j,k])                            
+        y.choicePPC[i,j,k] ~ dbern(prob[i,j,k])
+      }
+    }
+  }
+    # noise parameter
+    w ~ dgamma(2, 0.1)
+    a ~ dgamma(2, 1)
+    b ~ dgamma(2, 1)
+    sigma ~ dgamma(.0001, .0001)
+    tau <- 1/sigma^2
+    
+    # priors for logistic skew
+    # for (i in 1:n.subj) { k.skew[i] ~ dgamma(2, 1) }
+    k.skewGlobal ~ dgamma(5,5)
[…]

### 2016-01-25 prag_oldDev/plotting.R

@@ -1,0 +1,45 @@
+library('coda')
+library('ggmcmc')
+library('jagsUI') # for parallel computing
+source('helpers/helpers.R')
+source('~/Desktop/data/svn/ProComPrag/dev_tmp/typicality_quantifiers/model/helpers.r')
+source('process_data.R')
+
+readFlag = TRUE
+savePlots = FALSE
+
+if (readFlag){
+  load(file = "~/Desktop/Dropbox/priors_data/out25.Rdat")
+} else {
+  source('main.R')
+}
+
+
+csamples = tbl_df(melt(out$sims.list))
+colnames(csamples) = c("value", "step", "subject", "item", "bin", "variable")
+csamples$item = levels(factor(bin_dat$tag))[csamples$item]
+
+slider_ppv = filter(csamples, variable == "y.sliderPPC") %>% 
+  mutate(value = logistic(value)) %>%
+  group_by(subject, item, step) %>%
+  mutate(nvalue = value / sum(value))
+
+slider_aggr = slider_ppv %>% group_by(item, bin, step) %>%
+  summarise(y.rep.mean = mean(nvalue)) %>%
+  group_by(item, bin) %>%
+  summarise(mean = mean(y.rep.mean),
+            min = HDIofMCMC(y.rep.mean)[1],
+            max = HDIofMCMC(y.rep.mean)[2])
+slider_aggr = slider_aggr[order(slider_aggr$item,slider_aggr$bin),]
+slider_aggr$y_means = y.slider_means$mymean
+
+
+plotSliderPPC = ggplot(slider_aggr, aes(x = bin, y = mean)) + geom_line() + geom_point() + facet_wrap(~ item, scale = "free") + 
+  geom_errorbar(aes(ymin = min, ymax = max), width = .5, position = position_dodge(.1), color = 'gray') +
+  geom_line(aes(x = bin, y = y_means) , color = "red") + geom_point( aes(x = bin, y = y_means) , color = "red")
+
+show(plotSliderPPC)
+
+if (savePlots){
+  ggsave('plots/ppc_slider.pdf', ppc_slider, width=10, height = 8)
+}

### 2016-01-25 prag_oldDev/process_data.R

@@ -1,0 +1,42 @@
+dat <- read.csv('data/exp1.csv')
+bin_dat <- read.csv('data/bin_dat.csv')
+choice_dat <- read.csv('data/choice_dat.csv')
+number_dat <- read.csv('data/number_dat.csv')
+
+## slider rating data
+y.slider = array(0, dim = c(20,8,15))
+for (i in 1:nrow(bin_dat)){
+  y.slider[bin_dat$workerid[i] + 1, 
+           as.numeric(bin_dat$tag)[i],
+           bin_dat$bin_num[i]] = bin_dat$nresponse[i]
+}
+y.slider = logit(add.margin(y.slider))
+y.slider_means = bin_dat %>% group_by(bin_num, tag) %>% 
+  summarise(mymean = mean(nresponse)) %>% rename(bin = bin_num, item = tag)
+y.slider_means = y.slider_means[order(y.slider_means$item,y.slider_means$bin),]
+
+# number choice data
+y.number = array(0, dim = c(20,8,1))
+for (i in 1:nrow(number_dat)){
+  y.number[number_dat$workerid[i] + 1, 
+           as.numeric(number_dat$tag)[i],
+           1] = number_dat$chosen_bin[i]
+}
+
+## lightning round choice data
+choice_dat = choice_dat %>% mutate(condition = factor(ifelse(chosen_bin > unchosen_bin, unchosen_bin, chosen_bin)))
+choice_dat$condition = as.numeric(choice_dat$condition)
+y.choice = array(0, dim = c(20,8,5))
+higher   = array(0, dim = c(20,8,5))
+lower    = array(0, dim = c(20,8,5))
+for (i in 1:nrow(choice_dat)){
+  y.choice[choice_dat$workerid[i] + 1, 
+           as.numeric(choice_dat$tag)[i],
+           choice_dat$condition[i]] = ifelse(choice_dat$higher_chosen[i] , 1, 0)
+  higher[choice_dat$workerid[i] + 1, 
+         as.numeric(choice_dat$tag)[i],
+         choice_dat$condition[i]] = max(choice_dat$chosen_bin[i], choice_dat$unchosen_bin[i])
+  lower[choice_dat$workerid[i] + 1, 
+        as.numeric(choice_dat$tag)[i],
+        choice_dat$condition[i]] = min(choice_dat$chosen_bin[i], choice_dat$unchosen_bin[i])
+}

### 2016-01-26 prag_oldDev/plotting.R
plotting PPCs

@@ -9,21 +9,26 @@ readFlag = TRUE
 savePlots = FALSE
 
 if (readFlag){
-  load(file = "~/Desktop/Dropbox/priors_data/out25.Rdat")
+  # load(file = "/Users/micha/Desktop/Dropbox/priors_data/out_25.Rdat")
+  load(file = "/Users/micha/Desktop/Dropbox/priors_data/out.Rdat")
 } else {
   source('main.R')
 }
 
-
+# prepare samples
 csamples = tbl_df(melt(out$sims.list))
 colnames(csamples) = c("value", "step", "subject", "item", "bin", "variable")
 csamples$item = levels(factor(bin_dat$tag))[csamples$item]
 
+#######################
+# posterior predictives
+#######################
+
+# sliders
 slider_ppv = filter(csamples, variable == "y.sliderPPC") %>% 
   mutate(value = logistic(value)) %>%
   group_by(subject, item, step) %>%
   mutate(nvalue = value / sum(value))
-
 slider_aggr = slider_ppv %>% group_by(item, bin, step) %>%
   summarise(y.rep.mean = mean(nvalue)) %>%
   group_by(item, bin) %>%
@@ -32,14 +37,35 @@ slider_aggr = slider_ppv %>% group_by(item, bin, step) %>%
             max = HDIofMCMC(y.rep.mean)[2])
 slider_aggr = slider_aggr[order(slider_aggr$item,slider_aggr$bin),]
 slider_aggr$y_means = y.slider_means$mymean
-
-
 plotSliderPPC = ggplot(slider_aggr, aes(x = bin, y = mean)) + geom_line() + geom_point() + facet_wrap(~ item, scale = "free") + 
   geom_errorbar(aes(ymin = min, ymax = max), width = .5, position = position_dodge(.1), color = 'gray') +
   geom_line(aes(x = bin, y = y_means) , color = "red") + geom_point( aes(x = bin, y = y_means) , color = "red")
-
 show(plotSliderPPC)
 
+# numbers
+number_dat$chosen_bin = factor(number_dat$chosen_bin, levels = 1:15)
+DNm = tbl_df(melt(table(number_dat$tag, number_dat$chosen_bin))) %>%
+  rename(item = Var1, bin = Var2)
+DNm = DNm[order(DNm$item, DNm$bin),]
+number_ppv = filter(csamples, variable == "y.numberPPC")
+number_ppv$value = factor(number_ppv$value)
[…]

### 2016-01-26 prag_oldDev/helpers/helpers.R
stable model; most plotting functionality

@@ -175,7 +175,7 @@ plot_populationPriors <- function(slider_ppv, aggr_slider) {
     geom_line(aes(x = bin, y = y_emp) , color = "red")  + geom_point(aes(x = bin, y = y_emp ), color = "red")
 }
 
-plot_parameters <- function(p = c("a", "b", "w", "tau", "k.skewGlobal")) {
+plot_parameters <- function(p = c("a", "b", "w", "sigma", "k.skewGlobal")) {
   meansIP <- csamples %>% filter(variable %in% p) %>%
     group_by(variable) %>%
     summarise(

### 2016-01-26 prag_oldDev/main.R

@@ -6,33 +6,34 @@ source('~/Desktop/data/svn/ProComPrag/dev_tmp/typicality_quantifiers/model/helpe
 source('process_data.R')
 
 
-saveFlag = FALSE
+saveFlag = TRUE
 
-
-
-
-data_aggr <- list(y.slider = y.slider, y.numer = y.number, y.choice = y.choice, 
+data_aggr <- list(y.slider = y.slider, y.number = y.number, y.choice = y.choice, 
                   higher = higher, lower = lower,
                   nSubjs = 20, nItems = 8, nBins = 15, nLightConds = 5,
                   ones = rep(1,15))
 params <- c('w', 
             'a', 
             'sigma', 
-            'item.pop', 
-            'k.skewGlobal', 
+            'item.pop',
+            'subj',
+            'k', 
             'b',
-            "y.choicePPC", "y.numberPPC", "y.sliderPPC")
+            "y.choicePPC", 
+            "y.numberPPC",
+            "y.sliderPPC"
+            )
 
 model = "models/model.jags.R"
-burnin = 2000
-iter = 2500
+burnin = 100000
+iter = 25000
 out = jags(data = data_aggr,
             inits = NULL,
             parameters.to.save = params,
-            codaOnly = c("y.choicePPC", "y.numberPPC", "y.sliderPPC"),
+            codaOnly = c("y.choicePPC", "y.numberPPC", "y.sliderPPC", "subj"),
             model.file = model,
             n.chains = 2,
-            n.adapt = 500,
+            n.adapt = 50000,
             n.iter = iter + burnin,
             n.burnin = burnin,
             n.thin = 2, 
@@ -40,48 +41,17 @@ out = jags(data = data_aggr,
             verbose = TRUE,
             parallel = TRUE)
 
-if (saveFlag) { save(out, file = "~/Desktop/Dropbox/priors_data/out_25.Rdat") }
-
+if (saveFlag) { save(out, file = "~/Desktop/Dropbox/priors_data/out.Rdat") }
 
 stop()
 
+p = out$sims.list$item.pop[1,1,]
 
-# construct all the posterior predictive samples; needs to be adjusted! (esp. slider ...)
-slider_ppv <- construct_ppvs(samples)
[…]

### 2016-01-26 prag_oldDev/models/model.jags.R

@@ -12,11 +12,15 @@ model {
   ## NUMBERS
   for (i in 1:nSubjs) {
     for (j in 1:nItems) {
-        y.number[i,j,1]    ~ dcat(pow(subj[i,j, 1:15], a))
-        y.numberPPC[i,j,1] ~ dcat(pow(subj[i,j, 1:15], a))
+      for (k in 1:nBins){
+        cp[i,j,k] = exp(a * subj[i,j,k])
+      }
+      y.number[i,j,1]    ~ dcat(cp[i,j,1:15])
+      y.numberPPC[i,j,1] ~ dcat(cp[i,j,1:15])
     }
   }
-
+  x[1:15] = cp[1,1,1:15]
+  
   ## CHOICE
   for (i in 1:nSubjs) {
     for (j in 1:nItems) {
@@ -27,13 +31,20 @@ model {
       mode[i,j] ~ dcat(pTempNorm[i,j,1:15])
       for (k in 1:nLightConds) {
         # difference between choice alternatives and modal bin
-        d.lo[i,j,k] <- abs(mode[i,j] - lower[i,j,k])
-        d.hi[i,j,k] <- abs(mode[i,j] - higher[i,j,k])
-        probT[i,j,k,1] <- exp(b * (15 - d.lo[i,j,k]))
-        probT[i,j,k,2] <- exp(b * (15 - d.hi[i,j,k]))
+        hi[i,j,k] = ifelse(abs(mode[i,j] - lower[i,j,k]) > abs(mode[i,j] - higher[i,j,k]), 1, 0) 
+        lo[i,j,k] = ifelse(abs(mode[i,j] - lower[i,j,k]) < abs(mode[i,j] - higher[i,j,k]), 1, 0)
+        probT[i,j,k,1] <- exp(b * (1 + lo[i,j,k] - hi[i,j,k]))
+        probT[i,j,k,2] <- exp(b * (1 + hi[i,j,k] - lo[i,j,k]))
         prob[i,j,k] <- probT[i,j,k,2] / sum(probT[i,j,k,1:2])
         y.choice[i,j,k] ~ dbern(prob[i,j,k])                            
         y.choicePPC[i,j,k] ~ dbern(prob[i,j,k])
+#         d.lo[i,j,k] <- abs(mode[i,j] - lower[i,j,k])
+#         d.hi[i,j,k] <- abs(mode[i,j] - higher[i,j,k])
+#         probT[i,j,k,1] <- exp(b * (15 - d.lo[i,j,k]))
[…]

### 2016-01-26 prag_oldDev/plotting.R

@@ -5,20 +5,76 @@ source('helpers/helpers.R')
 source('~/Desktop/data/svn/ProComPrag/dev_tmp/typicality_quantifiers/model/helpers.r')
 source('process_data.R')
 
+theme_set(theme_bw() + theme(plot.background=element_blank()) )
+
 readFlag = TRUE
-savePlots = FALSE
+savePlots = TRUE
 
 if (readFlag){
   # load(file = "/Users/micha/Desktop/Dropbox/priors_data/out_25.Rdat")
   load(file = "/Users/micha/Desktop/Dropbox/priors_data/out.Rdat")
 } else {
-  source('main.R')
+  # source('main.R')
 }
 
 # prepare samples
 csamples = tbl_df(melt(out$sims.list))
 colnames(csamples) = c("value", "step", "subject", "item", "bin", "variable")
 csamples$item = levels(factor(bin_dat$tag))[csamples$item]
+
+#######################
+# posteriors
+#######################
+
+# item pop
+meansIP = tbl_df(melt(out$sims.list$item.pop)) %>%
+  rename(step = Var1, item = Var2, bin = Var3) %>%
+  mutate(item = levels(factor(bin_dat$tag))[item] ) %>%
+  group_by(item, bin) %>%
+  summarise(
+    mean = mean(value),
+    max = HDIofMCMC(value)[2],
+    min = HDIofMCMC(value)[1]
+  )
+meansIP$y_emp = y.slider_means$mymean
+pop_priors = ggplot(meansIP, aes(x = bin, y = mean)) + geom_line() + geom_point() + facet_wrap(~ item, scale = "free") +
+  geom_ribbon(aes(ymin=min, ymax=max), fill="gray", alpha="0.5") +
+  geom_line(aes(x = bin, y = y_emp) , color = "red")  + geom_point(aes(x = bin, y = y_emp ), color = "red")
+show(pop_priors)
+
+# subjective priors
+# meansIPSubj = tbl_df(melt(out$sims.list$subj)) %>%
+#   rename(step = Var1, subject = Var2, item = Var3, bin = Var4) %>%
+#   mutate(item = levels(factor(bin_dat$tag))[item] ) %>%
+#   filter(subject <= 8) %>%
+#   group_by(subject, item, bin) %>%
+#   summarise(
+#     mean = mean(value),
+#     max = HDIofMCMC(value)[2],
+#     min = HDIofMCMC(value)[1]
+#   )
+# meansIPSubj$y_emp = y.slider_means$mymean
[…]

### 2016-01-26 prag_oldDev/process_data.R

@@ -1,7 +1,16 @@
+library(bootstrap)
+
 dat <- read.csv('data/exp1.csv')
 bin_dat <- read.csv('data/bin_dat.csv')
 choice_dat <- read.csv('data/choice_dat.csv')
 number_dat <- read.csv('data/number_dat.csv')
+
+## for bootstrapping 95% confidence intervals
+theta <- function(x,xdata,na.rm=T) {mean(xdata[x],na.rm=na.rm)}
+ci.low <- function(x,na.rm=T) {
+  mean(x,na.rm=na.rm) - quantile(bootstrap(1:length(x),1000,theta,x,na.rm=na.rm)$thetastar,.025,na.rm=na.rm)}
+ci.high <- function(x,na.rm=T) {
+  quantile(bootstrap(1:length(x),1000,theta,x,na.rm=na.rm)$thetastar,.975,na.rm=na.rm) - mean(x,na.rm=na.rm)}
 
 ## slider rating data
 y.slider = array(0, dim = c(20,8,15))
@@ -12,7 +21,9 @@ for (i in 1:nrow(bin_dat)){
 }
 y.slider = logit(add.margin(y.slider))
 y.slider_means = bin_dat %>% group_by(bin_num, tag) %>% 
-  summarise(mymean = mean(nresponse)) %>% rename(bin = bin_num, item = tag)
+  summarise(mymean = mean(nresponse),
+            cilow  = mean(nresponse) - ci.low(nresponse),
+            cihigh = mean(nresponse) + ci.high(nresponse)) %>% rename(bin = bin_num, item = tag)
 y.slider_means = y.slider_means[order(y.slider_means$item,y.slider_means$bin),]
 
 # number choice data

### 2016-01-27 prag_oldDev/plotting.R
polished plotting

@@ -39,28 +39,35 @@ meansIP = tbl_df(melt(out$sims.list$item.pop)) %>%
 meansIP$y_emp = y.slider_means$mymean
 pop_priors = ggplot(meansIP, aes(x = bin, y = mean)) + geom_line() + geom_point() + facet_wrap(~ item, scale = "free") +
   geom_ribbon(aes(ymin=min, ymax=max), fill="gray", alpha="0.5") +
-  geom_line(aes(x = bin, y = y_emp) , color = "red")  + geom_point(aes(x = bin, y = y_emp ), color = "red")
+  geom_line(aes(x = bin, y = y_emp) , color = "firebrick")  + geom_point(aes(x = bin, y = y_emp ), color = "firebrick")
 show(pop_priors)
 
 # subjective priors
-# meansIPSubj = tbl_df(melt(out$sims.list$subj)) %>%
-#   rename(step = Var1, subject = Var2, item = Var3, bin = Var4) %>%
-#   mutate(item = levels(factor(bin_dat$tag))[item] ) %>%
-#   filter(subject <= 8) %>%
-#   group_by(subject, item, bin) %>%
-#   summarise(
-#     mean = mean(value),
-#     max = HDIofMCMC(value)[2],
-#     min = HDIofMCMC(value)[1]
-#   )
-# meansIPSubj$y_emp = y.slider_means$mymean
-# meansIPSubj$subject = factor(meansIPSubj$subject)
-# pop_priorsSubj = ggplot(meansIPSubj, aes(x = bin, y = mean, group = subject)) + geom_line(color="lightgray") + geom_point(color="lightgray", size = 1.5) + facet_wrap(~ item, scale = "free") +
-#   geom_line(aes(x = bin, y = y_emp) , color = "black")  + geom_point(aes(x = bin, y = y_emp ), color = "black")
-# show(pop_priorsSubj)
+meansIPSubj = tbl_df(melt(out$sims.list$subj)) %>%
+  rename(step = Var1, subject = Var2, item = Var3, bin = Var4) %>%
+  mutate(item = levels(factor(bin_dat$tag))[item] ) %>%
+  filter(subject <= 20) %>%
+  group_by(subject, item, bin) %>%
+  summarise(
+    mean = mean(value),
+    max = HDIofMCMC(value)[2],
+    min = HDIofMCMC(value)[1]
+  )
+meansIPSubj$y_emp = y.slider_means$mymean
[…]

### 2016-01-27 prag_oldDev/plotting.R
CogSci paper start

@@ -40,7 +40,11 @@ meansIP$y_emp = y.slider_means$mymean
 pop_priors = ggplot(meansIP, aes(x = bin, y = mean)) + geom_line() + geom_point() + facet_wrap(~ item, scale = "free") +
   geom_ribbon(aes(ymin=min, ymax=max), fill="gray", alpha="0.5") +
   geom_line(aes(x = bin, y = y_emp) , color = "firebrick")  + geom_point(aes(x = bin, y = y_emp ), color = "firebrick")
-show(pop_priors)
+
+pop_priorsPaper = ggplot(meansIP, aes(x = bin, y = mean)) + geom_line() + geom_point() + facet_wrap(~ item, scales = "free_y", ncol=2) +
+  geom_ribbon(aes(ymin=min, ymax=max), fill="gray", alpha="0.5") +
+  geom_line(aes(x = bin, y = y_emp) , color = "firebrick")  + geom_point(aes(x = bin, y = y_emp ), color = "firebrick")
+
 
 # subjective priors
 meansIPSubj = tbl_df(melt(out$sims.list$subj)) %>%
@@ -106,10 +110,13 @@ plotSliderPPC = ggplot(slider_aggr, aes(x = bin, y = mean)) + geom_line() + geom
   geom_ribbon(aes(ymin=min, ymax=max), fill="gray", alpha="0.5") +
   # geom_errorbar(aes(ymin = min, ymax = max), width = .5, position = position_dodge(.1), color = 'gray') +
   geom_line(aes(x = bin, y = y_means) , color = "firebrick") + geom_point( aes(x = bin, y = y_means) , color = "firebrick")
-show(plotSliderPPC)
+plotSliderPPCPaper = ggplot(slider_aggr, aes(x = bin, y = mean)) + geom_line() + geom_point() + facet_wrap(~ item, scale = "free_y", ncol=2) + 
+  geom_ribbon(aes(ymin=min, ymax=max), fill="gray", alpha="0.5") +
+  # geom_errorbar(aes(ymin = min, ymax = max), width = .5, position = position_dodge(.1), color = 'gray') +
+  geom_line(aes(x = bin, y = y_means) , color = "firebrick") + geom_point( aes(x = bin, y = y_means) , color = "firebrick")
 plotSliderData = ggplot(slider_aggr, aes(x = bin, y = y_means)) + geom_line() + geom_point() + facet_wrap(~ item, scale = "free") + 
[…]

### 2016-01-27 prag_oldDev/postprocessing.R

@@ -1,0 +1,55 @@
+library('coda')
+library('ggmcmc')
+library('jagsUI') # for parallel computing
+source('helpers/helpers.R')
+# source('~/Desktop/data/svn/ProComPrag/dev_tmp/typicality_quantifiers/model/helpers.r')
+source('process_data.R')
+
+theme_set(theme_bw() + theme(plot.background=element_blank()) )
+
+readFlag = TRUE
+
+if (readFlag){
+  # load(file = "/Users/micha/Desktop/Dropbox/priors_data/out_25.Rdat")
+  load(file = "/Users/micha/Desktop/Dropbox/priors_data/out.Rdat")
+} else {
+  # source('main.R')
+}
+
+entropy = function(p){
+  - sum(p*log(p))
+}
+
+y.rep = logistic(out$sims.list$y.sliderPPC)
+y.emp = logistic(y.slider)
+
+posteriorMeans = array(0, dim = c(dim(y.rep)[1], dim(y.rep)[2], dim(y.rep)[3]))
+posteriorEntrs = array(0, dim = c(dim(y.rep)[1], dim(y.rep)[2], dim(y.rep)[3]))
+pValueMean = matrix(0, nrow = 20, ncol = 8)
+pValueEntr = matrix(0, nrow = 20, ncol = 8)
+
+for (i in 1:(dim(y.rep)[2])) {
+  for (j in 1:(dim(y.rep)[3])) {
+    for (s in 1:(dim(y.rep)[1])) {
+      posteriorMeans[s,i,j] = sum(1:15*y.rep[s,i,j,])
+      posteriorEntrs[s,i,j] = entropy(y.rep[s,i,j,])
+    }
+    empMean = sum(1:15*y.emp[i,j,])
+    empEntr = entropy(y.emp[i,j,])
+    biggerMean  = sum(ifelse(posteriorMeans[,i,j] > empMean, 1, 0)) / dim(y.rep)[1]
+    smallerMean = sum(ifelse(posteriorMeans[,i,j] < empMean, 1, 0)) / dim(y.rep)[1]
+    pValueMean[i,j] = min(biggerMean,smallerMean) * 2
+    biggerEntr  = sum(ifelse(posteriorEntrs[,i,j] > empEntr, 1, 0)) / dim(y.rep)[1]
+    smallerEntr = sum(ifelse(posteriorEntrs[,i,j] < empEntr, 1, 0)) / dim(y.rep)[1]
[…]

### 2016-01-28 prag_oldDev/main.R
CogSci paper motivation, model, notes

@@ -25,15 +25,15 @@ params <- c('w',
             )
 
 model = "models/model.jags.R"
-burnin = 100000
-iter = 25000
+burnin = 10000
+iter = 10000
 out = jags(data = data_aggr,
             inits = NULL,
             parameters.to.save = params,
             codaOnly = c("y.choicePPC", "y.numberPPC", "y.sliderPPC", "subj"),
             model.file = model,
             n.chains = 2,
-            n.adapt = 50000,
+            n.adapt = 5000,
             n.iter = iter + burnin,
             n.burnin = burnin,
             n.thin = 2, 
@@ -41,7 +41,7 @@ out = jags(data = data_aggr,
             verbose = TRUE,
             parallel = TRUE)
 
-if (saveFlag) { save(out, file = "~/Desktop/Dropbox/priors_data/out.Rdat") }
+if (saveFlag) { save(out, file = "~/Desktop/Dropbox/priors_data/outTMP.Rdat") }
 
 stop()
 

### 2016-01-28 prag_oldDev/plotting.R

@@ -12,6 +12,7 @@ savePlots = TRUE
 
 if (readFlag){
   # load(file = "/Users/micha/Desktop/Dropbox/priors_data/out_25.Rdat")
+  # load(file = "/Users/micha/Desktop/Dropbox/priors_data/outTMP.Rdat")
   load(file = "/Users/micha/Desktop/Dropbox/priors_data/out.Rdat")
 } else {
   # source('main.R')
@@ -39,11 +40,13 @@ meansIP = tbl_df(melt(out$sims.list$item.pop)) %>%
 meansIP$y_emp = y.slider_means$mymean
 pop_priors = ggplot(meansIP, aes(x = bin, y = mean)) + geom_line() + geom_point() + facet_wrap(~ item, scale = "free") +
   geom_ribbon(aes(ymin=min, ymax=max), fill="gray", alpha="0.5") +
-  geom_line(aes(x = bin, y = y_emp) , color = "firebrick")  + geom_point(aes(x = bin, y = y_emp ), color = "firebrick")
+  geom_line(aes(x = bin, y = y_emp) , color = "firebrick")  + geom_point(aes(x = bin, y = y_emp ), color = "firebrick") +
+  ylab(" mean posterior Q_{ijk} / average slider rating")
 
 pop_priorsPaper = ggplot(meansIP, aes(x = bin, y = mean)) + geom_line() + geom_point() + facet_wrap(~ item, scales = "free_y", ncol=2) +
   geom_ribbon(aes(ymin=min, ymax=max), fill="gray", alpha="0.5") +
-  geom_line(aes(x = bin, y = y_emp) , color = "firebrick")  + geom_point(aes(x = bin, y = y_emp ), color = "firebrick")
+  geom_line(aes(x = bin, y = y_emp) , color = "firebrick")  + geom_point(aes(x = bin, y = y_emp ), color = "firebrick") +
+  ylab(" mean posterior Q_{ijk} / average slider rating")
 
 
 # subjective priors
@@ -109,11 +112,13 @@ slider_aggr$cihigh = y.slider_means$cihigh
 plotSliderPPC = ggplot(slider_aggr, aes(x = bin, y = mean)) + geom_line() + geom_point() + facet_wrap(~ item, scale = "free") + 
   geom_ribbon(aes(ymin=min, ymax=max), fill="gray", alpha="0.5") +
   # geom_errorbar(aes(ymin = min, ymax = max), width = .5, position = position_dodge(.1), color = 'gray') +
[…]

### 2016-01-28 prag_oldDev/process_data.R

@@ -40,6 +40,8 @@ choice_dat$condition = as.numeric(choice_dat$condition)
 y.choice = array(0, dim = c(20,8,5))
 higher   = array(0, dim = c(20,8,5))
 lower    = array(0, dim = c(20,8,5))
+chosenBin      = array(0, dim = c(20,8,5))
+unchosenBin    = array(0, dim = c(20,8,5))
 for (i in 1:nrow(choice_dat)){
   y.choice[choice_dat$workerid[i] + 1, 
            as.numeric(choice_dat$tag)[i],
@@ -50,4 +52,10 @@ for (i in 1:nrow(choice_dat)){
   lower[choice_dat$workerid[i] + 1, 
         as.numeric(choice_dat$tag)[i],
         choice_dat$condition[i]] = min(choice_dat$chosen_bin[i], choice_dat$unchosen_bin[i])
+  chosenBin[choice_dat$workerid[i] + 1, 
+        as.numeric(choice_dat$tag)[i],
+        choice_dat$condition[i]] = choice_dat$chosen_bin[i]
+  unchosenBin[choice_dat$workerid[i] + 1, 
+             as.numeric(choice_dat$tag)[i],
+             choice_dat$condition[i]] = choice_dat$unchosen_bin[i]
 }

### 2016-01-29 text/01_CogSci_abstract/paper.tex
CogSci paper result plots minor edits

% 
% Annual Cognitive Science Conference
% Sample LaTeX Paper -- Proceedings Format
%

% Original : Ashwin Ram (ashwin@cc.gatech.edu)       04/01/1994
% Modified : Johanna Moore (jmoore@cs.pitt.edu)      03/17/1995
% Modified : David Noelle (noelle@ucsd.edu)          03/15/1996
% Modified : Pat Langley (langley@cs.stanford.edu)   01/26/1997
% Latex2e corrections by Ramin Charles Nakisa        01/28/1997 
% Modified : Tina Eliassi-Rad (eliassi@cs.wisc.edu)  01/31/1998
% Modified : Trisha Yannuzzi (trisha@ircs.upenn.edu) 12/28/1999 (in process)
% Modified : Mary Ellen Foster (M.E.Foster@ed.ac.uk) 12/11/2000
% Modified : Ken Forbus                              01/23/2004
% Modified : Eli M. Silk (esilk@pitt.edu)            05/24/2005
% Modified : Niels Taatgen (taatgen@cmu.edu)         10/24/2006
% Modified : David Noelle (dnoelle@ucmerced.edu)     11/19/2014

%% Change "letterpaper" in the following line to "a4paper" if you must.

\documentclass[10pt,letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{apacite}
\usepackage{subcaption}
\usepackage{todonotes}
\usepackage{amsmath}

%% commands
\newcommand{\set}[1]{\left\{#1\right\}}
\DeclareMathOperator{\expo}{exp}
\newcommand{\citet}[1]{\citeA{#1}}
\newcommand{\citep}[1]{\cite{#1}}

\title{What does the crowd believe? A hierarchical approach to estimating subjective beliefs
  from empirical data}

\author{{\large \bf Erin Bennett, Judith Degen, Michael Henry Tessler, Noah D. Goodman}\\
     \{ebennett,jdegen,mhtessler,ngoodman\}@stanford.edu \\
     Department of Psychology, 450 Serra Mall \\
  Stanford, CA 94305 USA \AND {\large \bf Michael Franke} \\
  michael.franke@uni-tuebingen.de \\
  Department of Linguistics, Wilhelmstra\ss e 19 \\
 72074 T\"{u}bingen, Germany
  \AND}

\begin{document}

\maketitle

\todo[inline]{Each author's name should appear
on a separate line, 11~point bold, and centered, with the author's
email address in parentheses. Under each author's name list the
author's affiliation and postal address in ordinary 10~point type.}

\todo[inline]{who is an author? which order?}

\begin{abstract}
  Subjects' beliefs about everyday events are of theoretical interest on their own and also an
  important ingredient in especially Bayesian models of such diverse phenomena as logical
  reasoning, future predictions or language use and interpretation. Here, we scrutinize one
  recently popular method for measuring subjective beliefs experimentally. We present a
  hierarchical Bayesian model for inferring likely ``population-level beliefs'' as the central
  tendency of subjects' individual-level beliefs. Individual-level beliefs define likelihood
  functions for three types of task measures. Our results suggest that the previous practice of
  using averaged normalized slider ratings for binned quantities is a practical and fairly good
  approximator of latently inferred population-level beliefs.

\textbf{Keywords:} subjective beliefs, hierarchical modeling, Bayesian data analysis, Bayesian
cognitive models 
\end{abstract}

\section{Motivation}

We cannot look into a person's head. But in making sense of observed behavior we readily
ascribe beliefs and desires to fellow agents. This happens intuitively, in folk psychology, but
also in science. Ascriptions of latent mental states play an important role as well in many
explanations of especially higher-order cognition, like decision making, planning, reasoning,
or language use. Naturally, then, it becomes important how to validate any putative ascription
of mental states for explanatory purposes.

A family of models where this is particularly pressing are Bayesian models of cognition which
seek to explain task behavior in a variety of domains as partially informed by what subjects
believe about mundane events. Take interpretation of language. Empirical data on whether a
statement like ``That watch cost $n$ dollars'' is understood to convey speaker affect, rather
than literal meaning, can be explained well by a Bayesian model of utterance interpretation
\citep{KaoWu2014:Nonliteral-Unde} in which a crucial role is played, as is quite intuitive, by
an empirical measure of subjects' expectations about the likely or normal price of a
watch. Other examples of domains in which empirically successful models have included some
measure of subjects' belief include making future predictions
\citep{GriffithsTenenbaum2006:Optimal-Predict}, the strength of pragmatic enrichments
\citep{DegenTessler2015:Wonky-worlds:-L}, the interpretation of vague quantifiers
\citep{SchollerFranke2015:Semantic-values}, \textcolor{red}{ \dots }

\todo[inline]{fill me}

Many methods of assessing subjective beliefs exist. One approach is to take actual frequencies
as an approximation \citep{GriffithsTenenbaum2006:Optimal-Predict}. This, however, does not
work for beliefs about one-shot or imaginary events. Moreover, even where available, real-world
frequencies may deviate systematically from subjects' beliefs, and this can be problematic for
the predictions of a model that relies on subjective beliefs
\citep{MarcusDavis2013:How-Robust-Are-}.

Another approach is to try to experimentally measure relevant beliefs. Many techniques for this
exist, especially in the economics literature
\citep{MorganHenrion1990:Uncertainty:-A-,Manski2004:Measuring-Expec,SchlagTremewan2014:A-penny-for-you,AndersenFountain2014:Estimating-Subj}. The
perhaps most prominent method uses so-called \emph{scoring rules}
\citep{Savage1971:Elicitation-of-,SchlagTremewan2014:A-penny-for-you}. Roughly speaking,
scoring rules are particular schemes of rewarding participants' answers in such a way as to
make sure that responses are what we want them to be: the actual subjective probability of an
event, the mean of a distribution, a credible interval etc. Scoring rules are highly faithful
but require participants to become sufficiently familiar with the (usually probabilistic)
payoff scheme. Scoring rules also must make assumptions about how much participants value
certain probabilistic gambles and may therefore require elaborate means of data analysis
\citep{AndersenFountain2014:Estimating-Subj}. Similar practical concerns apply to other methods
as well, such as the use of \emph{iterated learning} for prior estimation
\citep{LewandowskyGriffiths2009:The-Wisdom-of-I}.

For common applications, many of these procedures may be excessively complex or too demanding
on both researcher and participant. It would be beneficial if there was an easily applicable
technique of eliciting subjective beliefs that is good enough for testing the general
predictions of models that rely on these measures for their predictions. A further complication
is that it may be impractical to obtain, from the same participant, data on subjective beliefs
and data on the task of interest, especially if exposure to one of these measures is likely to
affect the other. This suggests that, for practical reasons, there is a demand of
reliable-enough measures of what we will sloppily call ``population-level beliefs'': estimates
of the central tendency or average of subjective beliefs in a given population.

A very simple technique that has been used with apparent success is to have subjects adjust
sliders in order to express their (relative) levels of support for values of uncertain
contiguous quantities
\citep{KaoWu2014:Nonliteral-Unde,SchollerFranke2015:Semantic-values}. \textcolor{red}{ MORE REFERENCES } Subjects are confronted with a short and simple description of a situation or an
event, such as ``John bought a watch,'' and asked for their estimates of the watch's price. To
give their estimate, they are asked to adjust sliders, whose ends are labelled as ``extremely
likely'' and ``extremely unlikely,'' for a number of bins that partition the range of plausible
values (established by a pre-test). E.g., for the price of a watch bins could be
\textcolor{red}{ FILL ME }

\begin{itemize}
\item explain slider rating task
\item our goal here is to validate whether ``population-level beliefs'' are approximated well
  enough by normalized average slider ratings
\item use hierarchical Bayesian modeling
  \begin{itemize}
  \item calculate likelihoods from subjective beliefs
  \item each subjective belief is drawn from a population-level central tendency, which is a
    ``belief'' as well
  \end{itemize}
\end{itemize}

\section{Experiments}

\begin{itemize}
\item why within-participant?
\item what was the exact randomization / procedure?
\item items are given in Table~\ref{tab:Items}
\item include bins as well?
\item what were the lighting round bin comparisons?
  \begin{itemize}
  \item 1 vs 2
  \item 2 vs 6
  \item 6 vs 11
  \item 11 vs 14
  \item 14 vs 15
  \end{itemize}
\item results are in Figure~\ref{fig:Results}
\end{itemize}

\begin{table}
  \centering
  \begin{tabular}{p{8cm}}
    1.  X has just fetched himself a cup of \textbf{coffee} from the office vending machine.
         What do you think the temperature of his coffee is?  \\
    2.  X \textbf{commuted} to work yesterday. 
        How many minutes do you think she spent commuting yesterday? \\
3.  X told a \textbf{joke} to N kids. 
    How many of the kids do you think laughed?\\

4.  X bought a \textbf{laptop}.    
    How much do you think it cost?\\

5.  X threw N \textbf{marbles} into a pool.
    How many of the marbles do you think sank?\\

6.  X just went to the \textbf{movies} to see a blockbuster.
    How many minutes long do you think the movie was?\\

7.  X watched \textbf{TV} last week.
    How many hours do you think he spent watching TV last week? \\

8.  X bought a \textbf{watch}.
    How much do you think it cost?
  \end{tabular}
  \caption{Experimental items}
  \label{tab:Items}
\end{table}

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.5\textwidth}
    \includegraphics[width = \textwidth]{plots/data_sliderNumber.pdf}
    \caption{slider ratings \& number choices}
    \label{fig:slider}
  \end{subfigure}

  \begin{subfigure}[b]{0.5\textwidth}
    \includegraphics[width = \textwidth]{plots/data_choice.pdf}
    \caption{bin comparison}
    \label{fig:lighting}
  \end{subfigure}

  \caption{Results from experiments. Error bars show bootstrapped 95\% confidence intervals.}
  \label{fig:Results}
\end{figure}

\section{Model}

The data we would like to explain are: (i) the normalized slider ratings $s_{ijk} \in [0;1]$ of
subject $i \in \set{1, \dots, 20}$ for item $j \in \set{1, \dots, 8}$ and bin
$k \in \set{1, \dots, 15}$; (ii) the bins $n_{ij} \in \set{1, \dots, 15}$ in which subject
$i$'s number choice for item $i$ was; and (iii) the binary choices $c_{ijl} \in \set{0,1}$ of
whether subject $i$ selected the higher bin for item $j$ in the bin comparison condition
$l$. There two simplifications in need of commenting. In (i), we focus on slider ratings after
normalizing for each subject, because we assume that slider adjustments reflect relative, not
absolute estimates of subjective beliefs. In (ii), we focus on bin choices, not actual number
choices, in order to avoid, as much as possible, considerations of salience of particular
numbers, and also because otherwise data from items with smaller domains of plausible number
would get more weight than data from items which allow for a wider set of number choices.

All three pieces of data are to be explained as functions of subjective beliefs $P_{ij}$, where
$P_{ij}$ is a probability vector of length 15 with $P_{ijk}$ being subject $i$'s belief about
the relative likelihood of bin $k$ for item $j$. Each $P_{ij}$ defines a likelihood for our
data, via appropriate link functions (to be spelled out presently). Variance in subjective
beliefs is harnessed by a population-level hyper-prior with central tendency $Q_{j}$, i.e., a
stochastic vector of length 15, which we call a ``population-level belief'' about item $j$. The
structure of this model is pictured in Figure~\ref{fig:ModelGraph}.

\begin{figure}[]
  \centering
  \includegraphics[width = 0.4\textwidth]{modelGraph/modelGraphNoMath.pdf}
  \caption{The data-generating model as a probabilistic graphical model, following conventions
    of \citet{LeeWagenmakers2013:Bayesian-Cognit}. Shaded nodes are observed, white nodes are
    latent variables. Square nodes represent categorical, round nodes continuous
    variables. Boxes indicate scope of indices.}
  \label{fig:ModelGraph}
\end{figure}

To fill the structure in Figure~\ref{fig:ModelGraph} with life, we need to spell out three
parameterized link functions, one for each task type, and the relation between population-level
belief $Q_j$ to individual beliefs $P_{ij}$. Let's start with the latter. For fixed $Q_j$, the
idea is that $P_{ij}$ are noise-perturbed variants scattered around $Q_j$, with some parameter
$w$ to determine how much perturbation we should expect. To realize this, the model assumes
that $P_{ij}$ are distributed according to a Dirichlet distribution with weights given by
$w Q_{j}$:
\begin{align*}
  Q_j & \sim \text{Dirichlet}(1, \dots, 1) & w & \sim \text{Gamma}(2,0.1) \\
  P_{ij} & \sim \text{Dirichlet}( w Q_j)
\end{align*}
The higher $w$, the more likely it is that $P_{ij}$ is ``close'' to $Q_j$.

The link function for the slider rating data uses a logit transformation to project observed
slider ratings $s_{ijk}$ and latent probabilities $P_{ijk}$, which are bound to lie between 0
and 1, to the reals. The likelihood of logit-transformed observation $s_{ijk}$ is given by a
Gaussian with standard deviation $\sigma$ around the logit-transformed predictor $P_{ijk}$. On
top of that, there is a parameter $\kappa$, the steepness of the logit transform of $P_{ijk}$,
that allows response likelihoods to capture end-point affinity for $\kappa >1$ (values of
$P_{ijk}$ close to 0 or 1 are likely mapped to 0 or 1) or end-point aversion for $\kappa <0$
(values of $P_{ijk}$ are likely to be realized as more median), with a prior that expects
$\kappa=1$.
\begin{align*}
  \text{logit}(s_{ijk}) &\sim
        \text{Norm}(\text{logit}(P_{ijk}, \kappa), \sigma) \\
        \sigma &  \sim \text{Gamma}(0.0001,0.0001) & \kappa &\sim \text{Gamma}(5,5)
\end{align*}

The link function for number choice data treats each bin $n_{ij}$ as a draw from a categorical
distribution where the probability of bin $k$ is proportional to $\expo(a P_{ij})$, i.e., a
soft-max choice from $P_{ij}$. The higher parameter $a$, the more likely $n_{ij}$ is the mode
of $P_{ij}$. For $a \rightarrow 0$, all bins become equiprobable.
\begin{align*}
  n_{ij} & \sim
        \text{Categorical}(\expo(a P_{ij})) &
 a & \sim \text{Gamma}(2,1)
\end{align*}

Finally, consider the link function for bin comparisons. We are after the likelihood with which
subject $i$ selects the higher bin for item $j$ in bin comparison condition $l$. Suppose $l$ is
about comparing the lower bin $b_l$ to the higher $b_h$. The perhaps most natural approach
would be to link the likelihood of choosing $b_h$ over $b_l$ to the difference between
$P_{ijb_h}$ and $P_{ijb_l}$. This, however, does not appear to be what subjects were doing. For
example, take the ``marbles'' case where we see almost unanimous choices of bin 11 over bin 6,
despite seeing any pronounced difference in slider ratings for these bins. Indeed, a model that
implements this idea failed to capture the relevant regularities in the data blatantly.

Another plausible link function that accommodates for this is to assume that what matters in
bin comparisons is the distance to the mode of $P_{ij}$: soft-max prefer the bin that is closer
to the mode of $P_{ij}$; select random if both bins are equally far from the
prototype.\footnote{Notice that
  $(1 + \exp(2b(1-p^{\text{high}}_{ijl})) )^{-1} = \frac{\exp(b p^{\text{high}})}{\exp(b
    p^{\text{high}}) + \exp(b p^{\text{low}})}$
  with $p^\text{low}_{ijl} = 2 - p^\text{high}_{ijl}$.}
\begin{align*}
  c_{ijl} & \sim \text{Bern}( (1 + \exp(2b(1-p^{\text{high}}_{ijl})) )^{-1} ) \ \ \ 
  b  \sim \text{Gamma}(2,1) \\
  p^\text{high}_{ijl} & = \begin{cases}
    2 & \text{if $mode(P_{ij})$ is closer to higher} \\
    & \text{ bin of $l$ than to lower bin } \\ 1 & \text{if equal
      distance} \\ 0 & \text{otherwise}  
  \end{cases}
\end{align*}

\section{Inference}

The model was implemented in JAGS \cite{Plummer2003:JAGS:-A-Program}. 50,000 samples were
obtained from two chains with a thinning rate of 2 after a burn-in of 100,000 that ensured
convergence according to $\hat{R}$ \cite{GelmanRubin1992:Inference-from-}.

Of most theoretical interest are posterior estimates of $Q_j$. Figure~\ref{fig:PosteriorQj}
shows means of the posterior $Q_j$, with their 95\% HDIs, alongside the averaged normalized
slider ratings. The latter provide a very good approximation of the latent, inferred
population-level beliefs. Inspection of posteriors of individual $P_{ij}$ shows that there is
ample variation between subjects. Nonetheless, the way the model suggest we should think about
harnessing the individual $P_{ij}$s under a population-level central tendency is closely
approximated by simply averaging over slider ratings. Although the match is clearly not
perfect, it is good enough to maintain that the latter are a reasonable (enough) and practical
way of assessing what the crowd believes despite individual differences.

\begin{figure}
  \centering
  \includegraphics[width = 0.5\textwidth]{plots/pop_priors.pdf}
  \caption{Means of posteriors over $Q_j$ in black with gray area indicating 95\% HDIs. Red
    lines give the average normalized slider ratings for comparison.}
  \label{fig:PosteriorQj}
\end{figure}

\section{Model criticism}

Inferences based on the model are only as reliable as the model itself is plausible. Model
criticism is therefore important. Figure~\ref{fig:PPCs} shows posterior predictive checks at
the population-aggregate level for all of our three task types. For the slider rating task,
posterior predictions are spot-on. Some of the number estimation data is surprising even for
the model trained on this very data. This could have various reasons: (i) the number data does
not have a huge influence on the posterior likelihood, (ii) number choices may be influenced by
saliency and/or roundness of numbers after all. Finally, there is one condition in the bin
comparison task that the model definitely got wrong. This is the choice of what is more likely:
that one or that none of 14 marbles thrown into a pool would sink. The model would predict that
almost everybody should answer that it is more likely that one marble sank. But that is not
what we observe. It may be that subjects revise beliefs about ``normality'' of the marbles,
while holding on to an assumption that all marbles behave the same
\cite{DegenTessler2015:Wonky-worlds:-L}.

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.5\textwidth}
    \includegraphics[width = \textwidth]{plots/ppc_slider.pdf}
    \caption{slider rating}
    \label{fig:sliderPPC}
  \end{subfigure}

  \begin{subfigure}[b]{0.5\textwidth}
    \includegraphics[width = \textwidth]{plots/ppc_number.pdf}
    \caption{number estimation}
    \label{fig:numberPPC}
  \end{subfigure}

  \begin{subfigure}[b]{0.5\textwidth}
    \includegraphics[width = \textwidth]{plots/ppc_choice.pdf}
    \caption{bin comparison}
    \label{fig:lightingPPC}
  \end{subfigure}

  \caption{Posterior predictive checks for aggregate data. Red lines give empirical
    observations. Black lines are means of posterior predictive samples, gray areas are
    95\% HDIs.}
  \label{fig:PPCs}
\end{figure}

Posterior predictive checks indicate that the trained model captures patterns of answers at the
aggregate population level well. To have a more fine-grained measure of model fit, we also
looked at posterior predictive $p$-values at the level of subjects and items. Unfortunately,
only the slider rating task, provides enough data for such a fine-grained analysis. For the
slider rating task, fixing a subject and an item, observations and replicates are probability
vectors of length 15. In a first analysis, we used the mean of these probability vectors as a
test statistic. The minimum posterior predictive $p$-value over all 20 (subjects) times 8
(items) cases was 0.13, suggesting that the means of observed $s_{ij}$ are non-surprising to
the trained model. In a second analysis, we used entropy as a test statistic. Two cases gave
posterior predictive $p$-values lower than 0.05. These were from the two participants who gave a
very extreme slider rating for the ``marbles'' item, basically assigning all ``mass'' to the
last bin. What this suggests is that the model can cope reasonably well also with
individual-level data, but has problems accounting for ``extreme'' choices, given that the
population-level hyper-prior on $P_{ij}$ will lead to shrinkage.

\section{Conclusions}

\section{Acknowledgments}

MF's work was supported by the Institutional Strategy of the University of T̈\"{u}bingen (Deutsche
Forschungsgemeinschaft, ZUK 63) and the Priority Program XPrag.de (DFG Schwerpunktprogramm
1727).

\bibliographystyle{apacite}

\setlength{\bibleftmargin}{.125in}
\setlength{\bibindent}{-\bibleftmargin}

\bibliography{MyRefGlobal}

\end{document}

### 2016-01-29 prag_oldDev/plotting.R

@@ -120,7 +120,12 @@ plotSliderPPCPaper = ggplot(slider_aggr, aes(x = bin, y = mean)) + geom_line() +
   geom_line(aes(x = bin, y = y_means) , color = "firebrick") + geom_point( aes(x = bin, y = y_means) , color = "firebrick") +
   ylab("slider rating")
 plotSliderData = ggplot(slider_aggr, aes(x = bin, y = y_means)) + geom_line() + geom_point() + facet_wrap(~ item, scale = "free") + 
-  geom_errorbar(aes(ymin = cilow, ymax = cihigh), width = .5, position = position_dodge(.1), color = 'gray') 
+  geom_errorbar(aes(ymin = cilow, ymax = cihigh), width = .5, position = position_dodge(.1), color = 'gray') +
+  ylab("mean slider rating")
+plotSliderDataPaper = ggplot(slider_aggr, aes(x = bin, y = y_means)) + geom_line() + geom_point() + facet_wrap(~ item, scale = "free_y", ncol = 2) + 
+  geom_errorbar(aes(ymin = cilow, ymax = cihigh), width = .5, position = position_dodge(.1), color = 'gray') +
+  ylab("mean slider rating")
+
 
 
 # numbers
@@ -150,8 +155,13 @@ plotNumbersPPCPaper = ggplot(number_aggr, aes(x = bin, y = mean)) + geom_point()
   geom_line( aes(x = bin, y = value) , color = "firebrick") + geom_point( aes(x = bin, y = value) , color = "firebrick") +
   ylab("frequency")
 plotNumbersData = ggplot(number_aggr, aes(x = bin, y = value)) + geom_bar(stat = "identity") +
-  facet_wrap(~ item, scale = "free")
-show(plotNumbersData)
+  facet_wrap(~ item, scale = "free") + ylab("frequency number choice")
+plotNumbersDataPaper = ggplot(number_aggr, aes(x = bin, y = value)) + geom_bar(stat = "identity") +
+  facet_wrap(~ item, scale = "free_y", ncol = 2) + ylab("frequency number choice")
+plotSliderNumberDataPaper = ggplot(slider_aggr, aes(x = bin, y = y_means)) + geom_line() + geom_point() + facet_wrap(~ item, scale = "free_y", ncol = 2) + 
+  geom_errorbar(aes(ymin = cilow, ymax = cihigh), width = .5, position = position_dodge(.1), color = 'gray') +
[…]

### 2016-01-29 text/01_CogSci_abstract/paper.tex
edits

@@
 \usepackage{amsmath}
+\usepackage{color}
+
+\definecolor{Green}{RGB}{10,200,100}
+\newcommand{\ndg}[1]{\textcolor{Green}{[ndg: #1]}}
 
@@
   Subjects' beliefs about everyday events are of theoretical interest on their own and also an
-  important ingredient in especially Bayesian models of such diverse phenomena as logical
-  reasoning, future predictions or language use and interpretation. Here, we scrutinize one
-  recently popular method for measuring subjective beliefs experimentally. We present a
-  hierarchical Bayesian model for inferring likely ``population-level beliefs'' as the central
-  tendency of subjects' individual-level beliefs. Individual-level beliefs define likelihood
-  functions for three types of task measures. Our results suggest that the previous practice of
+  important ingredient for model building---especially in Bayesian models of phenomena such as logical
+  reasoning, future predictions, and language use. Here, we explore several recently used 
+  methods for measuring subjective beliefs about unidimensional properties. \ndg{say what kind of subjective beliefs /distributions we look at more clearly.} 
+  We a hierarchical Bayesian data analysis model for inferring likely ``population-level beliefs'' as the central
+  tendency of subjects' individual-level beliefs. 
+  Three different dependent measures---\ndg{name them?}---are related to individual-level beliefs through likelihood ``linking functions.''
+%  functions for three types of task measures. 
+  Our results suggest that the practice of
   using averaged normalized slider ratings for binned quantities is a practical and fairly good
-  approximator of latently inferred population-level beliefs.
-
+  approximator of inferred population-level beliefs.
+  Other dependent measures yield less reliable results.
 \textbf{Keywords:} subjective beliefs, hierarchical modeling, Bayesian data analysis, Bayesian
@@
 ascribe beliefs and desires to fellow agents. This happens intuitively, in folk psychology, but
-also in science. Ascriptions of latent mental states play an important role as well in many
-explanations of especially higher-order cognition, like decision making, planning, reasoning,
-or language use. Naturally, then, it becomes important how to validate any putative ascription
-of mental states for explanatory purposes.
+also in science. Scientific ascription of latent mental states plays an important role in many
+explanations of higher-order cognition: decision making, planning, reasoning,
+and language use. Naturally, then, it becomes important to validate any method for eliciting mental states, 
+to be used for explanatory purposes.
 
@@
 seek to explain task behavior in a variety of domains as partially informed by what subjects
-believe about mundane events. Take interpretation of language. Empirical data on whether a
+believe about mundane events---their \emph{prior beliefs}. Take interpretation of language. Empirical data on whether a
 statement like ``That watch cost $n$ dollars'' is understood to convey speaker affect, rather
 than literal meaning, can be explained well by a Bayesian model of utterance interpretation
-\citep{KaoWu2014:Nonliteral-Unde} in which a crucial role is played, as is quite intuitive, by
+\citep{KaoWu2014:Nonliteral-Unde}. In this model a crucial role is played by
 an empirical measure of subjects' expectations about the likely or normal price of a
-watch. Other examples of domains in which empirically successful models have included some
+watch. 
+Other examples of domains in which empirically successful models have included some
 measure of subjects' belief include making future predictions
@@
 Many methods of assessing subjective beliefs exist. One approach is to take actual frequencies
-as an approximation \citep{GriffithsTenenbaum2006:Optimal-Predict}. This, however, does not
-work for beliefs about one-shot or imaginary events. Moreover, even where available, real-world
+as an approximation \citep{GriffithsTenenbaum2006:Optimal-Predict}. This will not
+work when sufficient real-world data is unavailable---for instance, for beliefs about imaginary events. 
+Even where available, real-world
 frequencies may deviate systematically from subjects' beliefs, and this can be problematic for
-the predictions of a model that relies on subjective beliefs
-\citep{MarcusDavis2013:How-Robust-Are-}.
+the predictions of a model that relies on subjective beliefs.
+%\citep{MarcusDavis2013:How-Robust-Are-}.
 
-Another approach is to try to experimentally measure relevant beliefs. Many techniques for this
+Another approach is to experimentally elicit relevant beliefs. Many techniques for this
 exist, especially in the economics literature
@@
 \textcolor{red}{ FILL ME }
+
+\ndg{should we also talk about the other dependent measures? i think part of what's interesting is comparing the different dependent measures to each other?}
+
+\ndg{say more clearly that we focus on unidimensional distributions and why.}
 
@@
 
-\bibliography{MyRefGlobal}
+\bibliography{references}
 

### 2016-01-30 text/01_CogSci_abstract/paper.tex

@@
 
+\ndg{wow, this is weird! you tried softmax ratio of bin probs?}
+
 \section{Inference}
@@
 way of assessing what the crowd believes despite individual differences.
+
+\ndg{it'd be nice to see some posteriors for some of the (more meaningful) parameters. eg w, kappa, ...}
+
+\ndg{what about the success of the other DMs? looks like give-n is not so hot. worth saying that if so.}
 
@@
 aggregate population level well. To have a more fine-grained measure of model fit, we also
-looked at posterior predictive $p$-values at the level of subjects and items. Unfortunately,
+looked at posterior predictive $p$-values at the level of subjects and items. \ndg{need to say what this means -- i'm not even sure that i know!} Unfortunately,
 only the slider rating task, provides enough data for such a fine-grained analysis. For the
@@
 population-level hyper-prior on $P_{ij}$ will lead to shrinkage.
+\ndg{can say that one possibility (future work) would be to switch to a fatter-tailed distribution relating Q to P, in order to deal with outliers.}
 

### 2016-01-31 text/01_CogSci_abstract/paper.tex
filled in experiment and minor edits to rest of paper

@@
 \usepackage{color}
+\usepackage{booktabs}
 
@@
 \newcommand{\ndg}[1]{\textcolor{Green}{[ndg: #1]}}  
+\definecolor{Blue}{RGB}{10,100,200}
+\newcommand{\jd}[1]{\textcolor{Blue}{[jd: #1]}}
 
@@
 
+\newcommand{\tableref}[1]{Table \ref{#1}}
+\newcommand{\figref}[1]{Fig.~\ref{#1}}
+\newcommand{\appref}[1]{Appendix \ref{#1}}
+\newcommand{\sectionref}[1]{Section \ref{#1}}
+
 \title{What does the crowd believe? A hierarchical approach to estimating subjective beliefs
@@
 
-\author{{\large \bf Erin Bennett, Judith Degen, Michael Henry Tessler, Noah D. Goodman}\\
+\author{{\large \bf Michael Franke} \\
+  michael.franke@uni-tuebingen.de \\
+  Department of Linguistics, Wilhelmstra\ss e 19 \\
+ 72074 T\"{u}bingen, Germany \AND 
+ {\large \bf Erin Bennett, Judith Degen, Michael Henry Tessler, Noah D. Goodman}\\
      \{ebennett,jdegen,mhtessler,ngoodman\}@stanford.edu \\
      Department of Psychology, 450 Serra Mall \\
-  Stanford, CA 94305 USA \AND {\large \bf Michael Franke} \\
-  michael.franke@uni-tuebingen.de \\
-  Department of Linguistics, Wilhelmstra\ss e 19 \\
- 72074 T\"{u}bingen, Germany
-  \AND}
+  Stanford, CA 94305 USA }
 
@@
 \begin{abstract}
-  Subjects' beliefs about everyday events are of theoretical interest on their own and also an
+  Experimental participants' beliefs about everyday events are of theoretical interest on their own and also an
   important ingredient for model building---especially in Bayesian models of phenomena such as logical
@@
   We a hierarchical Bayesian data analysis model for inferring likely ``population-level beliefs'' as the central
-  tendency of subjects' individual-level beliefs. 
+  tendency of participants' individual-level beliefs. 
   Three different dependent measures---\ndg{name them?}---are related to individual-level beliefs through likelihood ``linking functions.''
@@
 A family of models where this is particularly pressing are Bayesian models of cognition which
-seek to explain task behavior in a variety of domains as partially informed by what subjects
+seek to explain task behavior in a variety of domains as partially informed by what participants
 believe about mundane events---their \emph{prior beliefs}. Take interpretation of language. Empirical data on whether a
@@
 \citep{KaoWu2014:Nonliteral-Unde}. In this model a crucial role is played by
-an empirical measure of subjects' expectations about the likely or normal price of a
+an empirical measure of participants' expectations about the likely or normal price of a
 watch. 
 Other examples of domains in which empirically successful models have included some
-measure of subjects' belief include making future predictions
+measure of participants' belief include making predictions about everyday events 
 \citep{GriffithsTenenbaum2006:Optimal-Predict}, the strength of pragmatic enrichments
@@
 Even where available, real-world
-frequencies may deviate systematically from subjects' beliefs, and this can be problematic for
+frequencies may deviate systematically from participants' beliefs, and this can be problematic for
 the predictions of a model that relies on subjective beliefs.
@@
 scoring rules are particular schemes of rewarding participants' answers in such a way as to
-make sure that responses are what we want them to be: the actual subjective probability of an
+make sure that responses are what the experimenter wants them to be: the actual subjective probability of an
 event, the mean of a distribution, a credible interval etc. Scoring rules are highly faithful
@@
 and data on the task of interest, especially if exposure to one of these measures is likely to
-affect the other. This suggests that, for practical reasons, there is a demand of
-reliable-enough measures of what we will sloppily call ``population-level beliefs'': estimates
+affect the other. This suggests that, for practical reasons, there is a demand for
+reliable-enough measures of what we will  call ``population-level beliefs'': estimates
 of the central tendency or average of subjective beliefs in a given population.
 
-A very simple technique that has been used with apparent success is to have subjects adjust
+A very simple technique that has been used with apparent success is to have participants adjust
 sliders in order to express their (relative) levels of support for values of uncertain
 contiguous quantities
-\citep{KaoWu2014:Nonliteral-Unde,SchollerFranke2015:Semantic-values}. \textcolor{red}{ MORE REFERENCES } Subjects are confronted with a short and simple description of a situation or an
+\citep{KaoWu2014:Nonliteral-Unde,SchollerFranke2015:Semantic-values}. \textcolor{red}{ MORE REFERENCES } Participants are confronted with a short and simple description of a situation or an
 event, such as ``John bought a watch,'' and asked for their estimates of the watch's price. To
-give their estimate, they are asked to adjust sliders, whose ends are labelled as ``extremely
+give their estimate, they are asked to adjust sliders whose endpoints are labelled as ``extremely
 likely'' and ``extremely unlikely,'' for a number of bins that partition the range of plausible
-values (established by a pre-test). E.g., for the price of a watch bins could be
+values (established by a pre-test). E.g., for the price of a watch bins could be \jd{erin: fill in}
 \textcolor{red}{ FILL ME }
@@
 
-\section{Experiments}
+\section{Experiment}
+
+\subsection{Methods}
+
+\paragraph{Participants} We recruited 20 self-reported native speakers of English over Mechanical Turk.
+
+\paragraph{Procedure and materials}
+
+We elicited prior beliefs from each participant on eight different items (listed in \tableref{tab:Items}) using three different dependent measures, which we call \emph{give-a-number}, \emph{binned histograms}, and \emph{paired comparisons}.
+Each participant rated each item using each dependent measure.
+Trial order was randomized.
+
+On  \emph{give-a-number} (GAN) trials, participants saw the context sentence and GAN question for each item (see \tableref{tab:Items}) and were asked to provide one number by adjusting a slider with endpoints labeled as the lowest and highest number for that item.
+Min and max numbers are shown in \tableref{tab:Items} and were taken from previous studies that had used these items \cite{DegenTessler2015:Wonky-worlds:-L,SchollerFranke2015:Semantic-values,KaoWu2014:Nonliteral-Unde}\jd{any others?}.
+The selected number appeared above the slider so participants knew exactly what the value of the slider would be.
+
+On \emph{binned histogram} (BH) trials, participants saw an item's context sentence and were asked to \emph{Please rate how likely it is that Y} (where Y corresponded to  the appropriate BH frame in \tableref{tab:Items}).
+They then adjusted 15 sliders, one per bin, with five points labeled \emph{impossible}, \emph{not very likely}, \emph{neutral}, \emph{very likely}, and \emph{extremely likely}.
+Bins were determined by dividing up the entire interval spanned by the item's minimum and maximum value into equally sized bins (see \tableref{tab:Items} for each item's bin size).
+
+On \emph{paired comparison} (PC) trials, participants  saw an item's context sentence and were asked to click on one of two options which were shown side by side -- whichever one they thought was more likely.
+Each option used the appropriate PC frame in\tableref{tab:Items} and numbers were filled in by comparing the following bins: 1 vs 2, 2 vs 6, 6 vs 11, 11 vs 14, and 14 vs 15.
+Paired comparison trials always occurred as a block of five trials, with order of comparisons randomized within block.
+
+Results are shown in \figref{fig:Results}. Data analysis and modeling is reported in the next section.\jd{do we need to say anything about this or just go straight to the model?}
 
@@
 \item why within-participant?
-\item what was the exact randomization / procedure?
-\item items are given in Table~\ref{tab:Items}
-\item include bins as well?
-\item what were the lighting round bin comparisons?
-  \begin{itemize}
-  \item 1 vs 2
-  \item 2 vs 6
-  \item 6 vs 11
-  \item 11 vs 14
-  \item 14 vs 15
-  \end{itemize}
-\item results are in Figure~\ref{fig:Results}
+\jd{because doing it between participants would mean not knowing if any differences in results are due to differences in beliefs or differences in the measure.
+ie we didn't want to confound participants and measures.
+do we need to talk about this?}
+%\item what was the exact randomization / procedure?
+%\item items are given in Table~\ref{tab:Items}
+%\item include bins as well?
+
+%\item results are in Figure~\ref{fig:Results}
 \end{itemize}
 
-\begin{table}
+\begin{table*}
   \centering
-  \begin{tabular}{p{8cm}}
-    1.  X has just fetched himself a cup of \textbf{coffee} from the office vending machine.
-         What do you think the temperature of his coffee is?  \\
-    2.  X \textbf{commuted} to work yesterday. 
-        How many minutes do you think she spent commuting yesterday? \\
-3.  X told a \textbf{joke} to N kids. 
-    How many of the kids do you think laughed?\\
+  \begin{tabular}{l p{2.2cm} p{2.5cm} p{3.5cm} p{3cm} p{2.5cm}}
+	\toprule
+	Item & Bins (units, min, max, step) & Context sentence & GAN question & BH frame & PC frame \\
+	\midrule
+    coffee & degrees, $<$44, $>200$, 13& X has just fetched himself a cup of coffee from the office vending machine.
+& What do you think the temperature of his coffee is?
+& his coffee was the following temperatures & The temperature of his coffee is N degrees.\\
+	commute & minutes, 0, $>$98, 7 & X commuted to work yesterday.
+& How many minutes do you think she spent commuting yesterday?
+& she commuted for the following numbers of minutes yesterday & She spent N minutes commuting.\\
+	joke & children, 0, 14, 1  & X told a joke to 14 kids. & How many of the kids do you think laughed? & the following number of kids laughed & N of the children laughed.\\
 
-4.  X bought a \textbf{laptop}.    
-    How much do you think it cost?\\
+	laptop & dollars, 0, $>$7500, 500& X bought a laptop. & How much do you think it cost? & the laptop cost the following numbers of dollars & The laptop cost \$N.\\
+    marbles & marbles, 0, 14, 1 & X threw 14 marbles into a pool. & How many of the marbles do you think sank? & the following number of marbles sank & N of the marbles sank.\\
 
-5.  X threw N \textbf{marbles} into a pool.
-    How many of the marbles do you think sank?\\
+	movies & minutes, 0, $>210$, 16& X just went to the movies to see a blockbuster.
+& How many minutes long do you think the movie was?
+& the movie was the following numbers of minutes long & The movie was N minutes long.\\
 
-6.  X just went to the \textbf{movies} to see a blockbuster.
-    How many minutes long do you think the movie was?\\
+	TV & hours, 0, $>$43, 3& X watched TV last week.
+& How many hours do you think he spent watching TV last week?
+& he watched TV for the following numbers of hours last week & He spent N hours watching TV.\\
 
-7.  X watched \textbf{TV} last week.
-    How many hours do you think he spent watching TV last week? \\
-
-8.  X bought a \textbf{watch}.
-    How much do you think it cost?
+	watch & dollars, 0, $>$750, 50 & X bought a watch.& How much do you think it cost? & it cost the following numbers of dollars & The watch cost \$N.\\
+    \bottomrule
   \end{tabular}
-  \caption{Experimental items}
+  \caption{Experimental items. \emph{X} was a randomly generated name (different on each trial). \emph{N} was one of the bins.}
   \label{tab:Items}
-\end{table}
+\end{table*}
 
@@
     \includegraphics[width = \textwidth]{plots/data_sliderNumber.pdf}
-    \caption{slider ratings \& number choices}
+    \caption{Red: proportion of choosing a number in each bin on \emph{give-a-number} trials. Black: mean slider ratings on \emph{binned histogram} trials.}
     \label{fig:slider}
@@
     \includegraphics[width = \textwidth]{plots/data_choice.pdf}
-    \caption{bin comparison}
+    \caption{Proportion of choosing the higher bin on \emph{paired comparison} trials.}
     \label{fig:lighting}
@@
 
-  \caption{Results from experiments. Error bars show bootstrapped 95\% confidence intervals.}
+  \caption{Experimental results. Error bars show bootstrapped 95\% confidence intervals.}
   \label{fig:Results}
@@
 The data we would like to explain are: (i) the normalized slider ratings $s_{ijk} \in [0;1]$ of
-subject $i \in \set{1, \dots, 20}$ for item $j \in \set{1, \dots, 8}$ and bin
-$k \in \set{1, \dots, 15}$; (ii) the bins $n_{ij} \in \set{1, \dots, 15}$ in which subject
-$i$'s number choice for item $i$ was; and (iii) the binary choices $c_{ijl} \in \set{0,1}$ of
-whether subject $i$ selected the higher bin for item $j$ in the bin comparison condition
-$l$. There two simplifications in need of commenting. In (i), we focus on slider ratings after
-normalizing for each subject, because we assume that slider adjustments reflect relative, not
+participant $i \in \set{1, \dots, 20}$ for item $j \in \set{1, \dots, 8}$ and bin
+$k \in \set{1, \dots, 15}$ from the binned histogram task; (ii) the bins $n_{ij} \in \set{1, \dots, 15}$ in which participant
+$i$'s number choice for item $i$ was in the give-a-number task; and (iii) the binary choices $c_{ijl} \in \set{0,1}$ of
+whether participant $i$ selected the higher bin for item $j$ in the paired comparison task for each comparison
+$l$. There are two simplifications in need of commenting. In (i), we focus on slider ratings after
+normalizing for each participant, because we assume that slider adjustments reflect relative, not
 absolute estimates of subjective beliefs. In (ii), we focus on bin choices, not actual number
 choices, in order to avoid, as much as possible, considerations of salience of particular
-numbers, and also because otherwise data from items with smaller domains of plausible number
+numbers, and also because otherwise data from items with smaller domains of plausible numbers
 would get more weight than data from items which allow for a wider set of number choices.
@@
 All three pieces of data are to be explained as functions of subjective beliefs $P_{ij}$, where
-$P_{ij}$ is a probability vector of length 15 with $P_{ijk}$ being subject $i$'s belief about
+$P_{ij}$ is a probability vector of length 15 with $P_{ijk}$ being participant $i$'s belief about
 the relative likelihood of bin $k$ for item $j$. Each $P_{ij}$ defines a likelihood for our
@@
 stochastic vector of length 15, which we call a ``population-level belief'' about item $j$. The
-structure of this model is pictured in Figure~\ref{fig:ModelGraph}.
+structure of this model is pictured in \figref{fig:ModelGraph}.
 
@@
 
-Finally, consider the link function for bin comparisons. We are after the likelihood with which
-subject $i$ selects the higher bin for item $j$ in bin comparison condition $l$. Suppose $l$ is
+Finally, consider the link function for bin comparisons. We are interested in the likelihood with which
+participant $i$ selects the higher bin for item $j$ in bin comparison condition $l$. Suppose $l$ is
 about comparing the lower bin $b_l$ to the higher $b_h$. The perhaps most natural approach
 would be to link the likelihood of choosing $b_h$ over $b_l$ to the difference between
-$P_{ijb_h}$ and $P_{ijb_l}$. This, however, does not appear to be what subjects were doing. For
+$P_{ijb_h}$ and $P_{ijb_l}$. This, however, does not appear to be what participants were doing. For
 example, take the ``marbles'' case where we see almost unanimous choices of bin 11 over bin 6,
-despite seeing any pronounced difference in slider ratings for these bins. Indeed, a model that
+despite seeing no pronounced difference in slider ratings for these bins. Indeed, a model that
 implements this idea failed to capture the relevant regularities in the data blatantly.
@@
 bin comparisons is the distance to the mode of $P_{ij}$: soft-max prefer the bin that is closer
-to the mode of $P_{ij}$; select random if both bins are equally far from the
+to the mode of $P_{ij}$; select randomly if both bins are equally far from the
 prototype.\footnote{Notice that
@@
 
+\jd{hm, doesn't seem too weird to me!}
+
 \section{Inference}
@@
 population-level beliefs. Inspection of posteriors of individual $P_{ij}$ shows that there is
-ample variation between subjects. Nonetheless, the way the model suggest we should think about
+ample variation between participants. Nonetheless, the way the model suggest we should think about
 harnessing the individual $P_{ij}$s under a population-level central tendency is closely
@@
 criticism is therefore important. Figure~\ref{fig:PPCs} shows posterior predictive checks at
-the population-aggregate level for all of our three task types. For the slider rating task,
-posterior predictions are spot-on. Some of the number estimation data is surprising even for
-the model trained on this very data. This could have various reasons: (i) the number data does
+the population-aggregate level for all of our three task types. For the binned histogram task,
+posterior predictions are spot-on. Some of the give-a-number data is surprising even for
+the model trained on this very data. This could have various reasons: (i) the give-a-number data does
 not have a huge influence on the posterior likelihood, (ii) number choices may be influenced by
-saliency and/or roundness of numbers after all. Finally, there is one condition in the bin
+saliency and/or roundness of numbers after all. Finally, there is one condition in the paired
 comparison task that the model definitely got wrong. This is the choice of what is more likely:
@@
 almost everybody should answer that it is more likely that one marble sank. But that is not
-what we observe. It may be that subjects revise beliefs about ``normality'' of the marbles,
-while holding on to an assumption that all marbles behave the same
+what we observe. It may be that participants revise beliefs about ``normality'' of the marbles,
+while holding on to an assumption that all marbles behave in the same way 
 \cite{DegenTessler2015:Wonky-worlds:-L}.
@@
     \includegraphics[width = \textwidth]{plots/ppc_slider.pdf}
-    \caption{slider rating}
+    \caption{Binned histogram task.}
     \label{fig:sliderPPC}
@@
     \includegraphics[width = \textwidth]{plots/ppc_number.pdf}
-    \caption{number estimation}
+    \caption{Give-a-number task.}
     \label{fig:numberPPC}
@@
     \includegraphics[width = \textwidth]{plots/ppc_choice.pdf}
-    \caption{bin comparison}
+    \caption{Paired comparison task.}
     \label{fig:lightingPPC}
@@
 aggregate population level well. To have a more fine-grained measure of model fit, we also
-looked at posterior predictive $p$-values at the level of subjects and items. \ndg{need to say what this means -- i'm not even sure that i know!} Unfortunately,
-only the slider rating task, provides enough data for such a fine-grained analysis. For the
-slider rating task, fixing a subject and an item, observations and replicates are probability
+looked at posterior predictive $p$-values at the level of participants and items. \ndg{need to say what this means -- i'm not even sure that i know!} Unfortunately,
+only the slider rating task provides enough data for such a fine-grained analysis. For the
+slider rating task, fixing a participant and an item, observations and replicates are probability
 vectors of length 15. In a first analysis, we used the mean of these probability vectors as a
-test statistic. The minimum posterior predictive $p$-value over all 20 (subjects) times 8
+test statistic. The minimum posterior predictive $p$-value over all 20 (participants) times 8
 (items) cases was 0.13, suggesting that the means of observed $s_{ij}$ are non-surprising to
@@
 MF's work was supported by the Institutional Strategy of the University of T̈\"{u}bingen (Deutsche
-Forschungsgemeinschaft, ZUK 63) and the Priority Program XPrag.de (DFG Schwerpunktprogramm
-1727).
+Forschungsgemeinschaft, ZUK 63). MF and JD's work was supported by the Priority Program XPrag.de (DFG Schwerpunktprogramm
+1727). This work was further supported by ONR grant N00014-13-1-0788 and a James S. McDonnell Foundation Scholar Award to NDG and an SNF Early Postdoc.Mobility Award to JD.
 

### 2016-01-31 text/01_CogSci_abstract/paper.tex
add watch example bins
intro rewrite

@@
 \newcommand{\jd}[1]{\textcolor{Blue}{[jd: #1]}} 
+\definecolor{Red}{RGB}{178,34,34}
+\newcommand{\mf}[1]{\textcolor{Red}{[mf: #1]}}
 
@@
 
-\todo[inline]{Each author's name should appear
-on a separate line, 11~point bold, and centered, with the author's
-email address in parentheses. Under each author's name list the
-author's affiliation and postal address in ordinary 10~point type.}
+\begin{abstract}
+  Experimental participants' beliefs about everyday events are of theoretical interest on their
+  own and also an important ingredient for model building---especially in Bayesian models of
+  phenomena such as logical reasoning, future predictions, and language use. Here, we explore
+  several recently used methods for measuring subjective beliefs about unidimensional
+  contiguous properties, such as the likely price of a new watch. We use a hierarchical
+  Bayesian model for inferring likely ``population-level beliefs'' as the central tendency of
+  participants' individual-level beliefs.  Three different dependent measures are related to
+  individual-level beliefs through likelihood ``linking functions:'' (i) judgements of
+  (relative) likelihood of intervals of values via continuous sliders, (ii) a give-a-number
+  task, and (iii) choice of the more likely of two intervals of values. Our results suggest
+  that using averaged normalized slider ratings for binned quantities is a practical and fairly
+  good approximator of inferred population-level beliefs.
 
-\todo[inline]{who is an author? which order?}
-
-\begin{abstract}
-  Experimental participants' beliefs about everyday events are of theoretical interest on their own and also an
-  important ingredient for model building---especially in Bayesian models of phenomena such as logical
-  reasoning, future predictions, and language use. Here, we explore several recently used 
-  methods for measuring subjective beliefs about unidimensional properties. \ndg{say what kind of subjective beliefs /distributions we look at more clearly.} 
-  We a hierarchical Bayesian data analysis model for inferring likely ``population-level beliefs'' as the central
-  tendency of participants' individual-level beliefs. 
-  Three different dependent measures---\ndg{name them?}---are related to individual-level beliefs through likelihood ``linking functions.''
-%  functions for three types of task measures. 
-  Our results suggest that the practice of
-  using averaged normalized slider ratings for binned quantities is a practical and fairly good
-  approximator of inferred population-level beliefs.
-  Other dependent measures yield less reliable results.
-\textbf{Keywords:} subjective beliefs, hierarchical modeling, Bayesian data analysis, Bayesian
-cognitive models 
+  \textbf{Keywords:} subjective beliefs, hierarchical modeling, Bayesian data analysis,
+  Bayesian cognitive models
 \end{abstract}
@@
 also in science. Scientific ascription of latent mental states plays an important role in many
-explanations of higher-order cognition: decision making, planning, reasoning,
-and language use. Naturally, then, it becomes important to validate any method for eliciting mental states, 
-to be used for explanatory purposes.
+explanations of higher-order cognition: decision making, planning, reasoning, and language
+use. It is therefore vital to have methods for validating explanatory mental states
+ascriptions.
 
 A family of models where this is particularly pressing are Bayesian models of cognition which
-seek to explain task behavior in a variety of domains as partially informed by what participants
-believe about mundane events---their \emph{prior beliefs}. Take interpretation of language. Empirical data on whether a
-statement like ``That watch cost $n$ dollars'' is understood to convey speaker affect, rather
-than literal meaning, can be explained well by a Bayesian model of utterance interpretation
-\citep{KaoWu2014:Nonliteral-Unde}. In this model a crucial role is played by
-an empirical measure of participants' expectations about the likely or normal price of a
-watch. 
-Other examples of domains in which empirically successful models have included some
-measure of participants' belief include making predictions about everyday events 
-\citep{GriffithsTenenbaum2006:Optimal-Predict}, the strength of pragmatic enrichments
-\citep{DegenTessler2015:Wonky-worlds:-L}, the interpretation of vague quantifiers
-\citep{SchollerFranke2015:Semantic-values}, \textcolor{red}{ \dots }
-
-\todo[inline]{fill me}
+seek to explain task behavior in a variety of domains as partially informed by what
+participants believe about mundane events---their \emph{prior beliefs}. Take interpretation of
+language. Empirical data on whether a statement like ``That watch cost $n$ dollars'' is
+understood to convey speaker affect, rather than literal meaning, can be explained well by a
+Bayesian model of utterance interpretation \citep{KaoWu2014:Nonliteral-Unde}. In this model a
+crucial role is played by an empirical measure of participants' expectations about the likely
+or normal price of a watch.  Other examples of domains in which empirically successful models
+have included a measure of participants' prior beliefs include making predictions about
+everyday events \citep{GriffithsTenenbaum2006:Optimal-Predict}, referential communication
+\citep{FrankGoodman2012:Predicting-Prag}, the strength of pragmatic enrichments
+\citep{DegenTessler2015:Wonky-worlds:-L}, or the interpretation of vague quantifiers
+\citep{SchollerFranke2015:Semantic-values}. \mf{anyone wants any other references here?}
 
 Many methods of assessing subjective beliefs exist. One approach is to take actual frequencies
-as an approximation \citep{GriffithsTenenbaum2006:Optimal-Predict}. This will not
-work when sufficient real-world data is unavailable---for instance, for beliefs about imaginary events. 
-Even where available, real-world
-frequencies may deviate systematically from participants' beliefs, and this can be problematic for
-the predictions of a model that relies on subjective beliefs.
-%\citep{MarcusDavis2013:How-Robust-Are-}.
+as an approximation to subjective beliefs
+\citep{GriffithsTenenbaum2006:Optimal-Predict}. Unfortunately, the relevant data may not be
+available (e.g., for one-shot or imaginary events) and real-world frequencies may deviate from
+participants' subjective beliefs for many reasons.
 
-Another approach is to experimentally elicit relevant beliefs. Many techniques for this
-exist, especially in the economics literature
-\citep{MorganHenrion1990:Uncertainty:-A-,Manski2004:Measuring-Expec,SchlagTremewan2014:A-penny-for-you,AndersenFountain2014:Estimating-Subj}. The
-perhaps most prominent method uses so-called \emph{scoring rules}
-\citep{Savage1971:Elicitation-of-,SchlagTremewan2014:A-penny-for-you}. Roughly speaking,
-scoring rules are particular schemes of rewarding participants' answers in such a way as to
-make sure that responses are what the experimenter wants them to be: the actual subjective probability of an
-event, the mean of a distribution, a credible interval etc. Scoring rules are highly faithful
-but require participants to become sufficiently familiar with the (usually probabilistic)
-payoff scheme. Scoring rules also must make assumptions about how much participants value
-certain probabilistic gambles and may therefore require elaborate means of data analysis
-\citep{AndersenFountain2014:Estimating-Subj}. Similar practical concerns apply to other methods
-as well, such as the use of \emph{iterated learning} for prior estimation
-\citep{LewandowskyGriffiths2009:The-Wisdom-of-I}.
+Another possibility is to empirically measure subjective beliefs by ``give-a-number''
+tasks. The simplest version would be this: being told that John just bought a new watch,
+subjects are asked for a single numerical estimate of its price. This task is easy to
+comprehend and implement, but a single number does not provide much information about the
+subjective belief that it is (arguably) a manifestation of. One solution is to infer which
+parameterized distribution best explains the observed number choices, either at the individual
+level \citep{Manski2004:Measuring-Expec} or on the population level
+\citep{TauberSteyvers2013:Inferring-Subje}.
 
-For common applications, many of these procedures may be excessively complex or too demanding
-on both researcher and participant. It would be beneficial if there was an easily applicable
-technique of eliciting subjective beliefs that is good enough for testing the general
-predictions of models that rely on these measures for their predictions. A further complication
-is that it may be impractical to obtain, from the same participant, data on subjective beliefs
-and data on the task of interest, especially if exposure to one of these measures is likely to
-affect the other. This suggests that, for practical reasons, there is a demand for
-reliable-enough measures of what we will  call ``population-level beliefs'': estimates
-of the central tendency or average of subjective beliefs in a given population.
+More sophisticated ``give-a-number'' tasks give more information, but can be rather complex and
+more difficult to implement. A prominent method in economics uses \emph{scoring rules}
+\citep{Savage1971:Elicitation-of-,AndersenFountain2014:Estimating-Subj,SchlagTremewan2014:A-penny-for-you}. Roughly
+put, scoring rules are schemes of rewarding participants' answers that make sure that responses
+(by rational subjects) faithfully reflect a measure of interest: the actual subjective
+probability of an event, the mean of a distribution, a credible interval etc. Scoring-rule
+tasks require participants to become sufficiently familiar with the (usually probabilistic)
+payoff scheme. Another insightful, but also complex approach is to use a cleverly designed
+\emph{iterated learning paradigm} \citep{LewandowskyGriffiths2009:The-Wisdom-of-I}.
 
-A very simple technique that has been used with apparent success is to have participants adjust
-sliders in order to express their (relative) levels of support for values of uncertain
-contiguous quantities
-\citep{KaoWu2014:Nonliteral-Unde,SchollerFranke2015:Semantic-values}. \textcolor{red}{ MORE REFERENCES } Participants are confronted with a short and simple description of a situation or an
-event, such as ``John bought a watch,'' and asked for their estimates of the watch's price. To
-give their estimate, they are asked to adjust sliders whose endpoints are labelled as ``extremely
-likely'' and ``extremely unlikely,'' for a number of bins that partition the range of plausible
-values (established by a pre-test). E.g., for the price of a watch bins could be \jd{erin: fill in}
-\textcolor{red}{ FILL ME }
+In sum, there is tradeoff between simplicity of paradigms and their informativity. Ideally, we
+would like to have an experimental method for measuring subjective beliefs that (i) gives us
+sufficient and reliable-enough information about subjective beliefs to derive testable
+predictions from cognitive models that rely on such information, (ii) is easy to understand by
+the participants, (iii) is easy to implement by researchers whose focus and expertise is not
+belief-elicitation as such, and that (iv) does not rely on sophisticated methods of extracting
+the relevant information from the observed data. Moreover, the ideal method would be (v)
+flexible enough to allow inferred subjective belief distributions beyond standard parameterized
+distributions and (vi) would provide us with a reasonable approximation of a ``population-level
+belief,'' i.e., an estimate of the central tendency or average of subjective beliefs in a given
+population. The latter allows to use one set of subjects for measuring prior beliefs and
+another for whatever other task is of interest.
+
+A simple technique that seems to fit this bill is the \emph{binned histograms} task that has
+recently been used with apparent success
+\citep{KaoWu2014:Nonliteral-Unde,SchollerFranke2015:Semantic-values}. \mf{someone please add
+  references: where did you use binned sliders?} Participants adjust sliders in order to
+express their (relative) subjective beliefs about how likely it is that the value of an
+uncertain contiguous quantity lies in some interval of possible values (a bin). E.g.,
+participants read that John bought a watch.  To report their beliefs about the watch's price,
+they adjust sliders whose endpoints are labelled as ``extremely likely'' and ``extremely
+unlikely.'' Each slider corresponds to on out of 15 bins that partition the range of plausible
+values (established by a pre-test), such as 0--50\$, 50--100\$, \dots 700--750\$. Each
+subject's slider ratings are normalized within each item, and the results are averaged across
+subjects to obtain an estimate of the population's average belief about each item.
 
 \ndg{should we also talk about the other dependent measures? i think part of what's interesting is comparing the different dependent measures to each other?}
-
-\ndg{say more clearly that we focus on unidimensional distributions and why.}
 
@@
 Trial order was randomized.
-
-On  \emph{give-a-number} (GAN) trials, participants saw the context sentence and GAN question for each item (see \tableref{tab:Items}) and were asked to provide one number by adjusting a slider with endpoints labeled as the lowest and highest number for that item.
-Min and max numbers are shown in \tableref{tab:Items} and were taken from previous studies that had used these items \cite{DegenTessler2015:Wonky-worlds:-L,SchollerFranke2015:Semantic-values,KaoWu2014:Nonliteral-Unde}\jd{any others?}.
-The selected number appeared above the slider so participants knew exactly what the value of the slider would be.
-
-On \emph{binned histogram} (BH) trials, participants saw an item's context sentence and were asked to \emph{Please rate how likely it is that Y} (where Y corresponded to  the appropriate BH frame in \tableref{tab:Items}).
-They then adjusted 15 sliders, one per bin, with five points labeled \emph{impossible}, \emph{not very likely}, \emph{neutral}, \emph{very likely}, and \emph{extremely likely}.
-Bins were determined by dividing up the entire interval spanned by the item's minimum and maximum value into equally sized bins (see \tableref{tab:Items} for each item's bin size).
-
-On \emph{paired comparison} (PC) trials, participants  saw an item's context sentence and were asked to click on one of two options which were shown side by side -- whichever one they thought was more likely.
-Each option used the appropriate PC frame in\tableref{tab:Items} and numbers were filled in by comparing the following bins: 1 vs 2, 2 vs 6, 6 vs 11, 11 vs 14, and 14 vs 15.
-Paired comparison trials always occurred as a block of five trials, with order of comparisons randomized within block.
-
-Results are shown in \figref{fig:Results}. Data analysis and modeling is reported in the next section.\jd{do we need to say anything about this or just go straight to the model?}
-
-\begin{itemize}
-\item why within-participant?
-\jd{because doing it between participants would mean not knowing if any differences in results are due to differences in beliefs or differences in the measure.
-ie we didn't want to confound participants and measures.
-do we need to talk about this?}
-%\item what was the exact randomization / procedure?
-%\item items are given in Table~\ref{tab:Items}
-%\item include bins as well?
-
-%\item results are in Figure~\ref{fig:Results}
-\end{itemize}
 
@@
 \end{table*}
+
+On  \emph{give-a-number} (GAN) trials, participants saw the context sentence and GAN question for each item (see \tableref{tab:Items}) and were asked to provide one number by adjusting a slider with endpoints labeled as the lowest and highest number for that item.
+Min and max numbers are shown in \tableref{tab:Items} and were taken from previous studies that had used these items \cite{DegenTessler2015:Wonky-worlds:-L,SchollerFranke2015:Semantic-values,KaoWu2014:Nonliteral-Unde}\jd{any others?}.
+The selected number appeared above the slider so participants knew exactly what the value of the slider would be.
+
+On \emph{binned histogram} (BH) trials, participants saw an item's context sentence and were asked to \emph{Please rate how likely it is that Y} (where Y corresponded to  the appropriate BH frame in \tableref{tab:Items}).
+They then adjusted 15 sliders, one per bin, with five points labeled \emph{impossible}, \emph{not very likely}, \emph{neutral}, \emph{very likely}, and \emph{extremely likely}.
+Bins were determined by dividing up the entire interval spanned by the item's minimum and maximum value into equally sized bins (see \tableref{tab:Items} for each item's bin size).
+
+On \emph{paired comparison} (PC) trials, participants  saw an item's context sentence and were asked to click on one of two options which were shown side by side -- whichever one they thought was more likely.
+Each option used the appropriate PC frame in\tableref{tab:Items} and numbers were filled in by comparing the following bins: 1 vs 2, 2 vs 6, 6 vs 11, 11 vs 14, and 14 vs 15.
+Paired comparison trials always occurred as a block of five trials, with order of comparisons randomized within block.
+
+Results are shown in \figref{fig:Results}. Data analysis and modeling is reported in the next section.\jd{do we need to say anything about this or just go straight to the model?}
+
+\begin{itemize}
+\item why within-participant?
+\jd{because doing it between participants would mean not knowing if any differences in results are due to differences in beliefs or differences in the measure.
+ie we didn't want to confound participants and measures.
+do we need to talk about this?}
+%\item what was the exact randomization / procedure?
+%\item items are given in Table~\ref{tab:Items}
+%\item include bins as well?
+
+%\item results are in Figure~\ref{fig:Results}
+\end{itemize}
 
@@
 
+\begin{itemize}
+\item \ndg{say more clearly that we focus on unidimensional distributions and why.}
+\end{itemize}
+
 \section{Acknowledgments}

### 2016-01-31 text/01_CogSci_abstract/paper.tex
motivation minor edits

@@
 subject's slider ratings are normalized within each item, and the results are averaged across
-subjects to obtain an estimate of the population's average belief about each item.
+subjects to obtain an estimate of the population's average belief about each item. These
+averaged normalized slider ratings per bin (henceforth: mean slider ratings) look intuitively
+plausible and give good results if fed into cognitive models that predict behavior in language
+or reasoning tasks, for instance.
 
-\ndg{should we also talk about the other dependent measures? i think part of what's interesting is comparing the different dependent measures to each other?}
-
-\begin{itemize}
-\item explain slider rating task
-\item our goal here is to validate whether ``population-level beliefs'' are approximated well
-  enough by normalized average slider ratings
-\item use hierarchical Bayesian modeling
-  \begin{itemize}
-  \item calculate likelihoods from subjective beliefs
-  \item each subjective belief is drawn from a population-level central tendency, which is a
-    ``belief'' as well
-  \end{itemize}
-\end{itemize}
+Explanatory success notwithstanding, the question remains whether mean slider ratings measure
+what we would like them to. Are these good approximations of a central tendency of subjects'
+individual beliefs? Are these measures consistent with subjects' behavior in other tasks, such
+as \emph{give-a-number}, as well? To address this issue we collected data in a within-subject
+from three task types: (i) \emph{binned histograms}, (ii) \emph{give-a-number} and (iii)
+\emph{paired comparisons}, i.e., a direct comparison of two bins from the \emph{binned
+  histograms} task. Using a Bayesian hierarchical model, we used the data from all three
+dependent measures to infer likely latent subjective beliefs, where each subjective belief is,
+intuitively, a noisy perturbation of a population-level belief. We find that mean \emph{a
+  posteriori} population-level beliefs are approximated well by mean slider ratings, suggesting
+that the latter may indeed be a plausible, yet cheap measure of ``what the crowd beliefs.''
+Moreover, the model is able to capture subjects' behavior in all three task types reasonably
+well, suggesting that what mean sliders measure is what we think it is: a population-level
+central tendency of latent subjective beliefs.
 
@@
 \begin{table*}
+  \footnotesize
   \centering

### 2016-01-31 text/01_CogSci_abstract/paper.tex

@@
 
-Another possibility is to empirically measure subjective beliefs by ``give-a-number''
+Another possibility is to empirically measure subjective beliefs by \emph{give-a-number}
 tasks. The simplest version would be this: being told that John just bought a new watch,
@@
 
-More sophisticated ``give-a-number'' tasks give more information, but can be rather complex and
-more difficult to implement. A prominent method in economics uses \emph{scoring rules}
+More sophisticated \emph{give-a-number} tasks give more information, but can be rather complex
+and more difficult to implement. A prominent method in economics uses \emph{scoring rules}
 \citep{Savage1971:Elicitation-of-,AndersenFountain2014:Estimating-Subj,SchlagTremewan2014:A-penny-for-you}. Roughly
@@
 they adjust sliders whose endpoints are labelled as ``extremely likely'' and ``extremely
-unlikely.'' Each slider corresponds to on out of 15 bins that partition the range of plausible
-values (established by a pre-test), such as 0--50\$, 50--100\$, \dots 700--750\$. Each
-subject's slider ratings are normalized within each item, and the results are averaged across
-subjects to obtain an estimate of the population's average belief about each item. These
-averaged normalized slider ratings per bin (henceforth: mean slider ratings) look intuitively
-plausible and give good results if fed into cognitive models that predict behavior in language
-or reasoning tasks, for instance.
+unlikely.'' Each slider corresponds to one out of 15 bins that partition the range of plausible
+values (established by a pre-test). E.g., for the price of a watch bins could be ``\$0-\$50'',
+``\$50-\$100'', etc. up to ``\$700-\$750'' and ``more than \$750''. Each subject's slider
+ratings are normalized within each item, and the results are averaged across subjects to obtain
+an estimate of the population's average belief about each item. These averaged normalized
+slider ratings per bin (henceforth: mean slider ratings) look intuitively plausible and give
+good results if fed into cognitive models that predict behavior in language or reasoning tasks,
+for instance.
 

### 2016-01-31 text/01_CogSci_abstract/paper.tex
minor edits model section

@@
 
-We elicited prior beliefs from each participant on eight different items (listed in \tableref{tab:Items}) using three different dependent measures, which we call \emph{give-a-number}, \emph{binned histograms}, and \emph{paired comparisons}.
-Each participant rated each item using each dependent measure.
-Trial order was randomized.
+We elicited prior beliefs for eight different items (listed in
+\tableref{tab:Items}) using three different dependent measures: \emph{binned histograms},
+\emph{give-a-number} and \emph{paired comparisons}. Each participant rated each item using each
+dependent measure. Trial order was randomized.
 
@@
 	\midrule
-    coffee & degrees, $<$44, $>200$, 13& X has just fetched himself a cup of coffee from the office vending machine.
+    coffee & degrees, $<$44, $>200$, 12& X has just fetched himself a cup of coffee from the office vending machine.
 & What do you think the temperature of his coffee is?
-& his coffee was the following temperatures & The temperature of his coffee is N degrees.\\
+& His coffee was the following temperatures & The temperature of his coffee is N degrees.\\
 	commute & minutes, 0, $>$98, 7 & X commuted to work yesterday.
 & How many minutes do you think she spent commuting yesterday?
-& she commuted for the following numbers of minutes yesterday & She spent N minutes commuting.\\
+& She commuted for the following numbers of minutes yesterday & She spent N minutes commuting.\\
 	joke & children, 0, 14, 1  & X told a joke to 14 kids. & How many of the kids do you think laughed? & the following number of kids laughed & N of the children laughed.\\
 
-	laptop & dollars, 0, $>$7500, 500& X bought a laptop. & How much do you think it cost? & the laptop cost the following numbers of dollars & The laptop cost \$N.\\
-    marbles & marbles, 0, 14, 1 & X threw 14 marbles into a pool. & How many of the marbles do you think sank? & the following number of marbles sank & N of the marbles sank.\\
+	laptop & dollars, 0, $>$7500, 500& X bought a laptop. & How much do you think it cost? & The laptop cost the following numbers of dollars & The laptop cost \$N.\\
+    marbles & marbles, 0, 14, 1 & X threw 14 marbles into a pool. & How many of the marbles do you think sank? & The following number of marbles sank & N of the marbles sank.\\
 
@@
 & How many minutes long do you think the movie was?
-& the movie was the following numbers of minutes long & The movie was N minutes long.\\
+& The movie was the following numbers of minutes long & The movie was N minutes long.\\
 
@@
 & How many hours do you think he spent watching TV last week?
-& he watched TV for the following numbers of hours last week & He spent N hours watching TV.\\
+& He watched TV for the following numbers of hours last week & He spent N hours watching TV.\\
 
-	watch & dollars, 0, $>$750, 50 & X bought a watch.& How much do you think it cost? & it cost the following numbers of dollars & The watch cost \$N.\\
+	watch & dollars, 0, $>$750, 50 & X bought a watch.& How much do you think it cost? & It cost the following numbers of dollars & The watch cost \$N.\\
     \bottomrule
@@
 
-On  \emph{give-a-number} (GAN) trials, participants saw the context sentence and GAN question for each item (see \tableref{tab:Items}) and were asked to provide one number by adjusting a slider with endpoints labeled as the lowest and highest number for that item.
-Min and max numbers are shown in \tableref{tab:Items} and were taken from previous studies that had used these items \cite{DegenTessler2015:Wonky-worlds:-L,SchollerFranke2015:Semantic-values,KaoWu2014:Nonliteral-Unde}\jd{any others?}.
-The selected number appeared above the slider so participants knew exactly what the value of the slider would be.
+On \emph{give-a-number} (GAN) trials, participants saw the context sentence and GAN question
+for each item (see \tableref{tab:Items}) and were asked to provide one number by adjusting a
+slider with endpoints labeled as the lowest and highest number for that item. Min and max
+numbers are shown in \tableref{tab:Items} and were taken from previous studies that had used
+these items
+\cite{DegenTessler2015:Wonky-worlds:-L,SchollerFranke2015:Semantic-values,KaoWu2014:Nonliteral-Unde}. The
+selected number appeared above the slider so participants knew exactly what the value of the
+slider would be.
 
-On \emph{binned histogram} (BH) trials, participants saw an item's context sentence and were asked to \emph{Please rate how likely it is that Y} (where Y corresponded to  the appropriate BH frame in \tableref{tab:Items}).
-They then adjusted 15 sliders, one per bin, with five points labeled \emph{impossible}, \emph{not very likely}, \emph{neutral}, \emph{very likely}, and \emph{extremely likely}.
-Bins were determined by dividing up the entire interval spanned by the item's minimum and maximum value into equally sized bins (see \tableref{tab:Items} for each item's bin size).
+On \emph{binned histogram} (BH) trials, participants saw an item's context sentence and were
+asked to \emph{Please rate how likely it is that Y} (where Y corresponded to the appropriate BH
+frame in \tableref{tab:Items}). They then adjusted 15 sliders, one per bin, with five points
+labeled \emph{impossible}, \emph{not very likely}, \emph{neutral}, \emph{very likely}, and
+\emph{extremely likely}. Bins were determined by dividing up the entire interval spanned by the
+item's minimum and maximum value into equally sized bins (\tableref{tab:Items}).
 
-On \emph{paired comparison} (PC) trials, participants  saw an item's context sentence and were asked to click on one of two options which were shown side by side -- whichever one they thought was more likely.
-Each option used the appropriate PC frame in\tableref{tab:Items} and numbers were filled in by comparing the following bins: 1 vs 2, 2 vs 6, 6 vs 11, 11 vs 14, and 14 vs 15.
-Paired comparison trials always occurred as a block of five trials, with order of comparisons randomized within block.
+On \emph{paired comparison} (PC) trials, participants saw an item's context sentence and had to
+click on one of two options shown side by side -- whichever one they thought was more likely.
+Each option used the appropriate PC frame in\tableref{tab:Items} and numbers were filled in by
+comparing the following bins: 1 vs 2, 2 vs 6, 6 vs 11, 11 vs 14, and 14 vs 15. Paired
+comparison trials always occurred as a block of five trials, with order of comparisons
+randomized within block.
 
-Results are shown in \figref{fig:Results}. Data analysis and modeling is reported in the next section.\jd{do we need to say anything about this or just go straight to the model?}
-
-\begin{itemize}
-\item why within-participant?
-\jd{because doing it between participants would mean not knowing if any differences in results are due to differences in beliefs or differences in the measure.
-ie we didn't want to confound participants and measures.
-do we need to talk about this?}
-%\item what was the exact randomization / procedure?
-%\item items are given in Table~\ref{tab:Items}
-%\item include bins as well?
-
-%\item results are in Figure~\ref{fig:Results}
-\end{itemize}
+Results are shown in \figref{fig:Results}. \figref{fig:Results} shows mean slider ratings in
+relation to frequencies of number choices in the \emph{give-a-number} task. It looks like the
+latter could be samples from the former, with a tendency to modal choices. But, despite the use
+of slider, some \emph{give-a-number} results still seem to reflect a tendency towards round or
+salient numbers. E.g., in item \emph{coffee} more than 40\% of participants guessed that a
+coffee from the vending machine was ca. 150 degrees Fahrenheit (bin 10). The results from the
+\emph{paired comparison} seem consistent with the mean slider ratings as well, albeit not as
+straightforwardly as one might think. This is clearest from the \emph{marbles} case. While the
+mean slider ratings indicate that bins 2, 6 and 11 were all felt to be equally likely on
+average, almost every participant chose the higher bin in direct bin comparisons 2 vs 6 and 6
+vs 11. A potential explanation for this, which we will implement formally in the
+data-generating model, is that the \emph{paired comparison} condition invited subjects to think
+about what they thought was mostly likely the case (usually: bin 15 in the
+\emph{marbles} case) and to choose the bin that is closest to that.
 
@@
 
-All three pieces of data are to be explained as functions of subjective beliefs $P_{ij}$, where
-$P_{ij}$ is a probability vector of length 15 with $P_{ijk}$ being participant $i$'s belief about
-the relative likelihood of bin $k$ for item $j$. Each $P_{ij}$ defines a likelihood for our
-data, via appropriate link functions (to be spelled out presently). Variance in subjective
-beliefs is harnessed by a population-level hyper-prior with central tendency $Q_{j}$, i.e., a
-stochastic vector of length 15, which we call a ``population-level belief'' about item $j$. The
+All three pieces of data are to be explained as functions of subjective beliefs $P_{ij}$, with
+$P_{ijk}$ being participant $i$'s belief about the relative likelihood of bin $k$ for item
+$j$. Each $P_{ij}$ defines a likelihood for our data, via appropriate link functions (to be
+spelled out presently). Variance in subjective beliefs is harnessed by a population-level
+hyper-prior with central tendency $Q_{j}$, i.e., a stochastic vector of length 15. The
 structure of this model is pictured in \figref{fig:ModelGraph}.
@@
 
-Finally, consider the link function for bin comparisons. We are interested in the likelihood with which
-participant $i$ selects the higher bin for item $j$ in bin comparison condition $l$. Suppose $l$ is
-about comparing the lower bin $b_l$ to the higher $b_h$. The perhaps most natural approach
-would be to link the likelihood of choosing $b_h$ over $b_l$ to the difference between
-$P_{ijb_h}$ and $P_{ijb_l}$. This, however, does not appear to be what participants were doing. For
-example, take the ``marbles'' case where we see almost unanimous choices of bin 11 over bin 6,
-despite seeing no pronounced difference in slider ratings for these bins. Indeed, a model that
-implements this idea failed to capture the relevant regularities in the data blatantly.
-
-Another plausible link function that accommodates for this is to assume that what matters in
-bin comparisons is the distance to the mode of $P_{ij}$: soft-max prefer the bin that is closer
-to the mode of $P_{ij}$; select randomly if both bins are equally far from the
-prototype.\footnote{Notice that
+Finally, consider the link function for bin comparisons. We are interested in the likelihood
+with which participant $i$ selects the higher bin for item $j$ in bin comparison condition
+$l$. Suppose $l$ is about comparing the lower bin $b_l$ to the higher $b_h$. The perhaps most
+natural approach would be to link the likelihood of choosing $b_h$ over $b_l$ to the difference
+between $P_{ijb_h}$ and $P_{ijb_l}$. However, as discussed above, This does not appear to be
+what participants were doing. Indeed, a model that implements this idea failed to capture the
+relevant regularities in the data blatantly. Another plausible link function is to assume that
+what matters in bin comparisons is the distance to the mode of $P_{ij}$: soft-max prefer the
+bin that is closer to the mode of $P_{ij}$; select randomly if both bins are equally far from
+the prototype.\footnote{Notice that
   $(1 + \exp(2b(1-p^{\text{high}}_{ijl})) )^{-1} = \frac{\exp(b p^{\text{high}})}{\exp(b
@@
 
-\ndg{wow, this is weird! you tried softmax ratio of bin probs?}
-
-\jd{hm, doesn't seem too weird to me!}
-
 \section{Inference}
@@
 convergence according to $\hat{R}$ \cite{GelmanRubin1992:Inference-from-}.
+
+Means and bounds of 95\% HDIs of the posteriors for model parameters are in
+Table~\ref{tab:SummaryStats}. Posterior credible levels of $w$ allow for a limited amount of
+slack around $Q_j$. Values for $\kappa$ indicate that, on average, subjects may not have had a
+preference for or against extreme slider ratings. Relatively high values of $a$ indicate that
+subjects, on average, had a strong tendency to choose their mode in the \emph{give-a-number}
+task.
+\begin{table}[ht]
+\centering
+\begin{tabular}{rlllll}
+ & $w$ & $\kappa$ & $\sigma$ & $a$ & $b$ \\ \hline
+  lower & 14.65 &  0.98 &  0.26 & 22.04 &  1.11 \\ 
+  mean & 15.55 &  0.99 &  0.28 & 27.43 &  1.27 \\ 
+  upper & 16.48 &  0.99 &  0.31 & 32.95 &  1.42 \\ 
+   \hline
+\end{tabular}
+\caption{Summary statistics for posterior on parameters}
+\label{tab:SummaryStats}
+\end{table}
 
@@
 way of assessing what the crowd believes despite individual differences.
-
-\ndg{it'd be nice to see some posteriors for some of the (more meaningful) parameters. eg w, kappa, ...}
-
-\ndg{what about the success of the other DMs? looks like give-n is not so hot. worth saying that if so.}
 

### 2016-01-31 prag_oldDev/plotting.R

@@ -183,20 +183,44 @@ plotChoicesPPC = ggplot(choice_ppc, aes(x = bin, y = mean)) + geom_point() + geo
   facet_wrap(~ item, scale = "free") + 
   geom_ribbon(aes(ymin=low,ymax=high), fill="gray", alpha="0.5") +
   geom_line( aes(x = bin, y = yemp) , color = "firebrick") + geom_point( aes(x = bin, y = yemp) , color = "firebrick") +
-  xlab("condition") + ylab("choice prop. higher bin")
+  xlab("condition") + ylab("choice prop. higher bin") + 
+  scale_x_discrete(labels=c("1" = "1 vs 2", 
+                            "2" = "2 vs 6",
+                            "3" = "6 vs 11",
+                            "4" = "11 vs 14",
+                            "5" = "14 vs 15"
+  ))
 plotChoicesPPCPaper = ggplot(choice_ppc, aes(x = bin, y = mean)) + geom_point() + geom_line() +
   facet_wrap(~ item, scale = "fixed", ncol = 4) + 
   geom_ribbon(aes(ymin=low,ymax=high), fill="gray", alpha="0.5") +
   geom_line( aes(x = bin, y = yemp) , color = "firebrick") + geom_point( aes(x = bin, y = yemp) , color = "firebrick") +
-  xlab("condition") + ylab("choice prop. higher bin")
-plotChoicesData = ggplot(choice_ppc, aes(x = bin, y = yemp)) + geom_bar(stat = "identity", fill = "lightgray") +
+  xlab("condition") + ylab("choice prop. higher bin") + 
+  scale_x_discrete(labels=c("1" = "1&2", 
+                            "2" = "2&6",
+                            "3" = "6&11",
+                            "4" = "11&14",
+                            "5" = "14&15"
+  )) + theme(axis.text.x = element_text( angle=-25))
+plotChoicesData = ggplot(choice_ppc, aes(x = factor(bin), y = yemp)) + geom_bar(stat = "identity", fill = "lightgray") +
   facet_wrap(~ item, scale = "free") +
   geom_errorbar(aes(ymin = cilow, ymax = cihigh), color = "darkgray", width = .5) +
-  xlab("condition") + ylab("choce prop. higher bin")
-plotChoicesDataPaper = ggplot(choice_ppc, aes(x = bin, y = yemp)) + geom_bar(stat = "identity", fill = "lightgray") +
+  xlab("condition") + ylab("choce prop. higher bin") + 
[…]

### 2016-01-31 prag_oldDev/postprocessing.R

@@ -1,6 +1,7 @@
 library('coda')
 library('ggmcmc')
 library('jagsUI') # for parallel computing
+library('xtable')
 source('helpers/helpers.R')
 # source('~/Desktop/data/svn/ProComPrag/dev_tmp/typicality_quantifiers/model/helpers.r')
 source('process_data.R')
@@ -15,6 +16,31 @@ if (readFlag){
 } else {
   # source('main.R')
 }
+
+## get summary data
+statsSum = data.frame(
+  parameter = c("w", "kappa", "sigma", "a", "b"),
+  hdiMean = c(mean(out$sims.list$w), 
+              mean(out$sims.list$k), 
+              mean(out$sims.list$sigma),
+              mean(out$sims.list$a),
+              mean(out$sims.list$b)),
+  hdiLow = c(HDIofMCMC(out$sims.list$w)[1], 
+             HDIofMCMC(out$sims.list$k)[1], 
+             HDIofMCMC(out$sims.list$sigma)[1],
+             HDIofMCMC(out$sims.list$a)[1],
+             HDIofMCMC(out$sims.list$b)[1]),
+  hdiHigh = c(HDIofMCMC(out$sims.list$w)[2], 
+              HDIofMCMC(out$sims.list$k)[2], 
+              HDIofMCMC(out$sims.list$sigma)[2],
+              HDIofMCMC(out$sims.list$a)[2],
+              HDIofMCMC(out$sims.list$b)[2]))
+
+show(xtable(t(statsSum)))
+  
+
+
+
 
 entropy = function(p){
   - sum(p*log(p))

### 2016-01-31 text/01_CogSci_abstract/paper.tex
edits plotting for CogSci paper

@@
 aggregate population level well. To have a more fine-grained measure of model fit, we also
-looked at posterior predictive $p$-values at the level of participants and items. \ndg{need to say what this means -- i'm not even sure that i know!} Unfortunately,
-only the slider rating task provides enough data for such a fine-grained analysis. For the
-slider rating task, fixing a participant and an item, observations and replicates are probability
+looked at posterior predictive $p$-values \citep{GelmanCarlin2014:Bayesian-Data-A} at the level
+of participants and items. We focus on the binned histograms task, because here the
+population-level posterior predictive checks revealed no systematic deviance. For the slider
+rating task, fixing a participant and an item, observations and replicates are probability
 vectors of length 15. In a first analysis, we used the mean of these probability vectors as a
@@
 the trained model. In a second analysis, we used entropy as a test statistic. Two cases gave
-posterior predictive $p$-values lower than 0.05. These were from the two participants who gave a
-very extreme slider rating for the ``marbles'' item, basically assigning all ``mass'' to the
+posterior predictive $p$-values lower than 0.05. These were from the two participants who gave
+a very extreme slider rating for the ``marbles'' item, basically assigning all ``mass'' to the
 last bin. What this suggests is that the model can cope reasonably well also with
@@
 population-level hyper-prior on $P_{ij}$ will lead to shrinkage.
-\ndg{can say that one possibility (future work) would be to switch to a fatter-tailed distribution relating Q to P, in order to deal with outliers.}
 
@@
 
-\begin{itemize}
-\item \ndg{say more clearly that we focus on unidimensional distributions and why.}
-\end{itemize}
+There aspects of the data that the model does not capture well, but there are also natural
+explanations for these discrepancies. It therefore does not seem implausible that subjects'
+latent beliefs could have generated the data from all three task types roughly in the way that
+the model assumes, with each subjective belief being an expression of a population-level
+central tendency. If that is so, then mean slider ratings from the binned histogram task would
+seem to be a practical and reliable approximation of what the crowd believes.
+
+The data and model presented here suggest that mean slider ratings are consistent with other
+measures of subjective belief, namely from give-a-number and paired comparison tasks. Future
+research should evaluate whether this holds for other possible measures of subjective beliefs
+as well, such as from inferring parameterized distributions, or the use of scoring rules or the
+iterated learning paradigm.
+
+It also remains to be checked if and how our results can be extended to other types of
+uncertain variables. Here, we focused on uncertainty about unidimensional contiguous variables,
+because that was what many previous Bayesian models needed for checking their predictions. It
+remains to be seen whether the binned histogram task can be applied to higher-dimensional
+variables. Also, although it is suggestive that our results extend to cases where uncertainty
+about a binary event is measured with a single (similarly labelled slider), this should be
+checked carefully.
 

### 2016-01-31 prag_oldDev/plotting.R

@@ -205,12 +205,12 @@ plotChoicesData = ggplot(choice_ppc, aes(x = factor(bin), y = yemp)) + geom_bar(
   facet_wrap(~ item, scale = "free") +
   geom_errorbar(aes(ymin = cilow, ymax = cihigh), color = "darkgray", width = .5) +
   xlab("condition") + ylab("choce prop. higher bin") + 
-  scale_x_discrete(labels=c("1" = "1 vs 2", 
-                            "2" = "2 vs 6",
-                            "3" = "6 vs 11",
-                            "4" = "11 vs 14",
-                            "5" = "14 vs 15"
-  )) + theme(axis.text.x = element_text( angle=-25))
+  scale_x_discrete(labels=c("1" = "1 & 2", 
+                            "2" = "2 & 6",
+                            "3" = "6 & 11",
+                            "4" = "11 & 14",
+                            "5" = "14 & 15"
+  )) + theme(axis.text.x = element_text( angle=-30))
 plotChoicesDataPaper = ggplot(choice_ppc, aes(x = factor(bin), y = yemp)) + geom_bar(stat = "identity", fill = "lightgray") +
   facet_wrap(~ item, scale = "free_y", ncol = 2) +
   geom_errorbar(aes(ymin = cilow, ymax = cihigh), color = "darkgray", width = .5) +

### 2016-01-31 prag_oldDev/plotting.R
edits plotting for CogSci paper

@@ -200,7 +200,7 @@ plotChoicesPPCPaper = ggplot(choice_ppc, aes(x = bin, y = mean)) + geom_point() 
                             "3" = "6&11",
                             "4" = "11&14",
                             "5" = "14&15"
-  )) + theme(axis.text.x = element_text( angle=-25))
+  )) + theme(axis.text.x = element_text( angle=-35))
 plotChoicesData = ggplot(choice_ppc, aes(x = factor(bin), y = yemp)) + geom_bar(stat = "identity", fill = "lightgray") +
   facet_wrap(~ item, scale = "free") +
   geom_errorbar(aes(ymin = cilow, ymax = cihigh), color = "darkgray", width = .5) +

### 2016-01-31 text/01_CogSci_abstract/paper.tex
minor paper updates

@@
 \begin{abstract}
-  Experimental participants' beliefs about everyday events are of theoretical interest on their
-  own and also an important ingredient for model building---especially in Bayesian models of
+  Experimental participants' beliefs about everyday events are both of theoretical interest in their own right and an important ingredient in model building---especially in cognitive Bayesian models of
   phenomena such as logical reasoning, future predictions, and language use. Here, we explore
@@
   contiguous properties, such as the likely price of a new watch. We use a hierarchical
-  Bayesian model for inferring likely ``population-level beliefs'' as the central tendency of
+  Bayesian model for inferring likely population-level beliefs as the central tendency of
   participants' individual-level beliefs.  Three different dependent measures are related to
-  individual-level beliefs through likelihood ``linking functions:'' (i) judgements of
+  individual-level beliefs through likelihood linking functions: (i) judgements of
   (relative) likelihood of intervals of values via continuous sliders, (ii) a give-a-number
@@
 tasks. The simplest version would be this: being told that John just bought a new watch,
-subjects are asked for a single numerical estimate of its price. This task is easy to
+participants are asked for a single numerical estimate of its price. This task is easy to
 comprehend and implement, but a single number does not provide much information about the
-subjective belief that it is (arguably) a manifestation of. One solution is to infer which
+subjective belief that it is a manifestation of. One solution is to infer which
 parameterized distribution best explains the observed number choices, either at the individual
@@
 put, scoring rules are schemes of rewarding participants' answers that make sure that responses
-(by rational subjects) faithfully reflect a measure of interest: the actual subjective
+(by rational participants) faithfully reflect a measure of interest: the actual subjective
 probability of an event, the mean of a distribution, a credible interval etc. Scoring-rule
@@
 
+\begin{table*}
+  \footnotesize
+  \centering
+  \begin{tabular}{l p{2.2cm} p{2.5cm} p{3.5cm} p{3cm} p{2.5cm}}
+	\toprule
+	\textbf{Item} & \textbf{Bins (units, min, max, step)} & \textbf{Context sentence} & \textbf{GAN question} & \textbf{BH frame} & \textbf{PC frame} \\
+	\midrule
+    coffee & degrees, $<$44, $>200$, 12& X has just fetched himself a cup of coffee from the office vending machine.
+& What do you think the temperature of his coffee is?
+& His coffee was the following temperatures & The temperature of his coffee is N degrees.\\
+    	\midrule
+	commute & minutes, 0, $>$98, 7 & X commuted to work yesterday.
+& How many minutes do you think she spent commuting yesterday?
+& She commuted for the following numbers of minutes yesterday & She spent N minutes commuting.\\
+		\midrule
+	joke & children, 0, 14, 1  & X told a joke to 14 kids. & How many of the kids do you think laughed? & the following number of kids laughed & N of the children laughed.\\
+    	\midrule
+	laptop & dollars, 0, $>$7500, 500& X bought a laptop. & How much do you think it cost? & The laptop cost the following numbers of dollars & The laptop cost \$N.\\
+	\midrule
+    marbles & marbles, 0, 14, 1 & X threw 14 marbles into a pool. & How many of the marbles do you think sank? & The following number of marbles sank & N of the marbles sank.\\
+	\midrule
+	movies & minutes, 0, $>210$, 16& X just went to the movies to see a blockbuster.
+& How many minutes long do you think the movie was?
+& The movie was the following numbers of minutes long & The movie was N minutes long.\\
+	\midrule
+	TV & hours, 0, $>$43, 3& X watched TV last week.
+& How many hours do you think he spent watching TV last week?
+& He watched TV for the following numbers of hours last week & He spent N hours watching TV.\\
+	\midrule
+	watch & dollars, 0, $>$750, 50 & X bought a watch.& How much do you think it cost? & It cost the following numbers of dollars & The watch cost \$N.\\
+    \bottomrule
+  \end{tabular}
+  \caption{Experimental items. \emph{X} was a randomly generated name (different on each trial). \emph{N} was one of the bins.}
+  \label{tab:Items}
+\end{table*}
+
 In sum, there is tradeoff between simplicity of paradigms and their informativity. Ideally, we
-would like to have an experimental method for measuring subjective beliefs that (i) gives us
+would like to have an experimental method for measuring subjective beliefs that (i) provides
 sufficient and reliable-enough information about subjective beliefs to derive testable
 predictions from cognitive models that rely on such information, (ii) is easy to understand by
-the participants, (iii) is easy to implement by researchers whose focus and expertise is not
+ participants, (iii) is easy to implement by researchers whose focus and expertise is not
 belief-elicitation as such, and that (iv) does not rely on sophisticated methods of extracting
@@
 flexible enough to allow inferred subjective belief distributions beyond standard parameterized
-distributions and (vi) would provide us with a reasonable approximation of a ``population-level
-belief,'' i.e., an estimate of the central tendency or average of subjective beliefs in a given
-population. The latter allows to use one set of subjects for measuring prior beliefs and
+distributions and (vi) would provide us with a reasonable approximation of a population-level
+belief, i.e., an estimate of the central tendency or average of subjective beliefs in a given
+population. The latter would warrant using one set of participants for measuring prior beliefs and
 another for whatever other task is of interest.
 
-A simple technique that seems to fit this bill is the \emph{binned histograms} task that has
+A simple technique that seems to fit this bill is the \emph{binned histogram} task that has
 recently been used with apparent success
@@
 values (established by a pre-test). E.g., for the price of a watch bins could be ``\$0-\$50'',
-``\$50-\$100'', etc. up to ``\$700-\$750'' and ``more than \$750''. Each subject's slider
-ratings are normalized within each item, and the results are averaged across subjects to obtain
+``\$50-\$100'', etc. up to ``\$700-\$750'' and ``more than \$750''. Each participant's slider
+ratings are normalized within each item, and the results are averaged across participants to obtain
 an estimate of the population's average belief about each item. These averaged normalized
-slider ratings per bin (henceforth: mean slider ratings) look intuitively plausible and give
+slider ratings per bin (henceforth: mean slider ratings) look intuitively plausible and yield
 good results if fed into cognitive models that predict behavior in language or reasoning tasks,
@@
 Explanatory success notwithstanding, the question remains whether mean slider ratings measure
-what we would like them to. Are these good approximations of a central tendency of subjects'
-individual beliefs? Are these measures consistent with subjects' behavior in other tasks, such
+what we would like them to. Are these good approximations of a central tendency of participants'
+individual beliefs? Are these measures consistent with participants' behavior in other tasks, such
 as \emph{give-a-number}, as well? To address this issue we collected data in a within-subject
@@
   posteriori} population-level beliefs are approximated well by mean slider ratings, suggesting
-that the latter may indeed be a plausible, yet cheap measure of ``what the crowd beliefs.''
-Moreover, the model is able to capture subjects' behavior in all three task types reasonably
-well, suggesting that what mean sliders measure is what we think it is: a population-level
+that the latter may indeed be a plausible, yet cheap measure of what the crowd believes.
+Moreover, the model is able to capture participants' behavior in all three task types reasonably
+well, suggesting that what mean slider ratings measure is what we think it is: a population-level
 central tendency of latent subjective beliefs.
 
-\section{Experiment}
+\section{Experimental elicitation of prior beliefs}
 
@@
 
-\paragraph{Participants} We recruited 20 self-reported native speakers of English over Mechanical Turk.
-
-\paragraph{Procedure and materials}
+\paragraph{Participants, procedure and materials} We recruited 20 self-reported English native speakers  over Mechanical Turk.
 
@@
 dependent measure. Trial order was randomized.
-
-\begin{table*}
-  \footnotesize
-  \centering
-  \begin{tabular}{l p{2.2cm} p{2.5cm} p{3.5cm} p{3cm} p{2.5cm}}
-	\toprule
-	Item & Bins (units, min, max, step) & Context sentence & GAN question & BH frame & PC frame \\
-	\midrule
-    coffee & degrees, $<$44, $>200$, 12& X has just fetched himself a cup of coffee from the office vending machine.
-& What do you think the temperature of his coffee is?
-& His coffee was the following temperatures & The temperature of his coffee is N degrees.\\
-	commute & minutes, 0, $>$98, 7 & X commuted to work yesterday.
-& How many minutes do you think she spent commuting yesterday?
-& She commuted for the following numbers of minutes yesterday & She spent N minutes commuting.\\
-	joke & children, 0, 14, 1  & X told a joke to 14 kids. & How many of the kids do you think laughed? & the following number of kids laughed & N of the children laughed.\\
-
-	laptop & dollars, 0, $>$7500, 500& X bought a laptop. & How much do you think it cost? & The laptop cost the following numbers of dollars & The laptop cost \$N.\\
-    marbles & marbles, 0, 14, 1 & X threw 14 marbles into a pool. & How many of the marbles do you think sank? & The following number of marbles sank & N of the marbles sank.\\
-
-	movies & minutes, 0, $>210$, 16& X just went to the movies to see a blockbuster.
-& How many minutes long do you think the movie was?
-& The movie was the following numbers of minutes long & The movie was N minutes long.\\
-
-	TV & hours, 0, $>$43, 3& X watched TV last week.
-& How many hours do you think he spent watching TV last week?
-& He watched TV for the following numbers of hours last week & He spent N hours watching TV.\\
-
-	watch & dollars, 0, $>$750, 50 & X bought a watch.& How much do you think it cost? & It cost the following numbers of dollars & The watch cost \$N.\\
-    \bottomrule
-  \end{tabular}
-  \caption{Experimental items. \emph{X} was a randomly generated name (different on each trial). \emph{N} was one of the bins.}
-  \label{tab:Items}
-\end{table*}
 
@@
 
-Results are shown in \figref{fig:Results}. \figref{fig:Results} shows mean slider ratings in
-relation to frequencies of number choices in the \emph{give-a-number} task. It looks like the
+\figref{fig:Results} shows mean slider ratings alongside the frequencies of number choices in the \emph{give-a-number} task. It looks like the
 latter could be samples from the former, with a tendency to modal choices. But, despite the use
-of slider, some \emph{give-a-number} results still seem to reflect a tendency towards round or
+of sliders, some \emph{give-a-number} results still seem to reflect a tendency towards round or
 salient numbers. E.g., in item \emph{coffee} more than 40\% of participants guessed that a
@@
 vs 11. A potential explanation for this, which we will implement formally in the
-data-generating model, is that the \emph{paired comparison} condition invited subjects to think
+data-generating model, is that the \emph{paired comparison} condition invited participants to think
 about what they thought was mostly likely the case (usually: bin 15 in the
@@
 
-\begin{figure}
+\begin{figure}[t!]
   \centering
-  \begin{subfigure}[b]{0.5\textwidth}
+  \begin{subfigure}{0.5\textwidth}
     \includegraphics[width = \textwidth]{plots/data_sliderNumber.pdf}
@@
 
-  \begin{subfigure}[b]{0.5\textwidth}
+  \begin{subfigure}{0.5\textwidth}
     \includegraphics[width = \textwidth]{plots/data_choice.pdf}
@@
 
-\begin{figure}[]
-  \centering
-  \includegraphics[width = 0.4\textwidth]{modelGraph/modelGraphNoMath.pdf}
-  \caption{The data-generating model as a probabilistic graphical model, following conventions
-    of \citet{LeeWagenmakers2013:Bayesian-Cognit}. Shaded nodes are observed, white nodes are
-    latent variables. Square nodes represent categorical, round nodes continuous
-    variables. Boxes indicate scope of indices.}
-  \label{fig:ModelGraph}
-\end{figure}
-
-To fill the structure in Figure~\ref{fig:ModelGraph} with life, we need to spell out three
+To fill the structure in \figref{fig:ModelGraph} with life, we need to spell out three
 parameterized link functions, one for each task type, and the relation between population-level
@@
 natural approach would be to link the likelihood of choosing $b_h$ over $b_l$ to the difference
-between $P_{ijb_h}$ and $P_{ijb_l}$. However, as discussed above, This does not appear to be
+between $P_{ijb_h}$ and $P_{ijb_l}$. However, as discussed above, this does not appear to be
 what participants were doing. Indeed, a model that implements this idea failed to capture the
@@
 
+\begin{figure}[]
+  \centering
+  \includegraphics[width = 0.4\textwidth]{modelGraph/modelGraphNoMath.pdf}
+  \caption{The data-generating model as a probabilistic graphical model, following conventions
+    of \citet{LeeWagenmakers2013:Bayesian-Cognit}. Shaded nodes are observed, white nodes are
+    latent variables. Square nodes represent categorical, round nodes continuous
+    variables. Boxes indicate scope of indices.}
+  \label{fig:ModelGraph}
+\end{figure}
+
 \section{Inference}
@@
 Table~\ref{tab:SummaryStats}. Posterior credible levels of $w$ allow for a limited amount of
-slack around $Q_j$. Values for $\kappa$ indicate that, on average, subjects may not have had a
+slack around $Q_j$. Values for $\kappa$ indicate that, on average, participants may not have had a
 preference for or against extreme slider ratings. Relatively high values of $a$ indicate that
-subjects, on average, had a strong tendency to choose their mode in the \emph{give-a-number}
+participants, on average, had a strong tendency to choose their mode in the \emph{give-a-number}
 task.
@@
 
-Of most theoretical interest are posterior estimates of $Q_j$. Figure~\ref{fig:PosteriorQj}
+Of most theoretical interest are posterior estimates of $Q_j$. \figref{fig:PosteriorQj}
 shows means of the posterior $Q_j$, with their 95\% HDIs, alongside the averaged normalized
@@
 Inferences based on the model are only as reliable as the model itself is plausible. Model
-criticism is therefore important. Figure~\ref{fig:PPCs} shows posterior predictive checks at
+criticism is therefore important. \figref{fig:PPCs} shows posterior predictive checks at
 the population-aggregate level for all of our three task types. For the binned histogram task,
@@
 
-\section{Conclusions}
-
-There aspects of the data that the model does not capture well, but there are also natural
-explanations for these discrepancies. It therefore does not seem implausible that subjects'
-latent beliefs could have generated the data from all three task types roughly in the way that
-the model assumes, with each subjective belief being an expression of a population-level
-central tendency. If that is so, then mean slider ratings from the binned histogram task would
-seem to be a practical and reliable approximation of what the crowd believes.
+\section{Conclusion}
 
@@
 research should evaluate whether this holds for other possible measures of subjective beliefs
-as well, such as from inferring parameterized distributions, or the use of scoring rules or the
+as well, such as from inferring parameterized distributions, the use of scoring rules, or the
 iterated learning paradigm.
 
-It also remains to be checked if and how our results can be extended to other types of
-uncertain variables. Here, we focused on uncertainty about unidimensional contiguous variables,
-because that was what many previous Bayesian models needed for checking their predictions. It
+There are aspects of the data that the model does not capture well, but there are also natural
+explanations for these discrepancies. It therefore does not seem implausible that participants'
+latent beliefs could have generated the data from all three task types roughly in the way assumed by the model, with each subjective belief being an expression of a population-level
+central tendency. If that is so, then mean slider ratings from the binned histogram task are a practical and reliable approximation of what the crowd believes.
+
+Future research should investigate whether and how our results can be extended to other types of
+uncertain variables. Here, guided by the needs of many previous Bayesian cognitive models, we focused on uncertainty about unidimensional contiguous variables. It
 remains to be seen whether the binned histogram task can be applied to higher-dimensional
-variables. Also, although it is suggestive that our results extend to cases where uncertainty
+variables \jd{like what?}. Also, although it is suggestive that our results extend to cases where uncertainty
 about a binary event is measured with a single (similarly labelled slider), this should be
-checked carefully.
+checked carefully. \jd{what does this mean?}
 

### 2016-01-31 text/01_CogSci_abstract/paper.tex
edits to motivation

@@
 \newcommand{\mf}[1]{\textcolor{Red}{[mf: #1]}} 
+\definecolor{DarkOrange}{RGB}{255,100,50}
+\newcommand{\mht}[1]{\textcolor{DarkOrange}{[mht: #1]}}
 
@@
 explanations of higher-order cognition: decision making, planning, reasoning, and language
-use. It is therefore vital to have methods for validating explanatory mental states
+use. It is therefore vital to have methods for validating explanatory mental state
 ascriptions.
@@
 
-In sum, there is tradeoff between simplicity of paradigms and their informativity. Ideally, we
+In sum, there is tradeoff between simplicity of paradigms and their information content. Ideally, we
 would like to have an experimental method for measuring subjective beliefs that (i) provides
@@
 \citep{KaoWu2014:Nonliteral-Unde,SchollerFranke2015:Semantic-values}. \mf{someone please add
-  references: where did you use binned sliders?} Participants adjust sliders in order to
+  references: where did you use binned sliders?} \mht{I used ``categorical'' sliders for discrete priors in my syllogistic reasoning / belief bias stuff; and so did justine in her metaphor work...}Participants adjust sliders in order to
 express their (relative) subjective beliefs about how likely it is that the value of an
@@
 individual beliefs? Are these measures consistent with participants' behavior in other tasks, such
-as \emph{give-a-number}, as well? To address this issue we collected data in a within-subject
+as \emph{give-a-number}, as well? To address this issue we collected data in a within-subject paradigm 
 from three task types: (i) \emph{binned histograms}, (ii) \emph{give-a-number} and (iii)

### 2016-01-31 text/01_CogSci_abstract/paper.tex
make bin column of table more readable

@@
 	\toprule
-	\textbf{Item} & \textbf{Bins (units, min, max, step)} & \textbf{Context sentence} & \textbf{GAN question} & \textbf{BH frame} & \textbf{PC frame} \\
+	\textbf{Item} & \textbf{Bins} (\{min, max\} step; units)& \textbf{Context sentence} & \textbf{GAN question} & \textbf{BH frame} & \textbf{PC frame} \\
 	\midrule
-    coffee & degrees, $<$44, $>200$, 12& X has just fetched himself a cup of coffee from the office vending machine.
+    coffee &  \{$<$44, $>$200\} 2; degrees  & X has just fetched himself a cup of coffee from the office vending machine.
 & What do you think the temperature of his coffee is?
@@
     	\midrule
-	commute & minutes, 0, $>$98, 7 & X commuted to work yesterday.
+	commute &  \{0, $>$98\} 7; minutes & X commuted to work yesterday.
 & How many minutes do you think she spent commuting yesterday?
@@
 		\midrule
-	joke & children, 0, 14, 1  & X told a joke to 14 kids. & How many of the kids do you think laughed? & the following number of kids laughed & N of the children laughed.\\
+	joke &  \{0, 14\} 1; children  & X told a joke to 14 kids. & How many of the kids do you think laughed? & the following number of kids laughed & N of the children laughed.\\
     	\midrule
-	laptop & dollars, 0, $>$7500, 500& X bought a laptop. & How much do you think it cost? & The laptop cost the following numbers of dollars & The laptop cost \$N.\\
+	laptop &  \{0, $>$7500\} 500; dollars& X bought a laptop. & How much do you think it cost? & The laptop cost the following numbers of dollars & The laptop cost \$N.\\
 	\midrule
-    marbles & marbles, 0, 14, 1 & X threw 14 marbles into a pool. & How many of the marbles do you think sank? & The following number of marbles sank & N of the marbles sank.\\
+    marbles &  \{0, 14\} 1; marbles & X threw 14 marbles into a pool. & How many of the marbles do you think sank? & The following number of marbles sank & N of the marbles sank.\\
 	\midrule
-	movies & minutes, 0, $>210$, 16& X just went to the movies to see a blockbuster.
+	movies &  \{0, $>210$\} 16; minutes& X just went to the movies to see a blockbuster.
 & How many minutes long do you think the movie was?
@@
 	\midrule
-	TV & hours, 0, $>$43, 3& X watched TV last week.
+	TV &  \{0, $>$43\} 3; hours & X watched TV last week.
 & How many hours do you think he spent watching TV last week?
@@
 	\midrule
-	watch & dollars, 0, $>$750, 50 & X bought a watch.& How much do you think it cost? & It cost the following numbers of dollars & The watch cost \$N.\\
+	watch &  \{0, $>$750\} 50; dollars& X bought a watch.& How much do you think it cost? & It cost the following numbers of dollars & The watch cost \$N.\\
     \bottomrule

### 2016-02-01 text/01_CogSci_abstract/paper.tex
edits throughout methods and discussion

@@
  {\large \bf Erin Bennett, Judith Degen, Michael Henry Tessler, Noah D. Goodman}\\
-     \{ebennett,jdegen,mhtessler,ngoodman\}@stanford.edu \\
-     Department of Psychology, 450 Serra Mall \\
-  Stanford, CA 94305 USA }
+     \{ebennett, jdegen, mtessler, ngoodman\}@stanford.edu \\
+     Department of Psychology, Stanford University \\
+     450 Serra Mall, Stanford, CA 94305 USA }
 
@@
 $k \in \set{1, \dots, 15}$ from the binned histogram task; (ii) the bins $n_{ij} \in \set{1, \dots, 15}$ in which participant
-$i$'s number choice for item $i$ was in the give-a-number task; and (iii) the binary choices $c_{ijl} \in \set{0,1}$ of
+$i$'s number choice for item $j$ was in the give-a-number task; and (iii) the binary choices $c_{ijl} \in \set{0,1}$ of
 whether participant $i$ selected the higher bin for item $j$ in the paired comparison task for each comparison
@@
 idea is that $P_{ij}$ are noise-perturbed variants scattered around $Q_j$, with some parameter
-$w$ to determine how much perturbation we should expect. To realize this, the model assumes
+$w$ to determine how much perturbation we should expect.
+ To realize this, the model assumes
 that $P_{ij}$ are distributed according to a Dirichlet distribution with weights given by
@@
 
-The link function for the slider rating data uses a logit transformation to project observed
+The link function for the \textbf{slider rating data} uses a logit transformation to project observed
 slider ratings $s_{ijk}$ and latent probabilities $P_{ijk}$, which are bound to lie between 0
@@
 
-The link function for number choice data treats each bin $n_{ij}$ as a draw from a categorical
+The link function for \textbf{number choice data} treats each bin $n_{ij}$ as a draw from a categorical
 distribution where the probability of bin $k$ is proportional to $\expo(a P_{ij})$, i.e., a
@@
 
-Finally, consider the link function for bin comparisons. We are interested in the likelihood
+Finally, consider the link function for \textbf{bin comparisons}. We are interested in the likelihood
 with which participant $i$ selects the higher bin for item $j$ in bin comparison condition
@@
 between $P_{ijb_h}$ and $P_{ijb_l}$. However, as discussed above, this does not appear to be
-what participants were doing. Indeed, a model that implements this idea failed to capture the
-relevant regularities in the data blatantly. Another plausible link function is to assume that
+what participants were doing. Indeed, a model that implements this idea blatantly failed to capture the
+relevant regularities in the data. Another plausible link function is to assume that
 what matters in bin comparisons is the distance to the mode of $P_{ij}$: soft-max prefer the
@@
     p^{\text{high}}) + \exp(b p^{\text{low}})}$
-  with $p^\text{low}_{ijl} = 2 - p^\text{high}_{ijl}$.}
+  with $p^\text{low}_{ijl} = 2 - p^\text{high}_{ijl}$.
+  \mht{I am unable to notice this. Double check? Intuition tells me $p^\text{low}_{ijl} = 1 - p^\text{high}_{ijl}$.}
+  }
 \begin{align*}
@@
 population-level beliefs. Inspection of posteriors of individual $P_{ij}$ shows that there is
-ample variation between participants. Nonetheless, the way the model suggest we should think about
+ample variation between participants. Nonetheless, the way the model suggests we should think about
 harnessing the individual $P_{ij}$s under a population-level central tendency is closely
@@
 the population-aggregate level for all of our three task types. For the binned histogram task,
-posterior predictions are spot-on. Some of the give-a-number data is surprising even for
-the model trained on this very data. This could have various reasons: (i) the give-a-number data does
+posterior predictions are spot-on. For the give-a-number task, some of the data is surprising despite 
+the model being trained on this very data. This could have various reasons: (i) the give-a-number data does
 not have a huge influence on the posterior likelihood, (ii) number choices may be influenced by
-saliency and/or roundness of numbers after all. Finally, there is one condition in the paired
+saliency and/or roundness of numbers after all (and thus, not accurately reflect the true beliefs $Q_j$). Finally, there is one condition in the paired
 comparison task that the model definitely got wrong. This is the choice of what is more likely:
-that one or that none of 14 marbles thrown into a pool would sink. The model would predict that
+One or none of 14 marbles thrown into a pool would sink. The model would predict that
 almost everybody should answer that it is more likely that one marble sank. But that is not
@@
 population-level hyper-prior on $P_{ij}$ will lead to shrinkage.
-
+\mht{Is it worth mentioning about the possilbiity of including subject-specific link function parameters e.g., $\kappa$ and $\alpha$?}
 \section{Conclusion}

### 2016-02-01 text/01_CogSci_abstract/paper.tex
CogSci abstract final edits

@@
 \usepackage{booktabs}
+\usepackage{caption}
+\captionsetup{belowskip=0pt,aboveskip=2pt}
+\setlength{\textfloatsep}{9pt}
 
@@
 
-\author{{\large \bf Michael Franke} \\
-  michael.franke@uni-tuebingen.de \\
+\author{{\large \bf Michael Franke, Fabian Dablander, Anthea Sch\"{o}ller} \\
+  \{michael.franke, fabian.dablander, anthea.schoeller\}@uni-tuebingen.de \\
   Department of Linguistics, Wilhelmstra\ss e 19 \\
   72074 T\"{u}bingen, Germany \AND
- {\large \bf Erin Bennett, Judith Degen, Michael Henry Tessler, Noah D. Goodman}\\
-     \{ebennett, jdegen, mtessler, ngoodman\}@stanford.edu \\
+  {\large \bf Erin Bennett, Judith Degen, Michael Henry Tessler, Justine Kao, Ciyang Qing, Noah D. Goodman}\\
+  \{ebennett, jdegen, mtessler, justinek, qciyang, ngoodman\}@stanford.edu \\
   Department of Psychology, Stanford University \\
@@
 \begin{abstract}
-  Experimental participants' beliefs about everyday events are both of theoretical interest in their own right and an important ingredient in model building---especially in cognitive Bayesian models of
-  phenomena such as logical reasoning, future predictions, and language use. Here, we explore
-  several recently used methods for measuring subjective beliefs about unidimensional
-  contiguous properties, such as the likely price of a new watch. We use a hierarchical
-  Bayesian model for inferring likely population-level beliefs as the central tendency of
-  participants' individual-level beliefs.  Three different dependent measures are related to
-  individual-level beliefs through likelihood linking functions: (i) judgements of
-  (relative) likelihood of intervals of values via continuous sliders, (ii) a give-a-number
-  task, and (iii) choice of the more likely of two intervals of values. Our results suggest
-  that using averaged normalized slider ratings for binned quantities is a practical and fairly
-  good approximator of inferred population-level beliefs.
+  Experimental participants' beliefs about everyday events are both of theoretical interest in
+  their own right and an important ingredient in model building---especially in cognitive
+  Bayesian models of phenomena such as logical reasoning, future predictions, and language
+  use. Here, we explore several recently used methods for measuring subjective beliefs about
+  unidimensional contiguous properties, such as the likely price of a new watch. We use a
+  hierarchical Bayesian model for inferring likely population-level beliefs as the central
+  tendency of participants' individual-level beliefs.  Three different dependent measures are
+  used to infer latent beliefs: (i) slider ratings of (relative) likelihood of intervals of
+  values, (ii) a give-a-number task, and (iii) choice of the more likely of two intervals of
+  values. Our results suggest that using averaged normalized slider ratings for binned
+  quantities is a practical and fairly good approximator of inferred population-level beliefs.
 
@@
 also in science. Scientific ascription of latent mental states plays an important role in many
-explanations of higher-order cognition: decision making, planning, reasoning, and language
-use. It is therefore vital to have methods for validating explanatory mental state
-ascriptions.
+explanations of higher-order cognition: decision making, reasoning, language use etc. It is
+therefore vital to have methods for validating explanatory mental state ascriptions.
 
@@
 understood to convey speaker affect, rather than literal meaning, can be explained well by a
-Bayesian model of utterance interpretation \citep{KaoWu2014:Nonliteral-Unde}. In this model a
-crucial role is played by an empirical measure of participants' expectations about the likely
+Bayesian model of utterance interpretation \citep{KaoWu2014:Nonliteral-Unde}. This model
+assigns a crucial role to an empirical measure of participants' expectations about the likely
 or normal price of a watch.  Other examples of domains in which empirically successful models
@@
 \citep{FrankGoodman2012:Predicting-Prag}, the strength of pragmatic enrichments
-\citep{DegenTessler2015:Wonky-worlds:-L}, or the interpretation of vague quantifiers
-\citep{SchollerFranke2015:Semantic-values}. \mf{anyone wants any other references here?}
+\citep{DegenTessler2015:Wonky-worlds:-L}, or the interpretation of quantifiers
+\citep{SchollerFranke2015:Semantic-values}.
 
-Many methods of assessing subjective beliefs exist. One approach is to take actual frequencies
-as an approximation to subjective beliefs
-\citep{GriffithsTenenbaum2006:Optimal-Predict}. Unfortunately, the relevant data may not be
-available (e.g., for one-shot or imaginary events) and real-world frequencies may deviate from
-participants' subjective beliefs for many reasons.
+One method of getting at subjective beliefs is to take actual frequencies as an approximation
+to subjective beliefs \cite<e.g.>{GriffithsTenenbaum2006:Optimal-Predict}. Unfortunately,
+frequency data may be unavailable (e.g., one-shot events) or deviate from participants'
+subjective beliefs in crucial respects.
 
-Another possibility is to empirically measure subjective beliefs by \emph{give-a-number}
+Another frequent approach is to empirically measure subjective beliefs by \emph{give-a-number}
 tasks. The simplest version would be this: being told that John just bought a new watch,
@@
 comprehend and implement, but a single number does not provide much information about the
-subjective belief that it is a manifestation of. One solution is to infer which
-parameterized distribution best explains the observed number choices, either at the individual
-level \citep{Manski2004:Measuring-Expec} or on the population level
+subjective belief that it is a manifestation of. One solution is to infer which parameterized
+distribution best explains the observed number choices, either at the individual level
+\citep{Manski2004:Measuring-Expec} or at the population level
 \citep{TauberSteyvers2013:Inferring-Subje}.
 
-More sophisticated \emph{give-a-number} tasks give more information, but can be rather complex
-and more difficult to implement. A prominent method in economics uses \emph{scoring rules}
+More sophisticated \emph{give-a-number} tasks give more information, but can be difficult to
+implement and analyze. A prominent method in economics uses \emph{scoring rules}
 \citep{Savage1971:Elicitation-of-,AndersenFountain2014:Estimating-Subj,SchlagTremewan2014:A-penny-for-you}. Roughly
@@
 		\midrule
-	joke &  \{0, 14\} 1; children  & X told a joke to 14 kids. & How many of the kids do you think laughed? & the following number of kids laughed & N of the children laughed.\\
+	joke &  \{0, 14\} 1; children  & X told a joke to 14 kids. & How many of the kids do you think laughed? & The following number of kids laughed & N of the children laughed.\\
     	\midrule
@@
 
-In sum, there is tradeoff between simplicity of paradigms and their information content. Ideally, we
-would like to have an experimental method for measuring subjective beliefs that (i) provides
-sufficient and reliable-enough information about subjective beliefs to derive testable
-predictions from cognitive models that rely on such information, (ii) is easy to understand by
- participants, (iii) is easy to implement by researchers whose focus and expertise is not
-belief-elicitation as such, and that (iv) does not rely on sophisticated methods of extracting
-the relevant information from the observed data. Moreover, the ideal method would be (v)
-flexible enough to allow inferred subjective belief distributions beyond standard parameterized
-distributions and (vi) would provide us with a reasonable approximation of a population-level
-belief, i.e., an estimate of the central tendency or average of subjective beliefs in a given
-population. The latter would warrant using one set of participants for measuring prior beliefs and
-another for whatever other task is of interest.
+In sum, there is tradeoff between simplicity of paradigms and their information
+content. Ideally, we would like to have an experimental method for measuring subjective beliefs
+that (i) provides sufficient and reliable-enough information about subjective beliefs to derive
+testable predictions from cognitive models that rely on such information, (ii) is easy to
+understand by participants, (iii) is easy to implement, and that (iv) does not require
+sophisticated means of data analysis. Moreover, the ideal method would be (v) flexible enough
+to allow inferred subjective belief distributions beyond standard parameterized distributions
+and (vi) would provide a good approximation of the central tendency or average of subjective
+beliefs in a given population. The latter would allow using one set of participants for
+measuring prior beliefs and another for whatever other task is of interest, so as to avoid
+potential cross-over effects.
 
@@
 recently been used with apparent success
-\citep{KaoWu2014:Nonliteral-Unde,SchollerFranke2015:Semantic-values}. \mf{someone please add
-  references: where did you use binned sliders?} \mht{I used ``categorical'' sliders for discrete priors in my syllogistic reasoning / belief bias stuff; and so did justine in her metaphor work...}Participants adjust sliders in order to
-express their (relative) subjective beliefs about how likely it is that the value of an
-uncertain contiguous quantity lies in some interval of possible values (a bin). E.g.,
-participants read that John bought a watch.  To report their beliefs about the watch's price,
-they adjust sliders whose endpoints are labelled as ``extremely likely'' and ``extremely
-unlikely.'' Each slider corresponds to one out of 15 bins that partition the range of plausible
-values (established by a pre-test). E.g., for the price of a watch bins could be ``\$0-\$50'',
-``\$50-\$100'', etc. up to ``\$700-\$750'' and ``more than \$750''. Each participant's slider
-ratings are normalized within each item, and the results are averaged across participants to obtain
-an estimate of the population's average belief about each item. These averaged normalized
-slider ratings per bin (henceforth: mean slider ratings) look intuitively plausible and yield
-good results if fed into cognitive models that predict behavior in language or reasoning tasks,
-for instance.
+\cite<e.g.>{KaoWu2014:Nonliteral-Unde,KaoBergen2014:Formalizing--th,Tessler2015:Understanding-B,SchollerFranke2015:Semantic-values}. Participants
+adjust sliders to express (relative) subjective beliefs about how likely the value of an
+uncertain contiguous quantity lies in some interval of possible values (a bin). E.g., to report
+beliefs about the a watch's price, participants adjust sliders whose endpoints are labelled
+``extremely likely'' and ``impossible.'' Each slider corresponds to one of 15 bins that
+partition the range of plausible values (established by a pre-test), such as ``\$0-\$50'',
+``\$50-\$100'', etc. up to ``\$700-\$750'' and ``more than \$750.'' Each participant's slider
+ratings are normalized and the results averaged across participants. The resulting \emph{mean
+  slider ratings} look like plausible population-level beliefs and give good results when fed
+into cognitive models that predict task behavior based on prior expectations.
 
-Explanatory success notwithstanding, the question remains whether mean slider ratings measure
+Explanatory success aside, the question remains whether mean slider ratings measure
 what we would like them to. Are these good approximations of a central tendency of participants'
@@
 
-\subsection{Methods}
-
-\paragraph{Participants, procedure and materials} We recruited 20 self-reported English native speakers  over Mechanical Turk.
-
-We elicited prior beliefs for eight different items (listed in
+\paragraph{Participants, procedure and materials} We recruited 20 self-reported English native
+speakers over Mechanical Turk and collected responses for eight different items (listed in
 \tableref{tab:Items}) using three different dependent measures: \emph{binned histograms},
@@
 On \emph{give-a-number} (GAN) trials, participants saw the context sentence and GAN question
-for each item (see \tableref{tab:Items}) and were asked to provide one number by adjusting a
+for each item (see \tableref{tab:Items}) and provided one number by adjusting a
 slider with endpoints labeled as the lowest and highest number for that item. Min and max
@@
 
-\figref{fig:Results} shows mean slider ratings alongside the frequencies of number choices in the \emph{give-a-number} task. It looks like the
-latter could be samples from the former, with a tendency to modal choices. But, despite the use
-of sliders, some \emph{give-a-number} results still seem to reflect a tendency towards round or
-salient numbers. E.g., in item \emph{coffee} more than 40\% of participants guessed that a
-coffee from the vending machine was ca. 150 degrees Fahrenheit (bin 10). The results from the
-\emph{paired comparison} seem consistent with the mean slider ratings as well, albeit not as
-straightforwardly as one might think. This is clearest from the \emph{marbles} case. While the
-mean slider ratings indicate that bins 2, 6 and 11 were all felt to be equally likely on
-average, almost every participant chose the higher bin in direct bin comparisons 2 vs 6 and 6
-vs 11. A potential explanation for this, which we will implement formally in the
-data-generating model, is that the \emph{paired comparison} condition invited participants to think
-about what they thought was mostly likely the case (usually: bin 15 in the
-\emph{marbles} case) and to choose the bin that is closest to that.
+\paragraph{Results} \figref{fig:Results} shows mean slider ratings alongside the frequencies of
+number choices in the \emph{give-a-number} task. It looks like the latter could be samples from
+the former, with a tendency to modal choices. But some \emph{give-a-number} results seem
+influenced by a tendency towards round or salient numbers. E.g., in item \emph{coffee} more
+than 40\% of participants guessed that a coffee from the vending machine was ca.~150$^{\circ}$F
+(bin 10). The results from the \emph{paired comparison} seem consistent with the mean slider
+ratings as well, albeit not as straightforwardly as one might think. This is clearest from the
+\emph{marbles} case. While the mean slider ratings suggest that bins 2, 6 and 11 were
+considered roughly equiprobable on average, almost every participant chose the higher bin in
+direct bin comparisons 2 vs 6 and 6 vs 11. A potential explanation for this, which we will
+implement formally in the data-generating model, is that the \emph{paired comparison} condition
+invited participants to think about what they thought was mostly likely the case (usually: bin
+15 in the \emph{marbles} case) and to choose the bin that is closest to that.
 
@@
   \centering
-  \begin{subfigure}{0.5\textwidth}
+  \begin{subfigure}{0.475\textwidth}
     \includegraphics[width = \textwidth]{plots/data_sliderNumber.pdf}
-    \caption{Red: proportion of choosing a number in each bin on \emph{give-a-number} trials. Black: mean slider ratings on \emph{binned histogram} trials.}
+    \caption{Red: proportion of bins corresponding to number choices on \emph{give-a-number} trials. Black: mean slider ratings on \emph{binned histogram} trials.}
     \label{fig:slider}
@@
 
-  \begin{subfigure}{0.5\textwidth}
+  \begin{subfigure}{0.475\textwidth}
     \includegraphics[width = \textwidth]{plots/data_choice.pdf}
-    \caption{Proportion of choosing the higher bin on \emph{paired comparison} trials.}
+    \caption{Proportion of higher bin choices on \emph{paired comparison} trials.}
     \label{fig:lighting}
@@
 
-  \caption{Experimental results. Error bars show bootstrapped 95\% confidence intervals.}
+  \caption{Average data. Bars are bootstrapped 95\% CIs.}
   \label{fig:Results}
@@
 $P_{ijk}$ being participant $i$'s belief about the relative likelihood of bin $k$ for item
-$j$. Each $P_{ij}$ defines a likelihood for our data, via appropriate link functions (to be
-spelled out presently). Variance in subjective beliefs is harnessed by a population-level
-hyper-prior with central tendency $Q_{j}$, i.e., a stochastic vector of length 15. The
-structure of this model is pictured in \figref{fig:ModelGraph}.
+$j$. Each $P_{ij}$ defines a likelihood for our data, via appropriate link functions. Variance
+in subjective beliefs is harnessed by a population-level hyper-prior with central tendency
+$Q_{j}$. The structure of this model is pictured in \figref{fig:ModelGraph}.
 
@@
 parameterized link functions, one for each task type, and the relation between population-level
-belief $Q_j$ to individual beliefs $P_{ij}$. Let's start with the latter. For fixed $Q_j$, the
-idea is that $P_{ij}$ are noise-perturbed variants scattered around $Q_j$, with some parameter
-$w$ to determine how much perturbation we should expect.
- To realize this, the model assumes
-that $P_{ij}$ are distributed according to a Dirichlet distribution with weights given by
-$w Q_{j}$:
+belief $Q_j$ and individual beliefs $P_{ij}$. Let's start with the latter. The idea is that
+$P_{ij}$ are noise-perturbed variants scattered around $Q_j$, with some parameter $w$ to
+determine how much perturbation we should expect.  To realize this, the model assumes that
+$P_{ij}$ are distributed according to a Dirichlet distribution with weights given by $w Q_{j}$:
 \begin{align*}
@@
 
-Finally, consider the link function for \textbf{bin comparisons}. We are interested in the likelihood
-with which participant $i$ selects the higher bin for item $j$ in bin comparison condition
-$l$. Suppose $l$ is about comparing the lower bin $b_l$ to the higher $b_h$. The perhaps most
-natural approach would be to link the likelihood of choosing $b_h$ over $b_l$ to the difference
-between $P_{ijb_h}$ and $P_{ijb_l}$. However, as discussed above, this does not appear to be
-what participants were doing. Indeed, a model that implements this idea blatantly failed to capture the
-relevant regularities in the data. Another plausible link function is to assume that
-what matters in bin comparisons is the distance to the mode of $P_{ij}$: soft-max prefer the
-bin that is closer to the mode of $P_{ij}$; select randomly if both bins are equally far from
-the prototype.\footnote{Notice that
-  $(1 + \exp(2b(1-p^{\text{high}}_{ijl})) )^{-1} = \frac{\exp(b p^{\text{high}})}{\exp(b
-    p^{\text{high}}) + \exp(b p^{\text{low}})}$
-  with $p^\text{low}_{ijl} = 2 - p^\text{high}_{ijl}$.
-  \mht{I am unable to notice this. Double check? Intuition tells me $p^\text{low}_{ijl} = 1 - p^\text{high}_{ijl}$.}
-  }
+Finally, consider the link function for \textbf{bin comparisons}. We are interested in the
+likelihood with which participant $i$ selects the higher bin for item $j$ in bin comparison
+condition $l$. Suppose $l$ is about comparing the lower bin $b_l$ to the higher $b_h$. The
+perhaps most natural approach would be to link the likelihood of choosing $b_h$ over $b_l$ to
+the difference between $P_{ijb_h}$ and $P_{ijb_l}$. However, as discussed above, this does not
+appear to be what participants were doing. Indeed, a model that implements this idea blatantly
+fails to capture the relevant regularities in the data. Another plausible link function is to
+assume that what matters is the distance to the mode of $P_{ij}$: soft-max prefer the bin that
+is closer to the mode of $P_{ij}$; select randomly if both bins are equally far from the
+prototype.\footnote{$(1 + \exp(2b(1-x)) )^{-1} = \frac{\exp(b x)}{\exp(b
+    x) + \exp(b y)}$
+  if $x = 2 - y$.}
 \begin{align*}
@@
   \centering
-  \includegraphics[width = 0.4\textwidth]{modelGraph/modelGraphNoMath.pdf}
+  \includegraphics[width = 0.35\textwidth]{modelGraph/modelGraphNoMath.pdf}
   \caption{The data-generating model as a probabilistic graphical model, following conventions
@@
 Table~\ref{tab:SummaryStats}. Posterior credible levels of $w$ allow for a limited amount of
-slack around $Q_j$. Values for $\kappa$ indicate that, on average, participants may not have had a
+slack around $Q_j$. Values for $\kappa$ indicate that, on average, participants had no
 preference for or against extreme slider ratings. Relatively high values of $a$ indicate that
-participants, on average, had a strong tendency to choose their mode in the \emph{give-a-number}
+participants, on average, had a strong tendency to choose modal in the \emph{give-a-number}
 task.
@@
 \centering
+\footnotesize
 \begin{tabular}{rlllll}
- & $w$ & $\kappa$ & $\sigma$ & $a$ & $b$ \\ \hline
+ & $w$ & $\kappa$ & $\sigma$ & $a$ & $b$ \\ \midrule
   lower & 14.65 &  0.98 &  0.26 & 22.04 &  1.11 \\ 
@@
   upper & 16.48 &  0.99 &  0.31 & 32.95 &  1.42 \\ 
-   \hline
 \end{tabular}
-\caption{Summary statistics for posterior on parameters}
+\caption{Summary statistics for posteriors on parameters}
 \label{tab:SummaryStats}
@@
 
-Of most theoretical interest are posterior estimates of $Q_j$. \figref{fig:PosteriorQj}
-shows means of the posterior $Q_j$, with their 95\% HDIs, alongside the averaged normalized
-slider ratings. The latter provide a very good approximation of the latent, inferred
-population-level beliefs. Inspection of posteriors of individual $P_{ij}$ shows that there is
-ample variation between participants. Nonetheless, the way the model suggests we should think about
-harnessing the individual $P_{ij}$s under a population-level central tendency is closely
-approximated by simply averaging over slider ratings. Although the match is clearly not
-perfect, it is good enough to maintain that the latter are a reasonable (enough) and practical
-way of assessing what the crowd believes despite individual differences.
+Of most theoretical interest are posterior estimates of $Q_j$. \figref{fig:PosteriorQj} shows
+means of the posterior $Q_j$s, with their 95\% HDIs, alongside the mean slider ratings. The
+latter provide a very good approximation of the latent, inferred population-level
+beliefs. Inspection of posteriors of individual $P_{ij}$ shows that there is ample variation
+between participants. Nonetheless, the way the model suggests we should think about harnessing
+the individual $P_{ij}$s under a population-level central tendency is closely approximated by
+mean slider ratings. Although the match is not perfect, it is good enough to say that the
+latter are a practical way of approximating what the crowd believes despite individual
+differences.
 
@@
   \includegraphics[width = 0.5\textwidth]{plots/pop_priors.pdf}
-  \caption{Means of posteriors over $Q_j$ in black with gray area indicating 95\% HDIs. Red
-    lines give the average normalized slider ratings for comparison.}
+  \caption{Means of posteriors over $Q_j$ in black with gray area indicating 95\% HDIs. Red: mean slider ratings.}
   \label{fig:PosteriorQj}
@@
 Inferences based on the model are only as reliable as the model itself is plausible. Model
-criticism is therefore important. \figref{fig:PPCs} shows posterior predictive checks at
-the population-aggregate level for all of our three task types. For the binned histogram task,
-posterior predictions are spot-on. For the give-a-number task, some of the data is surprising despite 
-the model being trained on this very data. This could have various reasons: (i) the give-a-number data does
-not have a huge influence on the posterior likelihood, (ii) number choices may be influenced by
-saliency and/or roundness of numbers after all (and thus, not accurately reflect the true beliefs $Q_j$). Finally, there is one condition in the paired
+criticism is therefore important. \figref{fig:PPCs} shows posterior predictive checks at the
+population-aggregate level for all of our three task types. For the binned histogram task,
+posterior predictions are spot-on. For the give-a-number task, some of the data is surprising
+despite the model being trained on this very data. This could have various reasons: (i) the
+give-a-number data does not have a huge influence on the posterior likelihood, (ii) number
+choices may be influenced by saliency and/or roundness of numbers after all (and thus, not
+accurately reflect the true beliefs $Q_j$). Finally, there is one condition in the paired
 comparison task that the model definitely got wrong. This is the choice of what is more likely:
-One or none of 14 marbles thrown into a pool would sink. The model would predict that
+that one or that none of 14 marbles thrown into a pool would sink. The model predicts that
 almost everybody should answer that it is more likely that one marble sank. But that is not
@@
 of participants and items. We focus on the binned histograms task, because here the
-population-level posterior predictive checks revealed no systematic deviance. For the slider
-rating task, fixing a participant and an item, observations and replicates are probability
-vectors of length 15. In a first analysis, we used the mean of these probability vectors as a
-test statistic. The minimum posterior predictive $p$-value over all 20 (participants) times 8
-(items) cases was 0.13, suggesting that the means of observed $s_{ij}$ are non-surprising to
-the trained model. In a second analysis, we used entropy as a test statistic. Two cases gave
-posterior predictive $p$-values lower than 0.05. These were from the two participants who gave
-a very extreme slider rating for the ``marbles'' item, basically assigning all ``mass'' to the
-last bin. What this suggests is that the model can cope reasonably well also with
-individual-level data, but has problems accounting for ``extreme'' choices, given that the
-population-level hyper-prior on $P_{ij}$ will lead to shrinkage.
-\mht{Is it worth mentioning about the possilbiity of including subject-specific link function parameters e.g., $\kappa$ and $\alpha$?}
+population-level posterior predictive checks revealed no systematic deviance. For this task,
+fixing a participant and an item, observations and replicates are probability vectors of length
+15. In a first analysis, we used the mean of these probability vectors as a test statistic. The
+minimum posterior predictive $p$-value over all 20 (participants) times 8 (items) cases was
+0.13, suggesting that the means of observed $s_{ij}$ are non-surprising to the trained
+model. In a second analysis, we used entropy as a test statistic. Two cases gave posterior
+predictive $p$-values lower than 0.05. These were from the two participants who gave a very
+extreme slider rating for the ``marbles'' item, basically assigning all ``mass'' to the last
+bin. What this suggests is that the model can cope reasonably well also with individual-level
+data, but has problems accounting for ``extreme'' choices, given that the population-level
+hyper-prior on $P_{ij}$ will lead to shrinkage.
+
 \section{Conclusion}
@@
 research should evaluate whether this holds for other possible measures of subjective beliefs
-as well, such as from inferring parameterized distributions, the use of scoring rules, or the
-iterated learning paradigm.
+as well, such as iterated-learning or scoring-rule tasks.
 
@@
 explanations for these discrepancies. It therefore does not seem implausible that participants'
-latent beliefs could have generated the data from all three task types roughly in the way assumed by the model, with each subjective belief being an expression of a population-level
-central tendency. If that is so, then mean slider ratings from the binned histogram task are a practical and reliable approximation of what the crowd believes.
+latent beliefs could have generated the data from all three task types roughly in the way
+assumed by the model's link functions, with each subjective belief being an expression of a
+population-level central tendency. If that is so, then mean slider ratings from the binned
+histogram task are a practical and reliable approximation of what the crowd believes.
 
-Future research should investigate whether and how our results can be extended to other types of
-uncertain variables. Here, guided by the needs of many previous Bayesian cognitive models, we focused on uncertainty about unidimensional contiguous variables. It
-remains to be seen whether the binned histogram task can be applied to higher-dimensional
-variables \jd{like what?}. Also, although it is suggestive that our results extend to cases where uncertainty
-about a binary event is measured with a single (similarly labelled slider), this should be
-checked carefully. \jd{what does this mean?}
+Future research should investigate whether and how our results can be extended to other types
+of uncertain variables. Here, guided by the needs of many previous Bayesian cognitive models,
+we focused on uncertainty about unidimensional contiguous variables. It remains to be seen
+whether the binned histogram task can be applied to higher-dimensional variables, like joint
+prior beliefs over, say, \emph{height} and \emph{weight} of an individual or, more abstractly,
+different dimensions of a multi-dimensional property like \emph{intelligence}. Another
+challenge lies in devising means of eliciting prior expectations about properties that do not
+have clear and familiar measure terms, like the mentioned \emph{intelligence}.
+
+Finally, another aspect that our model has ignored so far are potential individual differences
+in the way subjective beliefs $P_{ij}$ generates responses. It is not unlikely that there is
+individual variation in at least some link function parameters, such as $\kappa$, which
+expresses end-point attraction or end-point aversion in the binned histograms task. A detailed
+investigation of such individual-level differences must be left to future work. Still, we
+believe that the model presented here is an important first step towards finding reliable and
+practical means of measuring what the crowd believes.
 
@@
 
-MF's work was supported by the Institutional Strategy of the University of T̈\"{u}bingen (Deutsche
-Forschungsgemeinschaft, ZUK 63). MF and JD's work was supported by the Priority Program XPrag.de (DFG Schwerpunktprogramm
-1727). This work was further supported by ONR grant N00014-13-1-0788 and a James S. McDonnell Foundation Scholar Award to NDG and an SNF Early Postdoc.Mobility Award to JD.
+MF's work was supported by the Institutional Strategy of the University of T̈\"{u}bingen
+(DFG, ZUK 63). MF, FD, AS and JD's work was supported by the
+Priority Program XPrag.de (DFG Schwerpunktprogramm 1727). This work was further supported by
+ONR grant N00014-13-1-0788 and a James S. McDonnell Foundation Scholar Award to NDG and an SNF
+Early Postdoc.Mobility Award to JD.
 

### 2016-02-01 text/01_CogSci_abstract/paper.tex
minor edits Anthea

@@
 implement and analyze. A prominent method in economics uses \emph{scoring rules}
-\citep{Savage1971:Elicitation-of-,AndersenFountain2014:Estimating-Subj,SchlagTremewan2014:A-penny-for-you}. Roughly
-put, scoring rules are schemes of rewarding participants' answers that make sure that responses
-(by rational participants) faithfully reflect a measure of interest: the actual subjective
-probability of an event, the mean of a distribution, a credible interval etc. Scoring-rule
-tasks require participants to become sufficiently familiar with the (usually probabilistic)
-payoff scheme. Another insightful, but also complex approach is to use a cleverly designed
-\emph{iterated learning paradigm} \citep{LewandowskyGriffiths2009:The-Wisdom-of-I}.
+\citep{Savage1971:Elicitation-of-,AndersenFountain2014:Estimating-Subj,SchlagTremewan2014:A-penny-for-you}. Scoring
+rules are complex schemes of assigning monetary rewards to participants' answers that make sure
+that responses (by rational participants) faithfully reflect a measure of interest: the actual
+subjective probability of an event, the mean of a distribution, a credible interval
+etc. Scoring-rule tasks require participants to become sufficiently familiar with the (usually
+probabilistic) payoff scheme. Another insightful, but also complex approach is to use a
+cleverly designed \emph{iterated learning paradigm}
+\citep{LewandowskyGriffiths2009:The-Wisdom-of-I}.
 
@@
 
-Explanatory success aside, the question remains whether mean slider ratings measure
-what we would like them to. Are these good approximations of a central tendency of participants'
-individual beliefs? Are these measures consistent with participants' behavior in other tasks, such
-as \emph{give-a-number}, as well? To address this issue we collected data in a within-subject paradigm 
-from three task types: (i) \emph{binned histograms}, (ii) \emph{give-a-number} and (iii)
-\emph{paired comparisons}, i.e., a direct comparison of two bins from the \emph{binned
-  histograms} task. Using a Bayesian hierarchical model, we used the data from all three
-dependent measures to infer likely latent subjective beliefs, where each subjective belief is,
-intuitively, a noisy perturbation of a population-level belief. We find that mean \emph{a
-  posteriori} population-level beliefs are approximated well by mean slider ratings, suggesting
-that the latter may indeed be a plausible, yet cheap measure of what the crowd believes.
-Moreover, the model is able to capture participants' behavior in all three task types reasonably
-well, suggesting that what mean slider ratings measure is what we think it is: a population-level
-central tendency of latent subjective beliefs.
+Explanatory success aside, the question remains whether mean slider ratings measure what we
+would like them to. Are these good approximations of a central tendency of participants'
+individual beliefs? Are these measures consistent with participants' behavior in other tasks,
+such as \emph{give-a-number}, as well? To address this issue we collected data in a
+within-subject paradigm from three task types: (i) \emph{binned histograms}, (ii)
+\emph{give-a-number} and (iii) \emph{paired comparisons}, i.e., a direct comparison of two bins
+from the \emph{binned histograms} task that included as a further consistency check. Using a
+Bayesian hierarchical model, we used the data from all three dependent measures to infer likely
+latent subjective beliefs, where each subjective belief is, intuitively, a noisy perturbation
+of a population-level belief. We find that mean \emph{a posteriori} population-level beliefs
+are approximated well by mean slider ratings, suggesting that the latter may indeed be a
+plausible, yet cheap measure of what the crowd believes.  Moreover, the model is able to
+capture participants' behavior in all three task types reasonably well, suggesting that what
+mean slider ratings measure is what we think it is: a population-level central tendency of
+latent subjective beliefs.
 
@@
 On \emph{binned histogram} (BH) trials, participants saw an item's context sentence and were
-asked to \emph{Please rate how likely it is that Y} (where Y corresponded to the appropriate BH
-frame in \tableref{tab:Items}). They then adjusted 15 sliders, one per bin, with five points
-labeled \emph{impossible}, \emph{not very likely}, \emph{neutral}, \emph{very likely}, and
-\emph{extremely likely}. Bins were determined by dividing up the entire interval spanned by the
-item's minimum and maximum value into equally sized bins (\tableref{tab:Items}).
+asked to \emph{Please rate how likely it is that Y} (where $Y$ came from the corresponding BH
+frame in \tableref{tab:Items}). They adjusted 15 continuous sliders, one per bin, with five
+points labeled \emph{impossible}, \emph{not very likely}, \emph{neutral}, \emph{very likely},
+and \emph{extremely likely}. Bins were determined by dividing the interval spanned by
+the item's minimum and maximum value into equally sized bins (\tableref{tab:Items}).
 

### 2016-02-01 text/01_CogSci_abstract/paper.tex
edit intro

@@
 \begin{abstract}
-  Experimental participants' beliefs about everyday events are both of theoretical interest in
-  their own right and an important ingredient in model building---especially in cognitive
-  Bayesian models of phenomena such as logical reasoning, future predictions, and language
+  People's beliefs about everyday events are both of theoretical interest in
+  their own right and an important ingredient in model building---especially in 
+  Bayesian cognitive models of phenomena such as logical reasoning, future predictions, and language
   use. Here, we explore several recently used methods for measuring subjective beliefs about
   unidimensional contiguous properties, such as the likely price of a new watch. We use a
-  hierarchical Bayesian model for inferring likely population-level beliefs as the central
+  hierarchical Bayesian data-analysis model for inferring likely population-level beliefs as the central
   tendency of participants' individual-level beliefs.  Three different dependent measures are
@@
 also in science. Scientific ascription of latent mental states plays an important role in many
-explanations of higher-order cognition: decision making, reasoning, language use etc. It is
+explanations of higher-order cognition: decision making, reasoning, language use, etc. It is
 therefore vital to have methods for validating explanatory mental state ascriptions.
@@
 participants believe about mundane events---their \emph{prior beliefs}. Take interpretation of
-language. Empirical data on whether a statement like ``That watch cost $n$ dollars'' is
-understood to convey speaker affect, rather than literal meaning, can be explained well by a
+language. 
+``That watch cost a million dollars,'' tends to be interpreted as hyperbole: conveying affect rather than literal truth.
+Empirical data on whether a statement like ``That watch cost $n$ dollars'' is
+understood as hyperbole can be explained well by a
 Bayesian model of utterance interpretation \citep{KaoWu2014:Nonliteral-Unde}. This model
 assigns a crucial role to an empirical measure of participants' expectations about the likely
-or normal price of a watch.  Other examples of domains in which empirically successful models
+or normal price of a watch---only when the uttered price is sufficiently unlikely \emph{a priori} will hyperbolic interpretation be possible.  
+Other examples of domains in which empirically successful models
 have included a measure of participants' prior beliefs include making predictions about
@@
 
+Many methods have been used to build prior distributions used in Bayesian cognitive models.
 One method of getting at subjective beliefs is to take actual frequencies as an approximation
@@
 
-Another frequent approach is to empirically measure subjective beliefs by \emph{give-a-number}
+Another common approach is to empirically measure subjective beliefs by \emph{give-a-number}
 tasks. The simplest version would be this: being told that John just bought a new watch,
@@
 \citep{TauberSteyvers2013:Inferring-Subje}.
+More sophisticated \emph{give-a-number} tasks give more information, but can be difficult to
+implement and analyze.
 
-More sophisticated \emph{give-a-number} tasks give more information, but can be difficult to
-implement and analyze. A prominent method in economics uses \emph{scoring rules}
-\citep{Savage1971:Elicitation-of-,AndersenFountain2014:Estimating-Subj,SchlagTremewan2014:A-penny-for-you}. Scoring
-rules are complex schemes of assigning monetary rewards to participants' answers that make sure
-that responses (by rational participants) faithfully reflect a measure of interest: the actual
-subjective probability of an event, the mean of a distribution, a credible interval
-etc. Scoring-rule tasks require participants to become sufficiently familiar with the (usually
-probabilistic) payoff scheme. Another insightful, but also complex approach is to use a
-cleverly designed \emph{iterated learning paradigm}
-\citep{LewandowskyGriffiths2009:The-Wisdom-of-I}.
+More complex elicitation methods include \emph{scoring rules} \citep{Savage1971:Elicitation-of-,AndersenFountain2014:Estimating-Subj,SchlagTremewan2014:A-penny-for-you}, prominent in economics, and the \emph{iterated learning paradigm} \citep{LewandowskyGriffiths2009:The-Wisdom-of-I}.
+These more complex methods are powerful but difficult to implement, and often assume very specific behavior from participants---such as optimal decision making under perfect knowledge of the payoff scheme.
+%}. Scoring
+%rules are complex schemes of assigning monetary rewards to participants' answers that make sure
+%that responses (by rational participants) faithfully reflect a measure of interest: the actual
+%subjective probability of an event, the mean of a distribution, a credible interval
+%etc. Scoring-rule tasks, however, require participants to become sufficiently familiar with the (usually
+%probabilistic) payoff scheme and assume close to optimal choices. 
+%
+%A final insightful, but also complex approach is to use a
+%cleverly designed
 
@@
 
+%
 In sum, there is tradeoff between simplicity of paradigms and their information
@@
 \cite<e.g.>{KaoWu2014:Nonliteral-Unde,KaoBergen2014:Formalizing--th,Tessler2015:Understanding-B,SchollerFranke2015:Semantic-values}. Participants
-adjust sliders to express (relative) subjective beliefs about how likely the value of an
+adjust sliders to express (relative) subjective beliefs about how likely it is that the value of an
 uncertain contiguous quantity lies in some interval of possible values (a bin). E.g., to report
@@
 individual beliefs? Are these measures consistent with participants' behavior in other tasks,
-such as \emph{give-a-number}, as well? To address this issue we collected data in a
+such as \emph{give-a-number}? To address this issue we collected data in a
 within-subject paradigm from three task types: (i) \emph{binned histograms}, (ii)
-\emph{give-a-number} and (iii) \emph{paired comparisons}, i.e., a direct comparison of two bins
-from the \emph{binned histograms} task that included as a further consistency check. Using a
-Bayesian hierarchical model, we used the data from all three dependent measures to infer likely
+\emph{give-a-number} and (iii) \emph{paired comparisons}. The latter asks participants for a direct comparison of two bins
+from the \emph{binned histograms} task, and was included as a further consistency check. 
+We used a Bayesian hierarchical model to analyze this data jointly.
+This model inferred likely
 latent subjective beliefs, where each subjective belief is, intuitively, a noisy perturbation
 of a population-level belief. We find that mean \emph{a posteriori} population-level beliefs
-are approximated well by mean slider ratings, suggesting that the latter may indeed be a
-plausible, yet cheap measure of what the crowd believes.  Moreover, the model is able to
+are approximated well by mean slider ratings from the binned-histograms task, suggesting that this may indeed be a
+sound and easy measure of what the crowd believes.  
+Moreover, the model is able to
 capture participants' behavior in all three task types reasonably well, suggesting that what

### 2016-02-01 text/01_CogSci_abstract/paper.tex
more edits everywhere

@@
 click on one of two options shown side by side -- whichever one they thought was more likely.
-Each option used the appropriate PC frame in\tableref{tab:Items} and numbers were filled in by
+Each option used the appropriate PC frame in \tableref{tab:Items} and numbers were filled in by
 comparing the following bins: 1 vs 2, 2 vs 6, 6 vs 11, 11 vs 14, and 14 vs 15. Paired
@@
 likelihood with which participant $i$ selects the higher bin for item $j$ in bin comparison
-condition $l$. Suppose $l$ is about comparing the lower bin $b_l$ to the higher $b_h$. The
-perhaps most natural approach would be to link the likelihood of choosing $b_h$ over $b_l$ to
+condition $l$. Suppose $l$ is about comparing the lower bin $b_l$ to the higher $b_h$. 
+Perhaps the most natural approach would be to link the likelihood of choosing $b_h$ over $b_l$ to
 the difference between $P_{ijb_h}$ and $P_{ijb_l}$. However, as discussed above, this does not
@@
 
-Means and bounds of 95\% HDIs of the posteriors for model parameters are in
+Means and bounds of 95\% high-density intervals (HDIs) of the posteriors for model parameters are in
 Table~\ref{tab:SummaryStats}. Posterior credible levels of $w$ allow for a limited amount of
@@
 preference for or against extreme slider ratings. Relatively high values of $a$ indicate that
-participants, on average, had a strong tendency to choose modal in the \emph{give-a-number}
+participants, on average, had a strong tendency to choose modal values in the \emph{give-a-number}
 task.
@@
 challenge lies in devising means of eliciting prior expectations about properties that do not
-have clear and familiar measure terms, like the mentioned \emph{intelligence}.
+have clear and familiar measure terms that can be used to label bins---again like \emph{intelligence}.
 

### 2016-02-02 text/01_CogSci_abstract/paper.tex
few typos

@@
   72074 T\"{u}bingen, Germany \AND
-  {\large \bf Erin Bennett, Judith Degen, Michael Henry Tessler, Justine Kao, Ciyang Qing, Noah D. Goodman}\\
-  \{ebennett, jdegen, mtessler, justinek, qciyang, ngoodman\}@stanford.edu \\
+  {\large \bf Erin Bennett, Judith Degen, Michael Henry Tessler, Justine Kao, Noah D. Goodman}\\
+  \{ebennett, jdegen, mtessler, justinek, ngoodman\}@stanford.edu \\
   Department of Psychology, Stanford University \\
@@
 %
-In sum, there is tradeoff between simplicity of paradigms and their information
+In sum, there is a tradeoff between simplicity of paradigms and their information
 content. Ideally, we would like to have an experimental method for measuring subjective beliefs
@@
 Finally, another aspect that our model has ignored so far are potential individual differences
-in the way subjective beliefs $P_{ij}$ generates responses. It is not unlikely that there is
+in the way subjective beliefs $P_{ij}$ generate responses. It is not unlikely that there is
 individual variation in at least some link function parameters, such as $\kappa$, which

### 2016-05-11 text/01_CogSci_abstract/paper.tex
CogSci revisions and list of future steps

@@
 \captionsetup{belowskip=0pt,aboveskip=2pt}
-\setlength{\textfloatsep}{9pt}
+\setlength{\textfloatsep}{8pt}
 
@@
 \begin{abstract}
-  People's beliefs about everyday events are both of theoretical interest in
-  their own right and an important ingredient in model building---especially in 
-  Bayesian cognitive models of phenomena such as logical reasoning, future predictions, and language
-  use. Here, we explore several recently used methods for measuring subjective beliefs about
-  unidimensional contiguous properties, such as the likely price of a new watch. We use a
-  hierarchical Bayesian data-analysis model for inferring likely population-level beliefs as the central
-  tendency of participants' individual-level beliefs.  Three different dependent measures are
-  used to infer latent beliefs: (i) slider ratings of (relative) likelihood of intervals of
-  values, (ii) a give-a-number task, and (iii) choice of the more likely of two intervals of
-  values. Our results suggest that using averaged normalized slider ratings for binned
-  quantities is a practical and fairly good approximator of inferred population-level beliefs.
+  People's beliefs about everyday events are both of theoretical interest in their own right
+  and an important ingredient in model building---especially in Bayesian cognitive models of
+  phenomena such as logical reasoning, future predictions, and language use. Here, we explore
+  several recently used methods for measuring subjective beliefs about unidimensional
+  contiguous properties, such as the likely price of a new watch. As a first step towards a way
+  of assessing and comparing belief elicitation methods, we use hierarchical Bayesian modeling
+  for inferring likely population-level beliefs as the central tendency of participants'
+  individual-level beliefs. Three different dependent measures are considered: (i) slider
+  ratings of (relative) likelihood of intervals of values, (ii) a give-a-number task, and (iii)
+  choice of the more likely of two intervals of values. Our results suggest that using averaged
+  normalized slider ratings for binned quantities is a practical and fairly good approximator
+  of inferred population-level beliefs.
 
@@
 
-We cannot look into a person's head. But in making sense of observed behavior we readily
-ascribe beliefs and desires to fellow agents. This happens intuitively, in folk psychology, but
-also in science. Scientific ascription of latent mental states plays an important role in many
-explanations of higher-order cognition: decision making, reasoning, language use, etc. It is
-therefore vital to have methods for validating explanatory mental state ascriptions.
+When trying to understand observed behavior we readily ascribe beliefs and desires to fellow
+agents. This happens intuitively, in folk psychology, but also in science. Scientific
+ascription of latent mental states plays an important role in many explanations of higher-order
+cognition: decision making, reasoning, language use, etc. It is therefore vital to have methods
+for validating explanatory mental state ascriptions.
 
@@
 participants believe about mundane events---their \emph{prior beliefs}. Take interpretation of
-language. 
-``That watch cost a million dollars,'' tends to be interpreted as hyperbole: conveying affect rather than literal truth.
-Empirical data on whether a statement like ``That watch cost $n$ dollars'' is
-understood as hyperbole can be explained well by a
-Bayesian model of utterance interpretation \citep{KaoWu2014:Nonliteral-Unde}. This model
-assigns a crucial role to an empirical measure of participants' expectations about the likely
-or normal price of a watch---only when the uttered price is sufficiently unlikely \emph{a priori} will hyperbolic interpretation be possible.  
-Other examples of domains in which empirically successful models
-have included a measure of participants' prior beliefs include making predictions about
-everyday events \citep{GriffithsTenenbaum2006:Optimal-Predict}, referential communication
-\citep{FrankGoodman2012:Predicting-Prag}, the strength of pragmatic enrichments
-\citep{DegenTessler2015:Wonky-worlds:-L}, or the interpretation of quantifiers
+language.  ``That watch cost a million dollars,'' tends to be interpreted as hyperbole:
+conveying affect rather than literal truth.  Empirical data on whether similar statements are
+understood as hyperbole can be explained well by a Bayesian model of utterance interpretation
+\citep{KaoWu2014:Nonliteral-Unde}. This model assigns a crucial role to an empirical measure of
+participants' expectations about the likely or normal price of a watch---only when the uttered
+price is sufficiently unlikely \emph{a priori} will hyperbolic interpretation be possible.
+Other examples of domains in which empirically successful models have included a measure of
+participants' prior beliefs include making predictions about everyday events
+\citep{GriffithsTenenbaum2006:Optimal-Predict}, referential reasoning
+\citep{FrankGoodman2012:Predicting-Prag}, strength of pragmatic enrichments
+\citep{DegenTessler2015:Wonky-worlds:-L}, or quantifier interpretation
 \citep{SchollerFranke2015:Semantic-values}.
@@
 
-More complex elicitation methods include \emph{scoring rules} \citep{Savage1971:Elicitation-of-,AndersenFountain2014:Estimating-Subj,SchlagTremewan2014:A-penny-for-you}, prominent in economics, and the \emph{iterated learning paradigm} \citep{LewandowskyGriffiths2009:The-Wisdom-of-I}.
-These more complex methods are powerful but difficult to implement, and often assume very specific behavior from participants---such as optimal decision making under perfect knowledge of the payoff scheme.
+More complex elicitation methods include \emph{scoring rules}
+\citep{Savage1971:Elicitation-of-,AndersenFountain2014:Estimating-Subj,SchlagTremewan2014:A-penny-for-you},
+prominent in economics, and the \emph{iterated learning paradigm}
+\citep{LewandowskyGriffiths2009:The-Wisdom-of-I}.  These methods are powerful but difficult to
+implement, and often assume very specific behavior from participants---such as optimal decision
+making under perfect knowledge of the payoff scheme.
 %}. Scoring
@@
 \cite<e.g.>{KaoWu2014:Nonliteral-Unde,KaoBergen2014:Formalizing--th,Tessler2015:Understanding-B,SchollerFranke2015:Semantic-values}. Participants
-adjust sliders to express (relative) subjective beliefs about how likely it is that the value of an
-uncertain contiguous quantity lies in some interval of possible values (a bin). E.g., to report
-beliefs about the a watch's price, participants adjust sliders whose endpoints are labelled
+adjust sliders to express (relative) subjective beliefs about how likely it is that the value
+of an uncertain contiguous quantity lies in some interval of possible values (a bin). E.g., to
+report beliefs about a watch's price, participants adjust sliders whose endpoints are labelled
 ``extremely likely'' and ``impossible.'' Each slider corresponds to one of 15 bins that
 partition the range of plausible values (established by a pre-test), such as ``\$0-\$50'',
-``\$50-\$100'', etc. up to ``\$700-\$750'' and ``more than \$750.'' Each participant's slider
+``\$50-\$100'', etc.~up to ``\$700-\$750'' and ``more than \$750.'' Each participant's slider
 ratings are normalized and the results averaged across participants. The resulting \emph{mean
@@
 individual beliefs? Are these measures consistent with participants' behavior in other tasks,
-such as \emph{give-a-number}? To address this issue we collected data in a
-within-subject paradigm from three task types: (i) \emph{binned histograms}, (ii)
-\emph{give-a-number} and (iii) \emph{paired comparisons}. The latter asks participants for a direct comparison of two bins
-from the \emph{binned histograms} task, and was included as a further consistency check. 
-We used a Bayesian hierarchical model to analyze this data jointly.
-This model inferred likely
-latent subjective beliefs, where each subjective belief is, intuitively, a noisy perturbation
-of a population-level belief. We find that mean \emph{a posteriori} population-level beliefs
-are approximated well by mean slider ratings from the binned-histograms task, suggesting that this may indeed be a
-sound and easy measure of what the crowd believes.  
-Moreover, the model is able to
-capture participants' behavior in all three task types reasonably well, suggesting that what
-mean slider ratings measure is what we think it is: a population-level central tendency of
-latent subjective beliefs.
+such as \emph{give-a-number}? To scrutinize the binned histogram task we address these issues
+by collecting data in a within-subject paradigm from three task types: (i) \emph{binned
+  histograms}, (ii) \emph{give-a-number} and (iii) \emph{paired comparisons}. The latter asks
+participants for a direct comparison of two bins from the \emph{binned histograms} task, and
+was included as a further consistency check.  We used a Bayesian hierarchical model to analyze
+this data jointly.  The model infers latent subjective beliefs, where each subjective
+belief is, intuitively, a noisy perturbation of a population-level belief. We find that mean
+\emph{a posteriori} population-level beliefs are approximated well by mean slider ratings from
+the binned-histograms task, suggesting that this may indeed be a sound and easy measure of what
+the crowd believes.  Moreover, the model is able to capture participants' behavior in all three
+task types reasonably well, suggesting that what mean slider ratings measure is what we think
+it is: a population-level central tendency of latent subjective beliefs.
 
@@
 speakers over Mechanical Turk and collected responses for eight different items (listed in
-\tableref{tab:Items}) using three different dependent measures: \emph{binned histograms},
-\emph{give-a-number} and \emph{paired comparisons}. Each participant rated each item using each
-dependent measure. Trial order was randomized.
+\tableref{tab:Items}) using the three different dependent measures mentioned above. Each
+participant rated each item using each dependent measure. Trial order was randomized.
 
@@
 participant $i \in \set{1, \dots, 20}$ for item $j \in \set{1, \dots, 8}$ and bin
-$k \in \set{1, \dots, 15}$ from the binned histogram task; (ii) the bins $n_{ij} \in \set{1, \dots, 15}$ in which participant
-$i$'s number choice for item $j$ was in the give-a-number task; and (iii) the binary choices $c_{ijl} \in \set{0,1}$ of
-whether participant $i$ selected the higher bin for item $j$ in the paired comparison task for each comparison
-$l$. There are two simplifications in need of commenting. In (i), we focus on slider ratings after
-normalizing for each participant, because we assume that slider adjustments reflect relative, not
-absolute estimates of subjective beliefs. In (ii), we focus on bin choices, not actual number
-choices, in order to avoid, as much as possible, considerations of salience of particular
-numbers, and also because otherwise data from items with smaller domains of plausible numbers
-would get more weight than data from items which allow for a wider set of number choices.
+$k \in \set{1, \dots, 15}$ from the binned histogram task; (ii) the bins
+$n_{ij} \in \set{1, \dots, 15}$ in which participant $i$'s number choice for item $j$ was in
+the give-a-number task; and (iii) the binary choices $c_{ijl} \in \set{0,1}$ of whether
+participant $i$ selected the higher bin for item $j$ in the paired comparison task for each
+comparison $l$. There are two simplifications in need of commenting. In (i), we focus on slider
+ratings after normalizing for each participant, because we assume that slider adjustments
+reflect relative, not absolute estimates of subjective beliefs. In (ii), we focus on bin
+choices, not actual number choices, in order to avoid, as much as possible, considerations of
+salience of particular numbers, and also because otherwise data from items with smaller domains
+of plausible numbers would get more weight than data from items with a wider range of
+number choices. (Future work should investigate the relation to an explicit \emph{give-a-bin}
+task.)
 
@@
 
-Of most theoretical interest are posterior estimates of $Q_j$. \figref{fig:PosteriorQj} shows
-means of the posterior $Q_j$s, with their 95\% HDIs, alongside the mean slider ratings. The
-latter provide a very good approximation of the latent, inferred population-level
-beliefs. Inspection of posteriors of individual $P_{ij}$ shows that there is ample variation
-between participants. Nonetheless, the way the model suggests we should think about harnessing
-the individual $P_{ij}$s under a population-level central tendency is closely approximated by
-mean slider ratings. Although the match is not perfect, it is good enough to say that the
-latter are a practical way of approximating what the crowd believes despite individual
-differences.
+Posterior estimates of $Q_j$ are the most relevant. \figref{fig:PosteriorQj} shows their means
+with their 95\% HDIs, alongside the mean slider ratings. The latter provide a very good
+approximation of the inferred population-level beliefs. Inspection of posteriors of
+individual $P_{ij}$ shows that there is ample variation between participants. Still, the
+way the model suggests we should think about harnessing the individual $P_{ij}$s under a
+population-level central tendency is closely approximated by mean slider ratings. Although the
+match is not perfect, it is good enough to say that the latter are a practical way of
+approximating what the crowd believes despite individual differences.
 
@@
 looked at posterior predictive $p$-values \citep{GelmanCarlin2014:Bayesian-Data-A} at the level
-of participants and items. We focus on the binned histograms task, because here the
-population-level posterior predictive checks revealed no systematic deviance. For this task,
-fixing a participant and an item, observations and replicates are probability vectors of length
-15. In a first analysis, we used the mean of these probability vectors as a test statistic. The
-minimum posterior predictive $p$-value over all 20 (participants) times 8 (items) cases was
-0.13, suggesting that the means of observed $s_{ij}$ are non-surprising to the trained
-model. In a second analysis, we used entropy as a test statistic. Two cases gave posterior
-predictive $p$-values lower than 0.05. These were from the two participants who gave a very
-extreme slider rating for the ``marbles'' item, basically assigning all ``mass'' to the last
-bin. What this suggests is that the model can cope reasonably well also with individual-level
-data, but has problems accounting for ``extreme'' choices, given that the population-level
-hyper-prior on $P_{ij}$ will lead to shrinkage.
+of participants and items. We look at the binned histograms task, because this is our main
+focus here and population-level posterior predictive checks for BH revealed no systematic
+deviance. Fixing a participant and an item, observations and replicates are probability vectors
+of length 15. In a first analysis, we used the mean of these probability vectors as a test
+statistic. The minimum posterior predictive $p$-value over all 20 (participants) times 8
+(items) cases was 0.13, suggesting that the means of observed $s_{ij}$ are non-surprising to
+the trained model. In a second analysis, we used entropy as a test statistic. Two cases gave
+posterior predictive $p$-values lower than 0.05. These were from the two participants who gave
+a very extreme slider rating for the ``marbles'' item, basically assigning all ``mass'' to the
+last bin. What this suggests is that the model can cope reasonably well also with
+individual-level data, but, somewhat unsurprisingly, has problems accounting for ``extreme''
+choices, given that the population-level hyper-prior on $P_{ij}$ will lead to shrinkage.
 
@@
 ONR grant N00014-13-1-0788 and a James S. McDonnell Foundation Scholar Award to NDG and an SNF
-Early Postdoc.Mobility Award to JD.
+Early Postdoc.Mobility Award to JD. Thanks to three anonymous reviewers for helpful suggestions
+and criticism.
 

### 2016-08-06 experiments/exp2_Manski/results/data/data_prep.R
renaming and restructuring of model files

@@ -29,7 +29,7 @@ binned_histogram = droplevels(subset(d, measure == "binned_histogram")) %>%
 binned_histogram$bin_num = 1:15
 
 y.slider = array(0, dim = c(50,8,15))
-for (i in 1:nrow(bin_dat)){
+for (i in 1:nrow(binned_histogram)){
   y.slider[binned_histogram$workerid[i] + 1, 
            as.numeric(binned_histogram$tag)[i],
            binned_histogram$bin_num[i]] = binned_histogram$response[i]

### 2016-08-06 models/04_Exp2_Manski_JAGS/helpers/helpers.R
modeling of Exp 2 data started

@@ -1,0 +1,216 @@
+library('dplyr')
+library('ggplot2')
+library('reshape2')
+
+
+
+logit = function(x, k = 1) k * log(x / (1-x))
+logistic = function(x, k = 1, zero = 0) 1 / (1 + exp(-1/k*(x - zero)))
+add.margin <- function(x, eps = 0.000001) (1 - 2*eps)*x + eps 
+# calibrate <- function(x) (x - min(x)) / (max(x) - min(x))
+normalize <- function(x) (x / sum(x))
+
+
+# Computes highest density interval from a sample of representative values,
+#   estimated as shortest credible interval.
+# Arguments:
+#   sampleVec
+#     is a vector of representative values from a probability distribution.
+#   credMass
+#     is a scalar between 0 and 1, indicating the mass within the credible
+#     interval that is to be estimated.
+# Value:
+#   HDIlim is a vector containing the limits of the HDI
+HDIofMCMC = function( sampleVec , credMass = 0.95 ) {
+  sortedPts = sort( sampleVec )
+  ciIdxInc = ceiling( credMass * length( sortedPts ) )
+  nCIs = length( sortedPts ) - ciIdxInc
+  ciWidth = rep( 0 , nCIs )
+  for ( i in 1:nCIs ) {
+    ciWidth[ i ] = sortedPts[ i + ciIdxInc ] - sortedPts[ i ]
+  }
+  HDImin = sortedPts[ which.min( ciWidth ) ]
+  HDImax = sortedPts[ which.min( ciWidth ) + ciIdxInc ]
+  HDIlim = c( HDImin , HDImax )
+  return( HDIlim )
+}
+
+
+# get the all the samples into a data.frame
+# first get the parameters, and the the posterior predictive values
+# otherwise it messes with the number of bins; until now: only sliderPPC
+construct_ppvs <- function(samples, ppv = 'y.sliderPPC') {
+  slist <- samples$BUGSoutput$sims.list
+  ppv_df <- slist[[ppv]]
+  ppv_df <- tbl_df(melt(ppv_df))
+  
+  # poor man's solution as of yet
+  if (ppv == 'y.sliderPPC') {
+    bins <- rep(bin_dat$bin_num, each = max(ppv_df$Var1))
+    items <- rep(bin_dat$tag, each = max(ppv_df$Var1))
+    workerid = rep(bin_dat$workerid, each = max(ppv_df$Var1))
+    ppv_df <- ppv_df %>%
[…]

### 2016-08-06 models/04_Exp2_Manski_JAGS/main.R

@@ -1,0 +1,57 @@
+library('coda')
+library('ggmcmc')
+library('jagsUI') # for parallel computing
+source('helpers/helpers.R')
+source('~/Desktop/data/svn/ProComPrag/dev_tmp/typicality_quantifiers/model/helpers.r')
+source('process_data.R')
+
+
+saveFlag = TRUE
+
+data_aggr <- list(y.slider = y.slider, y.number = y.number, y.choice = y.choice, 
+                  higher = higher, lower = lower,
+                  nSubjs = 20, nItems = 8, nBins = 15, nLightConds = 5,
+                  ones = rep(1,15))
+params <- c('w', 
+            'a', 
+            'sigma', 
+            'item.pop',
+            'subj',
+            'k', 
+            'b',
+            "y.choicePPC", 
+            "y.numberPPC",
+            "y.sliderPPC"
+            )
+
+model = "models/model.jags.R"
+burnin = 10000
+iter = 10000
+out = jags(data = data_aggr,
+            inits = NULL,
+            parameters.to.save = params,
+            codaOnly = c("y.choicePPC", "y.numberPPC", "y.sliderPPC", "subj"),
+            model.file = model,
+            n.chains = 2,
+            n.adapt = 5000,
+            n.iter = iter + burnin,
+            n.burnin = burnin,
+            n.thin = 2, 
+            DIC = TRUE,
+            verbose = TRUE,
+            parallel = TRUE)
+
+if (saveFlag) { save(out, file = "~/Desktop/Dropbox/priors_data/outTMP.Rdat") }
+
+stop()
+
+p = out$sims.list$item.pop[1,1,]
+
+ggplot(melt(rdirichlet(alpha= p*190 + 1, n = 10)), aes(x = Var2, y = value, fill = factor(Var1))) + geom_line(color = "gray") + geom_line(data = cbind(melt(p), x=1:15, Var1 = "a"), aes(x = x, y = value, color = "red"))
+
+x = out$sims.list$subj[,1,1,]
+y = out$sims.list$subj[,2,1,]
+z = out$sims.list$subj[,3,1,]
+
+ggplot(melt(x), aes(x = Var2, y = value, fill = factor(Var1))) + geom_line(color = "gray") + geom_line(data = melt(y), color = "darkgray") +
+  geom_line(data = melt(z), color = "green")

### 2016-08-06 models/04_Exp2_Manski_JAGS/models/model.jags.R

@@ -1,0 +1,71 @@
+model {
+  ## SLIDERS
+  for (i in 1:nSubjs) {
+    for (j in 1:nItems) {
+      for (k in 1:nBins){
+        y[i,j,k] <- k.skew[i] * log(subj[i,j,k] / (1 - subj[i,j,k]))
+        y.slider[i,j,k]    ~ dnorm(y[i,j,k], tau)
+        y.sliderPPC[i,j,k] ~ dnorm(y[i,j,k], tau)   
+      }
+    }
+  }
+  ## NUMBERS
+  for (i in 1:nSubjs) {
+    for (j in 1:nItems) {
+      for (k in 1:nBins){
+        cp[i,j,k] = exp(a * subj[i,j,k])
+      }
+      y.number[i,j,1]    ~ dcat(cp[i,j,1:15])
+      y.numberPPC[i,j,1] ~ dcat(cp[i,j,1:15])
+    }
+  }
+  x[1:15] = cp[1,1,1:15]
+  
+  ## CHOICE
+  for (i in 1:nSubjs) {
+    for (j in 1:nItems) {
+      for (b in 1:15) {
+        pTemp[i,j,b] <- exp( 1000 * subj[i,j,b] )
+        pTempNorm[i,j,b] = pTemp[i,j,b] / sum(pTemp[i,j,1:15])
+      }
+      mode[i,j] ~ dcat(pTempNorm[i,j,1:15])
+      for (k in 1:nLightConds) {
+        # difference between choice alternatives and modal bin
+        hi[i,j,k] = ifelse(abs(mode[i,j] - lower[i,j,k]) > abs(mode[i,j] - higher[i,j,k]), 1, 0) 
+        lo[i,j,k] = ifelse(abs(mode[i,j] - lower[i,j,k]) < abs(mode[i,j] - higher[i,j,k]), 1, 0)
+        probT[i,j,k,1] <- exp(b * (1 + lo[i,j,k] - hi[i,j,k]))
+        probT[i,j,k,2] <- exp(b * (1 + hi[i,j,k] - lo[i,j,k]))
+        prob[i,j,k] <- probT[i,j,k,2] / sum(probT[i,j,k,1:2])
+        y.choice[i,j,k] ~ dbern(prob[i,j,k])                            
+        y.choicePPC[i,j,k] ~ dbern(prob[i,j,k])
+#         d.lo[i,j,k] <- abs(mode[i,j] - lower[i,j,k])
[…]

### 2016-08-06 models/04_Exp2_Manski_JAGS/plotting.R

@@ -1,0 +1,265 @@
+library('coda')
+library('ggmcmc')
+library('jagsUI') # for parallel computing
+source('helpers/helpers.R')
+source('~/Desktop/data/svn/ProComPrag/dev_tmp/typicality_quantifiers/model/helpers.r')
+source('process_data.R')
+
+theme_set(theme_bw() + theme(plot.background=element_blank()) )
+
+readFlag = TRUE
+savePlots = TRUE
+
+if (readFlag){
+  # load(file = "/Users/micha/Desktop/Dropbox/priors_data/out_25.Rdat")
+  # load(file = "/Users/micha/Desktop/Dropbox/priors_data/outTMP.Rdat")
+  load(file = "/Users/micha/Desktop/Dropbox/priors_data/out.Rdat")
+} else {
+  # source('main.R')
+}
+
+# prepare samples
+csamples = tbl_df(melt(out$sims.list))
+colnames(csamples) = c("value", "step", "subject", "item", "bin", "variable")
+csamples$item = levels(factor(bin_dat$tag))[csamples$item]
+
+#######################
+# posteriors
+#######################
+
+# item pop
+meansIP = tbl_df(melt(out$sims.list$item.pop)) %>%
+  rename(step = Var1, item = Var2, bin = Var3) %>%
+  mutate(item = levels(factor(bin_dat$tag))[item] ) %>%
+  group_by(item, bin) %>%
+  summarise(
+    mean = mean(value),
+    max = HDIofMCMC(value)[2],
+    min = HDIofMCMC(value)[1]
+  )
+meansIP$y_emp = y.slider_means$mymean
+pop_priors = ggplot(meansIP, aes(x = bin, y = mean)) + geom_line() + geom_point() + facet_wrap(~ item, scale = "free") +
+  geom_ribbon(aes(ymin=min, ymax=max), fill="gray", alpha="0.5") +
+  geom_line(aes(x = bin, y = y_emp) , color = "firebrick")  + geom_point(aes(x = bin, y = y_emp ), color = "firebrick") +
+  ylab(" mean posterior Q_{ijk} / average slider rating")
+
+pop_priorsPaper = ggplot(meansIP, aes(x = bin, y = mean)) + geom_line() + geom_point() + facet_wrap(~ item, scales = "free_y", ncol=2) +
+  geom_ribbon(aes(ymin=min, ymax=max), fill="gray", alpha="0.5") +
[…]

### 2016-08-06 models/04_Exp2_Manski_JAGS/postprocessing.R

@@ -1,0 +1,81 @@
+library('coda')
+library('ggmcmc')
+library('jagsUI') # for parallel computing
+library('xtable')
+source('helpers/helpers.R')
+# source('~/Desktop/data/svn/ProComPrag/dev_tmp/typicality_quantifiers/model/helpers.r')
+source('process_data.R')
+
+theme_set(theme_bw() + theme(plot.background=element_blank()) )
+
+readFlag = TRUE
+
+if (readFlag){
+  # load(file = "/Users/micha/Desktop/Dropbox/priors_data/out_25.Rdat")
+  load(file = "/Users/micha/Desktop/Dropbox/priors_data/out.Rdat")
+} else {
+  # source('main.R')
+}
+
+## get summary data
+statsSum = data.frame(
+  parameter = c("w", "kappa", "sigma", "a", "b"),
+  hdiMean = c(mean(out$sims.list$w), 
+              mean(out$sims.list$k), 
+              mean(out$sims.list$sigma),
+              mean(out$sims.list$a),
+              mean(out$sims.list$b)),
+  hdiLow = c(HDIofMCMC(out$sims.list$w)[1], 
+             HDIofMCMC(out$sims.list$k)[1], 
+             HDIofMCMC(out$sims.list$sigma)[1],
+             HDIofMCMC(out$sims.list$a)[1],
+             HDIofMCMC(out$sims.list$b)[1]),
+  hdiHigh = c(HDIofMCMC(out$sims.list$w)[2], 
+              HDIofMCMC(out$sims.list$k)[2], 
+              HDIofMCMC(out$sims.list$sigma)[2],
+              HDIofMCMC(out$sims.list$a)[2],
+              HDIofMCMC(out$sims.list$b)[2]))
+
+show(xtable(t(statsSum)))
+  
+
+
+
+
+entropy = function(p){
+  - sum(p*log(p))
+}
+
+y.rep = logistic(out$sims.list$y.sliderPPC)
+y.emp = logistic(y.slider)
+
+posteriorMeans = array(0, dim = c(dim(y.rep)[1], dim(y.rep)[2], dim(y.rep)[3]))
+posteriorEntrs = array(0, dim = c(dim(y.rep)[1], dim(y.rep)[2], dim(y.rep)[3]))
+pValueMean = matrix(0, nrow = 20, ncol = 8)
[…]

### 2016-08-06 models/04_Exp2_Manski_JAGS/process_data.R

@@ -1,0 +1,45 @@
+require(dplyr)
+
+d = read.table("data/results.csv", header=T, sep=",",quote="\"")
+bins = read.csv("data/bins.txt",header=T)
+
+## number choices (treated as bins)
+give_number = d %>% filter(measure == "give_number") %>% droplevels() %>%
+  select(workerid, tag, response, max)  %>%
+  mutate(response = as.numeric(as.character(response)))
+bins$Bin = bins$Bin + 1
+give_number$chosen_bin = sapply(1:nrow(give_number), function(i) {
+  response = give_number$response[i]
+  tag = as.character(give_number$tag)[i]
+  bins[bins$Item == tag & bins$Min <= response & bins$Max >= response,]$Bin
+})
+
+y.number = array(0, dim = c(50,8,1))
+for (i in 1:nrow(give_number)){
+  y.number[give_number$workerid[i] + 1, 
+           as.numeric(give_number$tag)[i], # WTH does this enumerate starting at 2?
+           1] = give_number$chosen_bin[i]
+}
+dimnames(y.number)[[2]] = sapply(1:8, function(i) levels(give_number$tag)[i])
+
+## binned histograms
+binned_histogram = droplevels(subset(d, measure == "binned_histogram")) %>%
+  select(workerid, tag, bin, response) %>% 
+  mutate(response = as.numeric(as.character(response)))
+binned_histogram$bin_num = 1:15
+
+y.slider = array(0, dim = c(50,8,15))
+for (i in 1:nrow(binned_histogram)){
+  y.slider[binned_histogram$workerid[i] + 1, 
+           as.numeric(binned_histogram$tag)[i],
+           binned_histogram$bin_num[i]] = binned_histogram$response[i]
+}
+dimnames(y.slider)[[2]] = sapply(1:8, function(i) levels(binned_histogram$tag)[i])
+for (i in 1:dim(y.slider)[1]) {
+  for (j in 1:dim(y.slider)[2]) {
+    if (sum(y.slider[i,j,]) != 0) {
+      y.slider[i,j,] = y.slider[i,j,] / sum(y.slider[i,j,])
+    }
+  }
+}
+# next: add.margin, take logit, get summaries as well

### 2016-08-06 models/04_Exp2_Manski_JAGS/main.R
model fit Exp 2 data first stable

@@ -5,12 +5,11 @@ source('helpers/helpers.R')
 source('~/Desktop/data/svn/ProComPrag/dev_tmp/typicality_quantifiers/model/helpers.r')
 source('process_data.R')
 
-
 saveFlag = TRUE
 
 data_aggr <- list(y.slider = y.slider, y.number = y.number, y.choice = y.choice, 
                   higher = higher, lower = lower,
-                  nSubjs = 20, nItems = 8, nBins = 15, nLightConds = 5,
+                  nSubjs = 50, nItems = 8, nBins = 15, nLightConds = 5,
                   ones = rep(1,15))
 params <- c('w', 
             'a', 
@@ -25,6 +24,8 @@ params <- c('w',
             )
 
 model = "models/model.jags.R"
+# burnin = 100
+# iter = 100
 burnin = 10000
 iter = 10000
 out = jags(data = data_aggr,
@@ -41,7 +42,7 @@ out = jags(data = data_aggr,
             verbose = TRUE,
             parallel = TRUE)
 
-if (saveFlag) { save(out, file = "~/Desktop/Dropbox/priors_data/outTMP.Rdat") }
+if (saveFlag) { save(out, file = "~/Desktop/Dropbox/priors_data/outExp2.Rdat") }
 
 stop()
 

### 2016-08-06 models/04_Exp2_Manski_JAGS/plotting.R

@@ -21,7 +21,7 @@ if (readFlag){
 # prepare samples
 csamples = tbl_df(melt(out$sims.list))
 colnames(csamples) = c("value", "step", "subject", "item", "bin", "variable")
-csamples$item = levels(factor(bin_dat$tag))[csamples$item]
+csamples$item = levels(factor(binned_histogram$tag))[csamples$item]
 
 #######################
 # posteriors
@@ -30,7 +30,7 @@ csamples$item = levels(factor(bin_dat$tag))[csamples$item]
 # item pop
 meansIP = tbl_df(melt(out$sims.list$item.pop)) %>%
   rename(step = Var1, item = Var2, bin = Var3) %>%
-  mutate(item = levels(factor(bin_dat$tag))[item] ) %>%
+  mutate(item = levels(factor(binned_histogram$tag))[item] ) %>%
   group_by(item, bin) %>%
   summarise(
     mean = mean(value),
@@ -52,7 +52,7 @@ pop_priorsPaper = ggplot(meansIP, aes(x = bin, y = mean)) + geom_line() + geom_p
 # subjective priors
 meansIPSubj = tbl_df(melt(out$sims.list$subj)) %>%
   rename(step = Var1, subject = Var2, item = Var3, bin = Var4) %>%
-  mutate(item = levels(factor(bin_dat$tag))[item] ) %>%
+  mutate(item = levels(factor(binned_histogram$tag))[item] ) %>%
   filter(subject <= 20) %>%
   group_by(subject, item, bin) %>%
   summarise(
@@ -129,6 +129,7 @@ plotSliderDataPaper = ggplot(slider_aggr, aes(x = bin, y = y_means)) + geom_line
 
 
 # numbers
+number_dat = give_number
 number_dat$chosen_bin = factor(number_dat$chosen_bin, levels = 1:15)
 DNm = tbl_df(melt(table(number_dat$tag, number_dat$chosen_bin))) %>%
   rename(item = Var1, bin = Var2)
@@ -250,16 +251,16 @@ if (savePlots){
 
 # save plots for CogSciPaper  
 factor = 0.55
-if (savePlots){
-  # pdfs
-  ggsave('../text/01_CogSci_abstract/plots/pop_priors.pdf', pop_priorsPaper, width=10*factor, height = 8*factor)
-#   ggsave('../text/01_CogSci_abstract/plots/pop_priorsSubj.pdf', pop_priorsSubj, width=10*factor, height = 8*factor)
[…]

### 2016-08-06 models/04_Exp2_Manski_JAGS/process_data.R

@@ -1,4 +1,15 @@
 require(dplyr)
+require('bootstrap')
+library('rjson')
+library('stringr')
+source('helpers/helpers.R')
+
+## for bootstrapping 95% confidence intervals
+theta <- function(x,xdata,na.rm=T) {mean(xdata[x],na.rm=na.rm)}
+ci.low <- function(x,na.rm=T) {
+  mean(x,na.rm=na.rm) - quantile(bootstrap(1:length(x),1000,theta,x,na.rm=na.rm)$thetastar,.025,na.rm=na.rm)}
+ci.high <- function(x,na.rm=T) {
+  quantile(bootstrap(1:length(x),1000,theta,x,na.rm=na.rm)$thetastar,.975,na.rm=na.rm) - mean(x,na.rm=na.rm)}
 
 d = read.table("data/results.csv", header=T, sep=",",quote="\"")
 bins = read.csv("data/bins.txt",header=T)
@@ -23,23 +34,106 @@ for (i in 1:nrow(give_number)){
 dimnames(y.number)[[2]] = sapply(1:8, function(i) levels(give_number$tag)[i])
 
 ## binned histograms
+
 binned_histogram = droplevels(subset(d, measure == "binned_histogram")) %>%
   select(workerid, tag, bin, response) %>% 
-  mutate(response = as.numeric(as.character(response)))
+  mutate(response = as.numeric(as.character(response))) %>%
+  group_by(tag, workerid) %>% mutate(nresponse = normalize(response))
 binned_histogram$bin_num = 1:15
 
 y.slider = array(0, dim = c(50,8,15))
 for (i in 1:nrow(binned_histogram)){
   y.slider[binned_histogram$workerid[i] + 1, 
            as.numeric(binned_histogram$tag)[i],
-           binned_histogram$bin_num[i]] = binned_histogram$response[i]
+           binned_histogram$bin_num[i]] = binned_histogram$nresponse[i]
 }
 dimnames(y.slider)[[2]] = sapply(1:8, function(i) levels(binned_histogram$tag)[i])
-for (i in 1:dim(y.slider)[1]) {
-  for (j in 1:dim(y.slider)[2]) {
-    if (sum(y.slider[i,j,]) != 0) {
-      y.slider[i,j,] = y.slider[i,j,] / sum(y.slider[i,j,])
-    }
-  }
+y.slider = logit(add.margin(y.slider))
[…]

### 2016-08-06 models/04_Exp2_Manski_JAGS/plotting.R
Exp 2 model fits and plots

@@ -10,10 +10,12 @@ theme_set(theme_bw() + theme(plot.background=element_blank()) )
 readFlag = TRUE
 savePlots = TRUE
 
+nSubj = 50
+
 if (readFlag){
   # load(file = "/Users/micha/Desktop/Dropbox/priors_data/out_25.Rdat")
   # load(file = "/Users/micha/Desktop/Dropbox/priors_data/outTMP.Rdat")
-  load(file = "/Users/micha/Desktop/Dropbox/priors_data/out.Rdat")
+  load(file = "/Users/micha/Desktop/Dropbox/priors_data/outExp2.Rdat")
 } else {
   # source('main.R')
 }
@@ -161,7 +163,7 @@ plotNumbersDataPaper = ggplot(number_aggr, aes(x = bin, y = value)) + geom_bar(s
   facet_wrap(~ item, scale = "free_y", ncol = 2) + ylab("frequency number choice")
 plotSliderNumberDataPaper = ggplot(slider_aggr, aes(x = bin, y = y_means)) + geom_line() + geom_point() + facet_wrap(~ item, scale = "free_y", ncol = 2) + 
   geom_errorbar(aes(ymin = cilow, ymax = cihigh), width = .5, position = position_dodge(.1), color = 'gray') +
-  ylab("mean slider rating / bin proportion of number choice") + geom_bar(data = number_aggr, aes(x = bin, y = value/20), stat = "identity", fill = "firebrick", alpha = 0.5)
+  ylab("mean slider rating / bin proportion of number choice") + geom_bar(data = number_aggr, aes(x = bin, y = value/nSubj), stat = "identity", fill = "firebrick", alpha = 0.5)
 
 
 # lightning choices
@@ -251,16 +253,16 @@ if (savePlots){
 
 # save plots for CogSciPaper  
 factor = 0.55
-# if (savePlots){
-#   # pdfs
-#   ggsave('../text/01_CogSci_abstract/plots/pop_priors.pdf', pop_priorsPaper, width=10*factor, height = 8*factor)
-# #   ggsave('../text/01_CogSci_abstract/plots/pop_priorsSubj.pdf', pop_priorsSubj, width=10*factor, height = 8*factor)
-# #   ggsave('../text/01_CogSci_abstract/plots/posterior_parameters.pdf', posterior_parameters, width=10*factor, height = 8*factor)
[…]

### 2016-08-07 presentations/01_CogSci_2016/slides.Rmd
CogSci slides

---
title: "What does the crowd believe?"
subtitle: "A hierarchical approach to estimating subjective beliefs"
author: "Michael Franke, Fabian Dablander, Anthea Schöller (Tübingen)"
date : "Erin Bennett, Judith Degen, Michael Henry Tessler, Justine Kao & Noah D. Goodman (Stanford)"
runtime: shiny
output:
  ioslides_presentation:
    css: mistyle.css
    smaller: yes
    transition: faster
---
```{r setup, include=FALSE, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, dev.args = list(bg = 'transparent'), fig.align='center')
require('ggplot2')
require('reshape2')
require('coda')
require('ggmcmc')
require('rjags')
require('runjags')
require('dplyr')
require('gridExtra')
require('rstan')
theme_set(theme_bw() + theme(plot.background=element_blank()) )
```

## overview

<span style = "color:firebrick">recap</span>

- 3 types of Bayesianism:
    - Bayesian versions of classical statistical tests
    - Bayesian modeling of data-generating process
    - cognitive models that assume that people "are Bayesian"

<span style = "color:white"> dummy </span>

<span style = "color:firebrick">today</span>

- subjective beliefs for models of "Bayesian cognition"

## what, where from, where to?

<span style = "color:firebrick">the "priors consortium"</span>

<div style = "float:left; width:45%;">

  - Erin D. Bennett
  - Fabian Dablander
  - Judith Degen
  - Michael Franke
  - Noah Goodman

</div>

<div style = "float:right; width:45%;">

  - Justine Kao
  - Anthea Schoeller
  - Michael Henry Tessler
  - Ciyang Qing
  - ... ??? ...

</div>

<span style = "color:white"> dummy </span>

<span style = "color:white"> dummy </span>

<p>

<div>
<span style = "color:firebrick">disclaimer</span>

- this is mess in progress
- all follies are mine

</div>

# background

## "Bayesian cognition"

<span style = "color:firebrick">Bayes for learning</span>

- prior beliefs on what "structure" can be expected
    - (possibly innate) learning biases
- Bayes rule to learn from sparse input
    - language acquisition 
    - conceptualization, generalization

<span style = "color:firebrick">Bayes for performance</span>

- model task behavior as (partly) Bayesian inference
    - rationalistic explanation, not necessarily mechanism/process
- subject's beliefs about everyday events (task items) plays a role
    - reasoning from premisses, future predictions, etc.
    - language use and interpretation
        - vagueness, non-literal interpretation etc.

## example: vague <span style = "font-style: italic">many</span>

<span style = "color:firebrick">sentence</span>

Joe eats many burgers.

<span style = "color:firebrick">cardinal surprise reading</span>

Joe eats more bugers that we would expect of him.

<span style = "color:firebrick">theory</span>

- look at expectation $p$ about burger consumption for the "kind of guy" that Joe is
    - $p(n)$ is the probability that a guy like Joe eats $n$ burgers
    - let $P$ be the cumulative distribution of $p$
- let $b$ be the number of burgers that Joe actually eats
- "Joe eats many burgers" is true iff $P(b)$ is bigger than a fixed threshold
    - that threshold is the same for all sentences of this kind
    - contextual vagueness from uncertainty about $p$

<div style = "position:absolute; top: 620px; right:60px; color:darkblue;">
  (e.g., Schoeller & Franke 2015)
</div>

## enter subjective beliefs

<span style = "color:firebrick">desideratum</span>

- compact, versatile respresentation of subjective beliefs
- non-expert beliefs about possibly mundane affairs
- cheap, fast and reliable ways of assessment

<span style = "color:firebrick">previous literature</span>

- take real-world frequencies as stand-in for subjective beliefs
- measure beliefs experimentally, e.g.:
    - infer parameters of Gaussian from a judgements of cumulative probability (Manski, 2004)
    - infer parameters and type of distribution from predictions (Tauber & Steyvers 2013)
    - binned slider ratings (Kao et al., 2014)

## binned slider ratings

<div align = 'center'>
  <img src="pics/VQ_prior_elicitation_screen.png" alt="manyPriors" style="width: 800px;"/>
</div>

- discretize domain into bins; let subjects rate each bin
- take relative measure: normalize slider ratings per participant
- average (normalized) slider rating to reflect population-level belief
    - "wisdom of the crowds"

<div style = "position:absolute; top: 620px; right:60px; color:darkblue;">
  (e.g., Schoeller & Franke 2015)
</div>

## agenda

<span style = "color:firebrick">issues</span>

- wanted: reliable & practical way of measuring beliefs
    - consistency across measures
- relation subjects' answers, their beliefs and the population-level aggregate?
    - what of between-subject variance in task data?

<span style = "color:white"> dummy </span>

<span style = "color:firebrick">approach</span>

- belief elicitation with different tasks (within-subject)
- Bayesian hierarchical model to infer latent subjective & "population-level beliefs"
    - "population-level belief" => central tendency of subjective beliefs
    - think: "mean of a Gaussian hyperprior"

# experiments

## overview

- 20 participants recruited via MTurk
- each saw every condition of every task
- 8 items (from previous research)
- 3 task types:
    - slider ratings
    - number estimates
    - binary comparison (a.k.a. "lightning round")

## items

<div style = "position:absolute; top: 110px; left:50px;">

1. "X has just fetched himself a cup of <span style = "color:firebrick">coffee</span> from the office vending machine."
    - "What do you think the temperature of his coffee is?"

2. "X <span style = "color:firebrick">commuted</span> to work yesterday."  
    - "How many minutes do you think she spent commuting yesterday?"

3. "X told a <span style = "color:firebrick">joke</span> to N kids."   
    - "How many of the kids do you think laughed?"

4. "X bought a <span style = "color:firebrick">laptop</span>."         
    - "How much do you think it cost?"

5. "X threw N <span style = "color:firebrick">marbles</span> into a pool."
    - "How many of the marbles do you think sank?"

6. "X just went to the <span style = "color:firebrick">movies</span> to see a blockbuster."
    - "How many minutes long do you think the movie was?"

7. "X watched <span style = "color:firebrick">TV</span> last week."
    - "How many hours do you think he spent watching TV last week?"

8. "X bought a <span style = "color:firebrick">watch</span>."
    - "How much do you think it cost?"

</div>

## slider ratings

<span style = "color:white"> dummy </span>

<div align = 'center'>
  <img src="pics/priors_sliders.png" alt="priorsslider" style="width: 600px;"/>
</div>

## number estimates

<span style = "color:white"> dummy </span>

<div align = 'center'>
  <img src="pics/priors_numbers.png" alt="priorsnumbers" style="width: 600px;"/>
</div>

## binary choices

<span style = "color:white"> dummy </span>

<div align = 'center'>
  <img src="pics/priors_lightning.png" alt="priorslightning" style="width: 600px;"/>
</div>

# results

## slider ratings

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/data_slider.png" alt="dataslider" style="width: 600px;"/>
</div>

## number estimates

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/data_number.png" alt="dataslider" style="width: 600px;"/>
</div>

## lightning round

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/data_choice.png" alt="dataslider" style="width: 600px;"/>
</div>

# model

## questions

<span style = "color:firebrick">general</span>

- is there a consistent population-level average that explains participants choices?
- is the average slider rating a good approximation of a latent population-level average?
- how much individual variance is there?

<span style = "color:white"> dummy </span>

<span style = "color:firebrick">specific (technical)</span>

- hierarchical group-level prior for subjective beliefs?
- link functions for task types?

## slider data

$s_{ijk} \in [0;1]$ - normalized slider ratings of subject $i$ for item $j$ and bin $k$

```{r, echo = FALSE}
source('//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/process_data.R')
# bin_dat <- 
# choice_dat <- read.csv('//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/data/choice_dat.csv')
# number_dat <- read.csv('//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/data/number_dat.csv')
# 
# 
# ## slider rating data
# y.slider = array(0, dim = c(20,8,15))
# for (i in 1:nrow(bin_dat)){
#   y.slider[bin_dat$workerid[i] + 1, 
#            as.numeric(bin_dat$tag)[i],
#            bin_dat$bin_num[i]] = bin_dat$nresponse[i]
# }
# # y.slider = logit(add.margin(y.slider))
# 
# # number choice data
# y.number = array(0, dim = c(20,8,1))
# for (i in 1:nrow(number_dat)){
#   y.number[number_dat$workerid[i] + 1, 
#            as.numeric(number_dat$tag)[i],
#            1] = number_dat$chosen_bin[i]
# }
# y.number = y.number[,,1]
# 
# ## lightning round choice data
# choice_dat = choice_dat %>% mutate(condition = factor(ifelse(chosen_bin > unchosen_bin, unchosen_bin, chosen_bin)))
# choice_dat$condition = as.numeric(choice_dat$condition)
# y.choice = array(0, dim = c(20,8,5))
# higher   = array(0, dim = c(20,8,5))
# lower    = array(0, dim = c(20,8,5))
# for (i in 1:nrow(choice_dat)){
#   y.choice[choice_dat$workerid[i] + 1, 
#            as.numeric(choice_dat$tag)[i],
#            choice_dat$condition[i]] = ifelse(choice_dat$higher_chosen[i] , 1, 0)
# }
```

```{r}
dim(y.slider)
y.slider[1,1:4,1:5]
```

## number data

$n_{ij} \in \{1, \dots, 15\}$ - number of bin for number choice of subject $i$ and item $j$

```{r}
y.number[1:12,,]
```

## lightning choice data

$c_{ijl} \in \{ 0, 1\}$ - whether subject $i$ chose the higher bin in lightning round $l \in \{ 1, \dots, 5\}$ for item $j$

```{r}
y.choice[1,,]
```

## model

<div style = "position:absolute; top: 100px; left:100px;">
  <img src="//Users/micha/Desktop/data/svn/priors/text/01_CogSci_abstract/modelGraph/modelGraph.png" alt="modelGraph" style="width: 700px;"/>
</div>

## hierarchical population prior

- $w \sim \text{Gamma}(2,0.1)$
- $Q_{j}  \sim \text{Dirichlet}(1,\dots, 1)$
- $P_{ij} \sim \text{Dirichlet}(w Q_j)$

<span style = "color:white"> dummy </span>

<div style = "float:left; width:45%;">
`w = 20`
```{r, echo = FALSE, fig.align='center', fig.width=4, fig.height=3}
library('gtools')
p = c(0, 1, 1.4, 2, 4, 4.2, 4.1, 3.7, 4, 3.1, 2, 1.6, 0.8, 0.2)
p = p/sum(p)
ggplot(melt(rdirichlet(alpha= p*20, n = 10)), aes(x = Var2, y = value, fill = factor(Var1), color = "gray")) + geom_line(color = "gray") + geom_line(data = cbind(melt(p), x=1:14, Var1 = "a"), aes(x = x, y = value, color = "red")) + theme(legend.position="none") + xlab("") + ylab("")
```  
</div>
<div style = "float:right; width:45%;">
`w = 200`
```{r, echo = FALSE, fig.align='center', fig.width=4, fig.height=3}
ggplot(melt(rdirichlet(alpha= p*200, n = 10)), aes(x = Var2, y = value, fill = factor(Var1), color = "gray")) + geom_line(color = "gray") + geom_line(data = cbind(melt(p), x=1:14, Var1 = "a"), aes(x = x, y = value, color = "red")) + theme(legend.position="none") + xlab("") + ylab("")
```  
</div>

## link function: sliders

- $\kappa \sim \text{Gamma}(5,5)$
- $\sigma \sim \text{Gamma}(0.0001, 0.0001)$
- $s_{ijk} \sim  \text{logistic}(\text{Norm}(\text{logit}(P_{ijk}), \sigma), \kappa)$

```{r, echo = FALSE}
library(ggplot2)
library(reshape2)

# sequence of subjective probabilities for a given bin and item
pijks = seq(0.0000001,.9999999, length.out = 101)

logit = function(x, k = 1){
  k * log(x / (1-x))
}

logistic = function(x, k = 1, zero = 0) {
  # k is steepness of the curve
  return(1 / (1 + exp(-1/k*(x - zero))))
}

sd = 0.5 # global Gaussian noise
k = 1 # global steepness of logistic transform (has to be 1 in JAGS)
kLogit = 1.2 # global steepness of logit transform
        # (kLogit > k : end-point averse, kLogit < k: end-point loving )

noiseIV = function(x, sd = 0.25, percentile = .25){
  # get percentiles to see how much noise affects choices
  # per default we get the 95% CI about where noise-perturbations would lie
  return(sapply(x, function(y) qnorm(p = c(percentile, 1- percentile), mean = y, sd = sd)))
}

uplow = noiseIV(logit(pijks, kLogit), sd = sd)

plotData = data.frame(p = pijks,
                      lower = logistic(uplow[1,],k),
                      upper = logistic(uplow[2,],k),
                      mean = logistic(logit(pijks, kLogit),k))
plotDataMelted = melt(plotData, id.vars = c("p"))

myPlot = ggplot(plotDataMelted, aes(x = p, y = value, color = variable)) + geom_line() + xlab("P_ijk") + ylab("probability of S_ijk")

sd = 1.5 # global Gaussian noise
k = 1 # global steepness of logistic transform (has to be 1 in JAGS)
kLogit = 2 # global steepness of logit transform
        # (kLogit > k : end-point averse, kLogit < k: end-point loving )

noiseIV = function(x, sd = 0.25, percentile = .25){
  # get percentiles to see how much noise affects choices
  # per default we get the 95% CI about where noise-perturbations would lie
  return(sapply(x, function(y) qnorm(p = c(percentile, 1- percentile), mean = y, sd = sd)))
}

uplow = noiseIV(logit(pijks, kLogit), sd = sd)

plotData = data.frame(p = pijks,
                      lower = logistic(uplow[1,],k),
                      upper = logistic(uplow[2,],k),
                      mean = logistic(logit(pijks, kLogit),k))
plotDataMelted = melt(plotData, id.vars = c("p"))

myPlot2 = ggplot(plotDataMelted, aes(x = p, y = value, color = variable)) + geom_line() + xlab("P_ijk") + ylab("probability of S_ijk")

```

<span style = "color:white"> dummy </span>

<div style = "float:left; width:45%;">
`kappa = 1.2, sigma = 0.5`
```{r, echo = FALSE, fig.align='center', fig.width=4, fig.height=3}
show(myPlot)
```  
</div>
<div style = "float:right; width:45%;">
`kappa = 2, sigma = 1.5`
```{r, echo = FALSE, fig.align='center', fig.width=4, fig.height=3}
show(myPlot2)
```  
</div>

## link function: numbers

- $a \sim \text{Gamma}(2,1)$
- $n_{ij} \sim \text{Categorical}(\text{exp}(a P_{ij}))$

<span style = "color:white"> dummy </span>

```{r, echo = TRUE, eval = FALSE}
qplot(sample(x = 5, size = 1000, replace = TRUE, prob = exp(a *(1:5))))
```

<div style = "float:left; width:45%;">
`a = 0.5`
```{r, echo = FALSE, fig.align='center', fig.width=4, fig.height=3}
qplot(sample(x = 5, size = 1000, replace = TRUE, 
             prob = exp(0.5 *(1:5)))) + 
  xlab("choice") + ylab("frequency")
```
</div>
<div style = "float:right; width:45%;">
`a = 1.5`
```{r, echo = FALSE,fig.align='center', fig.width=4, fig.height=3}
qplot(sample(x = 5, size = 1000, replace = TRUE, 
             prob = exp(1.5 *(1:5)))) + 
  xlab("choice") + ylab("frequency")
```  
</div>

## link function: lightning

- $b \sim \text{Gamma}(2,1)$
- $c_{ijl} \sim \text{Bern}(\text{exp}(b \ (p^{\text{high}}_{ijl}, p^{\text{low}}_{ijl}))$

$$p^\text{high}_{ijl} = \begin{cases} 2 & \text{if $mode(P_{ij})$ is closer to higher} \\
& \text{ bin of $l$ than to lower bin }  \\ 1 & \text{if equal
            distance} \\ 0 & \text{otherwise}  \end{cases}$$

$$p^\text{low}_{ijl} = 2 - p^\text{high}_{ijk}$$

<span style = "color:firebrick">example</span>

- what's more likely: John scored 10-12, or John scored 26-28 points?
- subjective belief says: most likely John scored 14-16
- lower bin is closer, so

## model

<div style = "position:absolute; top: 100px; left:100px;">
  <img src="//Users/micha/Desktop/data/svn/priors/text/01_CogSci_abstract/modelGraph/modelGraph.png" alt="modelGraph" style="width: 700px;"/>
</div>

# Bayesian inference

## set-up

- implemented in JAGS
- 50,000 samples after a burn in of 100,000
- convergence checks: visually and $\hat{R}$

## posterior over parameters

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/posterior_parameters.png" alt="postParameters" style="width: 650px;"/>
</div>

## population-level beliefs $Q_j$

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/pop_priors.png" alt="postPriors" style="width: 580px;"/>
</div>

<div style = "position:absolute; top: 620px; right:60px; color:darkblue;">
  red: averaged normal. slider ratings; black: mean posterior $Q_i$ with 95% HDIs
</div>

## individual vs. population-level beliefs

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/pop_priorsSubj.png" alt="postSubjPriors" style="width: 580px;"/>
</div>

<div style = "position:absolute; top: 620px; right:60px; color:darkblue;">
  black: mean posterior $Q_j$ with 95% HDIs; dark gray: mean posterior $P_{ij}$
</div>

## impressions

<span style = "color:firebrick">from visual inspection</span>

- subjective beliefs differ from population-level mean (good!)
- avrgd normlzd slider ratings nicely approximate mean $Q_j$ (excellent!)

# model criticism

## PPC averaged normalized slider

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/ppc_slider.png" alt="ppcSlider" style="width: 560px;"/>
</div>

avrgd normlzd slider ratings we would expect from the model and the posterior distribution over parameters is virtually indistinguishable from the observed

## PPC number choice

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/ppc_number.png" alt="ppcNumber" style="width: 560px;"/>
</div>

some frequencies of number choices are suprising for the trained model (but: little data to go with; round or salient numbers may play a role)

## PPC lightning round

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/ppc_choice.png" alt="ppcChoice" style="width: 560px;"/>
</div>

miserable failure to predict why it should be more likely that no marble sank than that one marble sank (alternative explanation: subjects revise beliefs, assume homogeneous "wonkiness" of marbles)

## to do

- posterior predictive checks for individual-level $P_{ij}$

- posterior predictive <span style = "font-style: italic">p</span>-value with test statistics:
    - entropy of normalized slider ratings per subject
    - mean of normalized slider ratings per subject

# conclusions

## conclusions

<span style = "color:white"> dummy </span>

- avrgd normlzd slider ratings appear to be practical and reliable measures

<span style = "color:white"> dummy </span>

- slider ratings closely track pop-level central tendency of individual beliefs

<span style = "color:white"> dummy </span>

- hierarchical modeling of population-level beliefs is possible

<span style = "color:white"> dummy </span>

- link functions for task choices seem reliable (enough)

### 2016-08-07 models/04_Exp2_Manski_JAGS/process_data.R

@@ -2,7 +2,7 @@ require(dplyr)
 require('bootstrap')
 library('rjson')
 library('stringr')
-source('helpers/helpers.R')
+source('//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/helpers/helpers.R')
 
 ## for bootstrapping 95% confidence intervals
 theta <- function(x,xdata,na.rm=T) {mean(xdata[x],na.rm=na.rm)}
@@ -11,8 +11,8 @@ ci.low <- function(x,na.rm=T) {
 ci.high <- function(x,na.rm=T) {
   quantile(bootstrap(1:length(x),1000,theta,x,na.rm=na.rm)$thetastar,.975,na.rm=na.rm) - mean(x,na.rm=na.rm)}
 
-d = read.table("data/results.csv", header=T, sep=",",quote="\"")
-bins = read.csv("data/bins.txt",header=T)
+d = read.table("//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/data/results.csv", header=T, sep=",",quote="\"")
+bins = read.csv("//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/data/bins.txt",header=T)
 
 ## number choices (treated as bins)
 give_number = d %>% filter(measure == "give_number") %>% droplevels() %>%

### 2016-08-08 presentations/01_CogSci_2016/slides.Rmd
presentation shortened and polished

@@
 
-## overview
+# motivation
 
-<span style = "color:firebrick">recap</span>
+## point of departure
 
-- 3 types of Bayesianism:
-    - Bayesian versions of classical statistical tests
-    - Bayesian modeling of data-generating process
-    - cognitive models that assume that people "are Bayesian"
+<span style = "color:firebrick">relevance</span>
 
-<span style = "color:white"> dummy </span>
+- belief + intention + rationality $\Rightarrow$ action
 
-<span style = "color:firebrick">today</span>
+<span style = "color:white"> &nbsp; </span>
 
-- subjective beliefs for models of "Bayesian cognition"
+<span style = "color:firebrick">problem</span>
 
-## what, where from, where to?
+- beliefs are latent, only act is directly observable
 
-<span style = "color:firebrick">the "priors consortium"</span>
+<span style = "color:white"> &nbsp; </span>
+<span style = "color:white"> &nbsp; </span>
+<span style = "color:white"> &nbsp; </span>
 
-<div style = "float:left; width:45%;">
+<span style = "color:firebrick">INSERT A COOL PICTURE??</span>
 
-  - Erin D. Bennett
-  - Fabian Dablander
-  - Judith Degen
-  - Michael Franke
-  - Noah Goodman
-
-</div>
-
-<div style = "float:right; width:45%;">
-
-  - Justine Kao
-  - Anthea Schoeller
-  - Michael Henry Tessler
-  - Ciyang Qing
-  - ... ??? ...
-
-</div>
-
-<span style = "color:white"> dummy </span>
-
-<span style = "color:white"> dummy </span>
-
-<p>
-
-<div>
-<span style = "color:firebrick">disclaimer</span>
-
-- this is mess in progress
-- all follies are mine
-
-</div>
-
-# background
-
-## "Bayesian cognition"
-
-<span style = "color:firebrick">Bayes for learning</span>
-
-- prior beliefs on what "structure" can be expected
-    - (possibly innate) learning biases
-- Bayes rule to learn from sparse input
-    - language acquisition 
-    - conceptualization, generalization
-
-<span style = "color:firebrick">Bayes for performance</span>
-
-- model task behavior as (partly) Bayesian inference
-    - rationalistic explanation, not necessarily mechanism/process
-- subject's beliefs about everyday events (task items) plays a role
-    - reasoning from premisses, future predictions, etc.
-    - language use and interpretation
-        - vagueness, non-literal interpretation etc.
-
-## example: vague <span style = "font-style: italic">many</span>
+## mundane beliefs may matter
 
@@
 
+<span style = "color:white"> &nbsp; </span>
+
 <span style = "color:firebrick">cardinal surprise reading</span>
 
-Joe eats more bugers that we would expect of him.
+Joe eats more burgers that we would expect of him.
 
-<span style = "color:firebrick">theory</span>
+<span style = "color:white"> &nbsp; </span>
+<span style = "color:white"> &nbsp; </span>
+<span style = "color:white"> &nbsp; </span>
 
-- look at expectation $p$ about burger consumption for the "kind of guy" that Joe is
-    - $p(n)$ is the probability that a guy like Joe eats $n$ burgers
-    - let $P$ be the cumulative distribution of $p$
-- let $b$ be the number of burgers that Joe actually eats
-- "Joe eats many burgers" is true iff $P(b)$ is bigger than a fixed threshold
-    - that threshold is the same for all sentences of this kind
-    - contextual vagueness from uncertainty about $p$
+<span style = "color:firebrick">INSERT A COOL PICTURE</span>
 
@@
 
-## enter subjective beliefs
+## how to determine beliefs?
 
-<span style = "color:firebrick">desideratum</span>
+<span style = "color:firebrick">real-world frequencies</span>
 
-- compact, versatile respresentation of subjective beliefs
-- non-expert beliefs about possibly mundane affairs
-- cheap, fast and reliable ways of assessment
+- may not exist or may not be known
 
-<span style = "color:firebrick">previous literature</span>
+<span style = "color:white"> &nbsp; </span>
 
-- take real-world frequencies as stand-in for subjective beliefs
-- measure beliefs experimentally, e.g.:
-    - infer parameters of Gaussian from a judgements of cumulative probability (Manski, 2004)
-    - infer parameters and type of distribution from predictions (Tauber & Steyvers 2013)
-    - binned slider ratings (Kao et al., 2014)
+<span style = "color:firebrick">experimental measures</span>
 
-## binned slider ratings
+- give-number task & inference of parameterized distribution 
+    - e.g., Manski (2004); Tauber & Steyvers (2013)
+    - <span style = "font-style: italic">scoring rules</span>
+- binned histogram slider rating task (Kao et al., 2014)
+
+## binned histogram slider ratings
 
@@
 
-- discretize domain into bins; let subjects rate each bin
-- take relative measure: normalize slider ratings per participant
-- average (normalized) slider rating to reflect population-level belief
-    - "wisdom of the crowds"
-
-<div style = "position:absolute; top: 620px; right:60px; color:darkblue;">
-  (e.g., Schoeller & Franke 2015)
-</div>
+take average normalized slider ratings to reflect population-level belief
 
@@
 
-<span style = "color:firebrick">issues</span>
+<span style = "color:firebrick">goal: scrutinize BH task</span>
 
-- wanted: reliable & practical way of measuring beliefs
-    - consistency across measures
-- relation subjects' answers, their beliefs and the population-level aggregate?
-    - what of between-subject variance in task data?
+- relate subjects' answers, their beliefs & population-level aggregate
 
@@
 
-<span style = "color:firebrick">approach</span>
+<span style = "color:firebrick">approach: hieararchical Bayesian modeling</span>
 
-- belief elicitation with different tasks (within-subject)
-- Bayesian hierarchical model to infer latent subjective & "population-level beliefs"
+- data from multiple task types (within-subject)
+- infer latent subjective & "population-level beliefs"
     - "population-level belief" => central tendency of subjective beliefs
@@
 
-# experiments
+# experiment
 
@@
 
-- 20 participants recruited via MTurk
+- 50 participants recruited via MTurk
 - each saw every condition of every task
@@
 - 3 task types:
-    - slider ratings
-    - number estimates
-    - binary comparison (a.k.a. "lightning round")
+    - <span style = "color:firebrick">BH</span>: binned histogram
+    - <span style = "color:firebrick">GAN</span>: give-a-number
+    - <span style = "color:firebrick">PC</span>: paired comparison
 
@@
 
-## slider ratings
+## BH task
 
@@
 <div align = 'center'>
-  <img src="pics/priors_sliders.png" alt="priorsslider" style="width: 600px;"/>
+  <img src="pics/priors_sliders.png" alt="priorsslider" style="width: 750px;"/>
 </div>
 
-## number estimates
+## GAN task
 
@@
 <div align = 'center'>
-  <img src="pics/priors_numbers.png" alt="priorsnumbers" style="width: 600px;"/>
+  <img src="pics/priors_numbers.png" alt="priorsnumbers" style="width: 750px;"/>
 </div>
 
-## binary choices
+## PC task
 
@@
 <div align = 'center'>
-  <img src="pics/priors_lightning.png" alt="priorslightning" style="width: 600px;"/>
+  <img src="pics/priors_lightning.png" alt="priorslightning" style="width: 750px;"/>
 </div>
@@
 <div align = 'center'>
-  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/data_slider.png" alt="dataslider" style="width: 600px;"/>
+  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/data_slider.png" alt="dataslider" style="width: 750px;"/>
 </div>
@@
 <div align = 'center'>
-  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/data_number.png" alt="dataslider" style="width: 600px;"/>
+  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/data_number.png" alt="dataslider" style="width: 750px;"/>
 </div>
@@
 <div align = 'center'>
-  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/data_choice.png" alt="dataslider" style="width: 600px;"/>
+  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/data_choice.png" alt="dataslider" style="width: 750px;"/>
 </div>
 
-# model
+# Bayesian inference
 
@@
 
-- is there a consistent population-level average that explains participants choices?
-- is the average slider rating a good approximation of a latent population-level average?
-- how much individual variance is there?
+- do average BH ratings approximate latent population-level averages well?
+
+<span style = "color:white"> &nbsp; </span>
+
+<span style = "color:firebrick">specific</span>
+
+- relation: choice behavior, subjective & population-level beliefs
+
+```{r, echo = FALSE}
+load('//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/data/processed_data.RData')
+```
+
+## model
+
+<div style = "position:absolute; top: 100px; left:100px;">
+  <img src="//Users/micha/Desktop/data/svn/priors/text/01_CogSci_abstract/modelGraph/modelGraph.png" alt="modelGraph" style="width: 700px;"/>
+</div>
+
+## set-up
+
+- implemented in JAGS
+- 50,000 samples after a burn in of 100,000
+- convergence checks: visually and $\hat{R}$
+
+<!--
+
+## posterior over parameters
+
+<div align = 'center'>
+  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/posterior_parameters.png" alt="postParameters" style="width: 650px;"/>
+</div>
+-->
+
+## population-level beliefs $Q_j$
+
+<div align = 'center'>
+  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/pop_priors.png" alt="postPriors" style="width: 750px;"/>
+</div>
+
+<div style = "position:absolute; top: 620px; right:60px; color:darkblue;">
+  red: averaged normal. slider ratings; black: mean posterior $Q_i$ with 95% HDIs
+</div>
+
+## individual vs. population-level beliefs
+
+<div align = 'center'>
+  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/pop_priorsSubj.png" alt="postSubjPriors" style="width: 750px;"/>
+</div>
+
+<div style = "position:absolute; top: 620px; right:60px; color:darkblue;">
+  black: mean posterior $Q_j$ with 95% HDIs; dark gray: mean posterior $P_{ij}$
+</div>
+
+## upshot
+
+<span style = "color:white"> &nbsp; </span>
+
+- subjective beliefs differ from population-level mean (good!)
+- avrgd normlzd slider ratings reasonably approximate mean $Q_j$ (excellent!)
+
+# model criticism
+
+## PPC averaged normalized slider
+
+<div align = 'center'>
+  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/ppc_slider.png" alt="ppcSlider" style="width: 750px;"/>
+</div>
+
+avrgd normlzd slider ratings we would expect from the model and the posterior distribution over parameters is virtually indistinguishable from the observed
+
+## PPC number choice
+
+<div align = 'center'>
+  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/ppc_number.png" alt="ppcNumber" style="width: 750px;"/>
+</div>
+
+some frequencies of number choices are suprising for the trained model (but: little data to go with; round or salient numbers may play a role)
+
+## PPC lightning round
+
+<div align = 'center'>
+  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/ppc_choice.png" alt="ppcChoice" style="width: 750px;"/>
+</div>
+
+miserable failure to predict why it should be more likely that no marble sank than that one marble sank (alternative explanation: subjects revise beliefs, assume homogeneous "wonkiness" of marbles)
+
+## to do
+
+- posterior predictive checks for individual-level $P_{ij}$
+
+- posterior predictive <span style = "font-style: italic">p</span>-value with test statistics:
+    - entropy of normalized slider ratings per subject
+    - mean of normalized slider ratings per subject
+
+# conclusions
+
+## conclusions
 
@@
 
-<span style = "color:firebrick">specific (technical)</span>
+- avrgd normlzd slider ratings appear to be practical and reliable measures
 
-- hierarchical group-level prior for subjective beliefs?
-- link functions for task types?
+<span style = "color:white"> dummy </span>
 
-## slider data
+- slider ratings closely track pop-level central tendency of individual beliefs
 
-$s_{ijk} \in [0;1]$ - normalized slider ratings of subject $i$ for item $j$ and bin $k$
+<span style = "color:white"> dummy </span>
 
-```{r, echo = FALSE}
-source('//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/process_data.R')
-# bin_dat <- 
-# choice_dat <- read.csv('//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/data/choice_dat.csv')
-# number_dat <- read.csv('//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/data/number_dat.csv')
-# 
-# 
-# ## slider rating data
-# y.slider = array(0, dim = c(20,8,15))
-# for (i in 1:nrow(bin_dat)){
-#   y.slider[bin_dat$workerid[i] + 1, 
-#            as.numeric(bin_dat$tag)[i],
-#            bin_dat$bin_num[i]] = bin_dat$nresponse[i]
-# }
-# # y.slider = logit(add.margin(y.slider))
-# 
-# # number choice data
-# y.number = array(0, dim = c(20,8,1))
-# for (i in 1:nrow(number_dat)){
-#   y.number[number_dat$workerid[i] + 1, 
-#            as.numeric(number_dat$tag)[i],
-#            1] = number_dat$chosen_bin[i]
-# }
-# y.number = y.number[,,1]
-# 
-# ## lightning round choice data
-# choice_dat = choice_dat %>% mutate(condition = factor(ifelse(chosen_bin > unchosen_bin, unchosen_bin, chosen_bin)))
-# choice_dat$condition = as.numeric(choice_dat$condition)
-# y.choice = array(0, dim = c(20,8,5))
-# higher   = array(0, dim = c(20,8,5))
-# lower    = array(0, dim = c(20,8,5))
-# for (i in 1:nrow(choice_dat)){
-#   y.choice[choice_dat$workerid[i] + 1, 
-#            as.numeric(choice_dat$tag)[i],
-#            choice_dat$condition[i]] = ifelse(choice_dat$higher_chosen[i] , 1, 0)
-# }
-```
+- hierarchical modeling of population-level beliefs is possible
 
-```{r}
-dim(y.slider)
-y.slider[1,1:4,1:5]
-```
+<span style = "color:white"> dummy </span>
 
-## number data
+- link functions for task choices seem reliable (enough)
 
-$n_{ij} \in \{1, \dots, 15\}$ - number of bin for number choice of subject $i$ and item $j$
-
-```{r}
-y.number[1:12,,]
-```
-
-## lightning choice data
-
-$c_{ijl} \in \{ 0, 1\}$ - whether subject $i$ chose the higher bin in lightning round $l \in \{ 1, \dots, 5\}$ for item $j$
-
-```{r}
-y.choice[1,,]
-```
+# modeling details
 
@@
 - subjective belief says: most likely John scored 14-16
-- lower bin is closer, so
-
-## model
-
-<div style = "position:absolute; top: 100px; left:100px;">
-  <img src="//Users/micha/Desktop/data/svn/priors/text/01_CogSci_abstract/modelGraph/modelGraph.png" alt="modelGraph" style="width: 700px;"/>
-</div>
-
-# Bayesian inference
-
-## set-up
-
-- implemented in JAGS
-- 50,000 samples after a burn in of 100,000
-- convergence checks: visually and $\hat{R}$
-
-## posterior over parameters
-
-<div align = 'center'>
-  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/posterior_parameters.png" alt="postParameters" style="width: 650px;"/>
-</div>
-
-## population-level beliefs $Q_j$
-
-<div align = 'center'>
-  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/pop_priors.png" alt="postPriors" style="width: 580px;"/>
-</div>
-
-<div style = "position:absolute; top: 620px; right:60px; color:darkblue;">
-  red: averaged normal. slider ratings; black: mean posterior $Q_i$ with 95% HDIs
-</div>
-
-## individual vs. population-level beliefs
-
-<div align = 'center'>
-  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/pop_priorsSubj.png" alt="postSubjPriors" style="width: 580px;"/>
-</div>
-
-<div style = "position:absolute; top: 620px; right:60px; color:darkblue;">
-  black: mean posterior $Q_j$ with 95% HDIs; dark gray: mean posterior $P_{ij}$
-</div>
-
-## impressions
-
-<span style = "color:firebrick">from visual inspection</span>
-
-- subjective beliefs differ from population-level mean (good!)
-- avrgd normlzd slider ratings nicely approximate mean $Q_j$ (excellent!)
-
-# model criticism
-
-## PPC averaged normalized slider
-
-<div align = 'center'>
-  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/ppc_slider.png" alt="ppcSlider" style="width: 560px;"/>
-</div>
-
-avrgd normlzd slider ratings we would expect from the model and the posterior distribution over parameters is virtually indistinguishable from the observed
-
-## PPC number choice
-
-<div align = 'center'>
-  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/ppc_number.png" alt="ppcNumber" style="width: 560px;"/>
-</div>
-
-some frequencies of number choices are suprising for the trained model (but: little data to go with; round or salient numbers may play a role)
-
-## PPC lightning round
-
-<div align = 'center'>
-  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/ppc_choice.png" alt="ppcChoice" style="width: 560px;"/>
-</div>
-
-miserable failure to predict why it should be more likely that no marble sank than that one marble sank (alternative explanation: subjects revise beliefs, assume homogeneous "wonkiness" of marbles)
-
-## to do
-
-- posterior predictive checks for individual-level $P_{ij}$
-
-- posterior predictive <span style = "font-style: italic">p</span>-value with test statistics:
-    - entropy of normalized slider ratings per subject
-    - mean of normalized slider ratings per subject
-
-# conclusions
-
-## conclusions
-
-<span style = "color:white"> dummy </span>
-
-- avrgd normlzd slider ratings appear to be practical and reliable measures
-
-<span style = "color:white"> dummy </span>
-
-- slider ratings closely track pop-level central tendency of individual beliefs
-
-<span style = "color:white"> dummy </span>
-
-- hierarchical modeling of population-level beliefs is possible
-
-<span style = "color:white"> dummy </span>
-
-- link functions for task choices seem reliable (enough)
+- lower bin is closer, so it's more likely to be chosen

### 2016-08-08 models/04_Exp2_Manski_JAGS/plotting.R

@@ -13,8 +13,6 @@ savePlots = TRUE
 nSubj = 50
 
 if (readFlag){
-  # load(file = "/Users/micha/Desktop/Dropbox/priors_data/out_25.Rdat")
-  # load(file = "/Users/micha/Desktop/Dropbox/priors_data/outTMP.Rdat")
   load(file = "/Users/micha/Desktop/Dropbox/priors_data/outExp2.Rdat")
 } else {
   # source('main.R')
@@ -40,7 +38,7 @@ meansIP = tbl_df(melt(out$sims.list$item.pop)) %>%
     min = HDIofMCMC(value)[1]
   )
 meansIP$y_emp = y.slider_means$mymean
-pop_priors = ggplot(meansIP, aes(x = bin, y = mean)) + geom_line() + geom_point() + facet_wrap(~ item, scale = "free") +
+pop_priors = ggplot(meansIP, aes(x = bin, y = mean)) + geom_line() + geom_point() + facet_wrap(~ item, nrow = 2, scale = "free") +
   geom_ribbon(aes(ymin=min, ymax=max), fill="gray", alpha="0.5") +
   geom_line(aes(x = bin, y = y_emp) , color = "firebrick")  + geom_point(aes(x = bin, y = y_emp ), color = "firebrick") +
   ylab(" mean posterior Q_{ijk} / average slider rating")
@@ -70,7 +68,7 @@ meansIPSubj$maxPop = meansIP$max
 pop_priorsSubj = ggplot(meansIPSubj, aes(x = bin, y = mean, group = subject)) + 
   geom_ribbon(aes(ymin=minPop, ymax=maxPop), fill="lightgray", alpha=0.5) +
   geom_line(color="darkgray", size = 0.5, alpha = 0.7)  +
-  facet_wrap(~ item, scale = "free") +
+  facet_wrap(~ item, nrow = 2, scale = "free") +
   # geom_line(aes(x = bin, y = y_emp) , color = "firebrick")  + geom_point(aes(x = bin, y = y_emp ), color = "firebrick") +
   geom_line(aes(x = bin, y = meanPop) , color = "black", alpha = 0.7)
 show(pop_priorsSubj)
@@ -111,7 +109,7 @@ slider_aggr = slider_aggr[order(slider_aggr$item,slider_aggr$bin),]
 slider_aggr$y_means = y.slider_means$mymean
 slider_aggr$cilow = y.slider_means$cilow
 slider_aggr$cihigh = y.slider_means$cihigh
[…]

### 2016-08-08 models/04_Exp2_Manski_JAGS/postprocessing.R

@@ -11,8 +11,7 @@ theme_set(theme_bw() + theme(plot.background=element_blank()) )
 readFlag = TRUE
 
 if (readFlag){
-  # load(file = "/Users/micha/Desktop/Dropbox/priors_data/out_25.Rdat")
-  load(file = "/Users/micha/Desktop/Dropbox/priors_data/out.Rdat")
+  load(file = "/Users/micha/Desktop/Dropbox/priors_data/outExp2.Rdat")
 } else {
   # source('main.R')
 }
@@ -51,8 +50,8 @@ y.emp = logistic(y.slider)
 
 posteriorMeans = array(0, dim = c(dim(y.rep)[1], dim(y.rep)[2], dim(y.rep)[3]))
 posteriorEntrs = array(0, dim = c(dim(y.rep)[1], dim(y.rep)[2], dim(y.rep)[3]))
-pValueMean = matrix(0, nrow = 20, ncol = 8)
-pValueEntr = matrix(0, nrow = 20, ncol = 8)
+pValueMean = matrix(0, nrow = 50, ncol = 8)
+pValueEntr = matrix(0, nrow = 50, ncol = 8)
 
 for (i in 1:(dim(y.rep)[2])) {
   for (j in 1:(dim(y.rep)[3])) {
@@ -73,7 +72,7 @@ for (i in 1:(dim(y.rep)[2])) {
 
 ## two failures for means: two y.repects who gave "extreme" ratings for marble case fall out!!
 ## check empirical means for marbles:
-sapply(1:20, function(i) sum(1:15*y.emp[i,5,]))
+sapply(1:50, function(i) sum(1:15*y.emp[i,5,]))
 
 ## a few failures for "entropy: the two "extreme" cases plus another 18 cases of (almost) flat input
 

### 2016-08-08 models/04_Exp2_Manski_JAGS/process_data.R

@@ -137,3 +137,5 @@ for (i in 1:nrow(choice_dat)){
         as.numeric(choice_dat$tag)[i],
         choice_dat$condition[i]] = min(choice_dat$chosen_bin[i], choice_dat$unchosen_bin[i])
 }
+
+save(y.slider, y.slider_means, y.choice, y.number, file = "data/processed_data.RData")

### 2016-08-09 presentations/01_CogSci_2016/slides.Rmd
CogSci slides update; posterir P-value plots

@@
 
-## to do
+## posterior predictive p-values
 
-- posterior predictive checks for individual-level $P_{ij}$
-
-- posterior predictive <span style = "font-style: italic">p</span>-value with test statistics:
-    - entropy of normalized slider ratings per subject
-    - mean of normalized slider ratings per subject
+<div align = 'center'>
+  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/posteriorPvalues_individualSliders.png" alt="posteriorP" style="width: 750px;"/>
+</div>
 

### 2016-08-09 experiments/exp2_Manski/results/exp_manski.R

@@ -1,5 +1,9 @@
 library(ggplot2)
 library(rjson)
+library(tidyr)
+library(dplyr)
+library(lme4)
+library(lmerTest)
 # setwd("/Users/titlis/cogsci/projects/stanford/projects/priors/experiments/exp2_Manski/results")
 source("helpers.R")
 

### 2016-08-09 models/04_Exp2_Manski_JAGS/postprocessing.R

@@ -2,13 +2,15 @@ library('coda')
 library('ggmcmc')
 library('jagsUI') # for parallel computing
 library('xtable')
+library(grid)
+library(gridExtra)
 source('helpers/helpers.R')
 # source('~/Desktop/data/svn/ProComPrag/dev_tmp/typicality_quantifiers/model/helpers.r')
 source('process_data.R')
 
 theme_set(theme_bw() + theme(plot.background=element_blank()) )
 
-readFlag = TRUE
+readFlag = FALSE
 
 if (readFlag){
   load(file = "/Users/micha/Desktop/Dropbox/priors_data/outExp2.Rdat")
@@ -37,9 +39,6 @@ statsSum = data.frame(
 
 show(xtable(t(statsSum)))
   
-
-
-
 
 entropy = function(p){
   - sum(p*log(p))
@@ -78,3 +77,81 @@ sapply(1:50, function(i) sum(1:15*y.emp[i,5,]))
 
 show(pValueMean < 0.05)
 show(pValueEntr < 0.05)
+
+## make data frame with posterior p-value violations
+
+postPs = melt(pValueMean) %>% rename(subject = Var1, item = Var2, pMean = value) %>%
+  mutate(pMeanViol = as.numeric(pMean <= 0.05),
+         item = levels(factor(dimnames(y.emp)[[2]]))[item])
+postPsTMP = melt(pValueEntr) %>% rename(subject = Var1, item = Var2, pEntr = value) %>%
+  mutate(pEntrViol = as.numeric(pEntr <= 0.05))
+postPs$pEntr = postPsTMP$pEntr
+postPs$pEntrViol = postPsTMP$pEntrViol
+postPs$viol = paste0(as.character(postPs$pMeanViol), as.character(postPs$pEntrViol))
+postPs$viol = ifelse(postPs$viol == "00", "none", 
+                     ifelse(postPs$viol == "10", "mean", 
+                            ifelse(postPs$viol == "01", "entropy", "both")))
+postPs$bin = 1
+postPs$rating_scaled = 0.5
+
+
+## plot individual-level slider ratings
+
+rescale = function(x) {
+  if (max(x) == min(x)) {return(rep(0.5, length(x)))}
+  (x-min(x)) / (max(x) - min(x))
+}
+
+y.slider_individual = melt(y.emp) %>% rename(subject = Var1, item = Var2, bin = Var3, rating = value) %>%
[…]

### 2016-08-11 presentations/01_CogSci_2016/slides.Rmd
CogSci slides makeover

@@
 author: "Michael Franke, Fabian Dablander, Anthea Schöller (Tübingen)"
-date : "Erin Bennett, Judith Degen, Michael Henry Tessler, Justine Kao & Noah D. Goodman (Stanford)"
+date : "Erin Bennett, Judith Degen, MH Tessler, Justine Kao & Noah D. Goodman (Stanford)"
 runtime: shiny
@@
 
-## point of departure
+## outline
+
+<span style = "color:firebrick">Bayesian models of cognition</span>
+
+- may require specifications of subjective beliefs
+
+<span style = "color:white"> &nbsp; </span>
+
+<span style = "color:firebrick">New experimental measure</span>
+
+- binned histogram task
+
+<span style = "color:white"> &nbsp; </span>
+
+<span style = "color:firebrick">Bayesian data analysis</span>
+
+- infer population-level and subjective beliefs from data
+
+## beliefs are latent
 
@@
 
-<span style = "color:white"> &nbsp; </span>
-
 <span style = "color:firebrick">problem</span>
 
-- beliefs are latent, only act is directly observable
-
-<span style = "color:white"> &nbsp; </span>
-<span style = "color:white"> &nbsp; </span>
-<span style = "color:white"> &nbsp; </span>
-
-<span style = "color:firebrick">INSERT A COOL PICTURE??</span>
-
-## mundane beliefs may matter
-
-<span style = "color:firebrick">sentence</span>
-
-Joe eats many burgers.
+- beliefs are latent, only action is directly observable
 
@@
 
+<div align = 'center'>
+  <img src="pics/jones_umbrella.jpg" alt="JonesUmbrella" style="width: 600px;"/>
+</div>
+
+## focus: 1d continguous quantities
+
+<div style = "float:left; width:45%;">
+<span style = "color:firebrick">sentence</span>
+
+Joe ate many burgers.
+
+<span style = "color:white"> &nbsp; </span>
+</div>
+<div style = "float:right; width:45%;">
 <span style = "color:firebrick">cardinal surprise reading</span>
 
-Joe eats more burgers that we would expect of him.
+Joe ate more than expected.
 
 <span style = "color:white"> &nbsp; </span>
-<span style = "color:white"> &nbsp; </span>
+</div>
+
 <span style = "color:white"> &nbsp; </span>
 
-<span style = "color:firebrick">INSERT A COOL PICTURE</span>
+<div style = "float:left; width:45%;">
+```{r, echo = FALSE, fig.align='left', fig.width=3.5, fig.height=2.7}
+bb = function(n,a,b) {
+  sapply(0:n, function(k) choose(n,k) * beta(k+a,n-k+b) / beta(a,b))
+}
+theta = 0.8
+MyDF<-data.frame(x = 0:25,
+                 y = bb(25, 3, 10))
+cutoff = max(which(cumsum(MyDF$y) <= theta)) - 1
+p <-ggplot(MyDF, aes(x = x, y = y)) + geom_bar(stat = "identity", fill = "gray") +
+  # geom_bar(data = filter(MyDF, x <= cutoff), stat = "identity", fill = "firebrick") +
+  xlab("n") + ylab("") + 
+  ggtitle("P(Joe ate n)")
+show(p)
+```  
+</div>
+<div style = "float:right; width:45%;">
+```{r, echo = FALSE, fig.align='left', fig.width=3.5, fig.height=2.7}
+MyDF2 = data.frame(x = 0:25,
+                   y = cumsum(bb(25, 3, 10)))
+cutoff = max(which(MyDF2$y <= theta)) - 1
+p2 <-ggplot(MyDF2, aes(x = x, y = y)) + geom_bar(stat = "identity", fill = "gray") +
+  geom_bar(data = filter(MyDF2, x <= cutoff), stat = "identity", fill = "firebrick") +
+  xlab("n") + ylab("") +
+  geom_hline(aes(yintercept=theta), color = "firebrick") +
+  annotate("text", x = 3, y = 0.85, label = "theta = 0.8") +
+  ggtitle("P(Joe ate n or less)")
+show(p2)
+```  
+</div>
 
@@
 
-- may not exist or may not be known
+- may not exist
+- subjects may not be aware
 
@@
     - e.g., Manski (2004); Tauber & Steyvers (2013)
-    - <span style = "font-style: italic">scoring rules</span>
-- binned histogram slider rating task (Kao et al., 2014)
+    - <span style = "font-style: italic">scoring rules</span> (Savage, 1971)
+- iterated learning tasks (Lewandowsky, Griffiths, & Kalish, 2009)
+- <span style = "color:firebrick">binned histogram task</span> (Kao et al., 2014)
 
-## binned histogram slider ratings
+<!--
+## BH task
 
@@
 
-## agenda
-
-<span style = "color:firebrick">goal: scrutinize BH task</span>
-
-- relate subjects' answers, their beliefs & population-level aggregate
-
-<span style = "color:white"> dummy </span>
-
-<span style = "color:firebrick">approach: hieararchical Bayesian modeling</span>
-
-- data from multiple task types (within-subject)
-- infer latent subjective & "population-level beliefs"
-    - "population-level belief" => central tendency of subjective beliefs
-    - think: "mean of a Gaussian hyperprior"
+-->
 
@@
 
-## questions
+## agenda
 
-<span style = "color:firebrick">general</span>
+<span style = "color:firebrick">goal: scrutinize BH task</span>
 
@@
 
-<span style = "color:white"> &nbsp; </span>
+<span style = "color:white"> dummy </span>
 
-<span style = "color:firebrick">specific</span>
+<span style = "color:firebrick">approach: hieararchical Bayesian modeling</span>
 
-- relation: choice behavior, subjective & population-level beliefs
+- take data from all three tasks
+- infer latent subjective & "population-level beliefs"
+
+<span style = "color:white"> dummy </span>
+
+<span style = "color:firebrick">main technical questions</span>
+
+- how do individual-level beliefs relate to population level beliefs?
+    - "population-level belief" => central tendency of subjective beliefs
+    - think: "mean of a Gaussian hyperprior"
+- how do subjective beliefs give rise to observable choices in the three tasks?
 
@@
 
-<!--
-
-## posterior over parameters
-
-<div align = 'center'>
-  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/posterior_parameters.png" alt="postParameters" style="width: 650px;"/>
-</div>
--->
-
 ## population-level beliefs $Q_j$
@@
 
-<div style = "position:absolute; top: 620px; right:60px; color:darkblue;">
-  red: averaged normal. slider ratings; black: mean posterior $Q_i$ with 95% HDIs
+<div style = "position:absolute; top: 620px; right:60px;">
+<span style = "color:firebrick">red</span>: BH task averages; <span style = "color:black">black</span>: mean posterior $Q_i$; <span style = "color:grey">grey</span>: 95% HDIs
 </div>
@@
 
-<div style = "position:absolute; top: 620px; right:60px; color:darkblue;">
-  black: mean posterior $Q_j$ with 95% HDIs; dark gray: mean posterior $P_{ij}$
+<div style = "position:absolute; top: 620px; right:60px;">
+<span style = "color:black">black</span>: mean posterior $Q_i$; <span style = "color:grey">grey</span>: mean posterior $P_{ij}$
 </div>
@@
 
-- subjective beliefs differ from population-level mean (good!)
-- avrgd normlzd slider ratings reasonably approximate mean $Q_j$ (excellent!)
+<span style = "color:white"> &nbsp; </span>
+
+- subjective beliefs $P_{ij}$ may differ from population-level mean (good!)
+
+<span style = "color:white"> &nbsp; </span>
+
+- BH task averages reasonably approximate mean $Q_j$ (excellent!)
 
@@
 
-avrgd normlzd slider ratings we would expect from the model and the posterior distribution over parameters is virtually indistinguishable from the observed
+<div style = "position:absolute; top: 620px; right:60px;">
+<span style = "color:firebrick">red</span>: observed data; <span style = "color:black">black</span>: mean posterior prediction with 95% HDIs
+</div>
 
@@
 
-some frequencies of number choices are suprising for the trained model (but: little data to go with; round or salient numbers may play a role)
+<div style = "position:absolute; top: 620px; right:60px;">
+<span style = "color:firebrick">red</span>: observed data; <span style = "color:black">black</span>: mean posterior prediction with 95% HDIs
+</div>
 
@@
 
-miserable failure to predict why it should be more likely that no marble sank than that one marble sank (alternative explanation: subjects revise beliefs, assume homogeneous "wonkiness" of marbles)
-
-## posterior predictive p-values
-
-<div align = 'center'>
-  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/posteriorPvalues_individualSliders.png" alt="posteriorP" style="width: 750px;"/>
+<div style = "position:absolute; top: 620px; right:60px;">
+<span style = "color:firebrick">red</span>: observed data; <span style = "color:black">black</span>: mean posterior prediction with 95% HDIs
 </div>
@@
 
-- avrgd normlzd slider ratings appear to be practical and reliable measures
+<span style = "color:firebrick">general</span>
+
+- BH task averages are practical and reasonable approximations of the popuation-level central tendency of individual beliefs
+- Bayesian data analysis is a great tool to extract information from multiple sources (e.g., data from different task types)
 
@@
 
-- slider ratings closely track pop-level central tendency of individual beliefs
+<span style = "color:firebrick">specific</span>
 
-<span style = "color:white"> dummy </span>
-
-- hierarchical modeling of population-level beliefs is possible
-
-<span style = "color:white"> dummy </span>
-
+- hierarchical modeling of population-level beliefs seems possible
 - link functions for task choices seem reliable (enough)
@@
 
-## link function: sliders
+## link function: BH task
 
@@
 
-## link function: numbers
+## link function: GAN task
 
@@
 
-## link function: lightning
+## link function: PC task
 
@@
 - lower bin is closer, so it's more likely to be chosen
+
+## posterior over parameters
+
+<div align = 'center'>
+  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/posterior_parameters.png" alt="postParameters" style="width: 650px;"/>
+</div>
+
+## posterior predictive p-values
+
+<div align = 'center'>
+  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/posteriorPvalues_individualSliders.png" alt="posteriorP" style="width: 750px;"/>
+</div>

### 2016-08-11 models/04_Exp2_Manski_JAGS/plotting.R

@@ -264,3 +264,4 @@ if (savePlots){
   ggsave('plots/paper/data_number.pdf', plotNumbersDataPaper,  width=10*factor, height = 8*factor)
   ggsave('plots/paper/data_choice.pdf', plotChoicesDataPaper, width=10*factor, height = 8*factor)
 }
+

### 2016-08-11 presentations/01_CogSci_2016/slides.Rmd
CogSci slides final make-over

@@
 
-# motivation
-
 ## outline
@@
 
-<span style = "color:firebrick">New experimental measure</span>
+<span style = "color:firebrick">new experimental measure</span>
 
-- binned histogram task
+- binned histogram task (Kao et al., 2014)
 
@@
 
-- infer population-level and subjective beliefs from data
+- great tool to specify and test ideas about how unobservables (like beliefs) relate to each other and how they generate observable data
 
@@
 
-- belief + intention + rationality $\Rightarrow$ action
+- belief + desire + rationality $\Rightarrow$ action
 
@@
 ## focus: 1d continguous quantities
+
+<span style = "color:firebrick">context</span>
+
+Joe is an adult male from Texas. He went to a BBQ party yesterday.
 
@@
 
+<!--
+
 <div style = "position:absolute; top: 620px; right:60px; color:darkblue;">
@@
 
+-->
 ## how to determine beliefs?
@@
 
-<span style = "color:white"> dummy </span>
-
 <div align = 'center'>
@@
 </div>
+
+<span style = "color:white"> dummy </span>
+
+normalize slider rating by subject $\Rightarrow$ average over subjects $\Rightarrow$ <span style = "color:firebrick">BH task averages</span>
 
@@
 
-## slider ratings
+## BH task averages
 
@@
 
-## number estimates
+## GAN frequencies
 
@@
 
-## lightning round
+## PC choice proportions
 
@@
 
-- do average BH ratings approximate latent population-level averages well?
+- do BH task averages approximate the central tendency of beliefs in the population?
 
@@
 - infer latent subjective & "population-level beliefs"
+- specify "link functions"
+    - how do subjective beliefs generate observable data
 
 <span style = "color:white"> dummy </span>
+
+<!--
 
@@
 - how do subjective beliefs give rise to observable choices in the three tasks?
+
+-->
 
@@
 
-<span style = "color:firebrick">general</span>
-
-- BH task averages are practical and reasonable approximations of the popuation-level central tendency of individual beliefs
-- Bayesian data analysis is a great tool to extract information from multiple sources (e.g., data from different task types)
+- BH task averages are practical and reasonable approximations of the population-level central tendency of individual beliefs
 
@@
 
-<span style = "color:firebrick">specific</span>
-
-- hierarchical modeling of population-level beliefs seems possible
-- link functions for task choices seem reliable (enough)
+- Bayesian data analysis is a great tool to specify and test ideas about how unobservables (like beliefs) relate to each other and how they generate observable data
+    - inferences based on diverse data from different tasks
 
